### 2016-06-16 paper.tex
initial commit moving this into git

\documentclass[a4paper]{article}

\usepackage{microtype}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{acronym}
\acrodef{MDM}{mobile device management}
\acrodef{BYOD}{bring your own device}

\usepackage{algorithm}
\usepackage{algpseudocode}

% AppPAL and Listing formatting
\usepackage{listings}
\lstdefinelanguage{AppPAL}{%
  morekeywords={if,inf,says,where,true,false},
  otherkeywords={can-act-as,can-say},
  sensitive=true,
  morestring=[b]',
  literate={\ inf\ }{{$\infty$}}5
}[keywords,strings]
\lstset{%
  basicstyle=\ttfamily\footnotesize{},
  stringstyle=\sffamily\footnotesize{},
  keywordstyle=\ttfamily\slshape\footnotesize{},
  language=AppPAL,
  columns=flexible,
  frame=single,
  framesep=0.5pt,
  framexleftmargin=2.5pt,
  framexrightmargin=2.5pt,
}

\newcommand{\dotdotdot}[1]{#1}

\title{Case Study Formalising Mobile BYOD Policies with AppPAL}
\author{Joseph Hallett and David Aspinall}
\begin{document}
\maketitle

% 4 sentences
\begin{abstract}
  \begin{itemize}
  \item BYOD policies written infromally, with no real means to enforce.
  \item Created standard language by analyzing \emph{X} real policies, using AppPAL
  \item Using formal methods we found \emph{Y} problems.
  \item AppPAL is a helpful language for thinking about policies in. 
  \end{itemize}
\end{abstract}

% 1 page
\section{Introduction}
\label{sec:intro}
% 1. Describe the problem
% 2. State your contributions
% ...and that is all

Employees bring their personal devices to work.
This creates a problem for corporate IT departments:
  they want to secure their company's networks, but they have limited control over the personal devices running on them.
To ensure compliance IT departments pubilsh mobile device polices.
A recent survey found that around 70\% of companies published some form of mobile device policy~\cite{schulze_byod_2016}.
Employees must agree to follow these policies if they want to use their personal devices at work.

Often the policies are specified with natural language.
This makes compliance hard as the policies can contain ambiguities, contradictions and can change without the user's noticing there is a new policy to follow.
To solve this some companies offer \ac{MDM} software that can help enforce some of the policies, however their capabilities are usually limited to banning specific apps and configuring email and filesharing services.
Even when companies have \ac{MDM} software a survey found that over 50\% of companies have one or more devices on their network that did not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.

Formal languages can help to remove ambiguities in policies and identify when policies contain problems.
We show how the AppPAL language~\cite{hallett_apppal_2016} can be used to formalise \ac{BYOD} policies, and present a standard set of predicates and idioms from translating real \ac{BYOD} policies from the SANS institute and NHS.
By translating these policies into formal language we were able to identify cases where the policy rule were duplicated, and underspecified.

In this paper we present excerpts from the AppPAL translations of two BYOD policies.
In particular we focus on the translations of a policy from an NHS Trust hospital, and from an exemplar policy from the SANS institute designed to help companies devise their own policies.  These policies are more complex than a typical BYOD policies as they have rules which go beyond white and black listing.  We illustrate how AppPAL's checking methods can identify when the policies are under-specified and when the policies contain redundency, either through policies being duplicated in the documents or by policies with simplified rules in the same document.

% 1 page
\section{The Problem}
\label{sec:problem}

Consider a policy for describing what apps can be installed on a mobile device for use at work.
One approach might be to provide a list of authorized and unauthorized apps.
Employees should cosult this list before running any apps, and any \ac{MDM} software ought to enforce the list automatically.
Both the SANS and NHS policies give requirements implementing these policies.
In the SANS policy the IT department is responsible for maintaining the lists, and in the NHS policy (which only uses whitelisting) an app is only approved if the Care and Clinical Policy group approves it as well as another group depending on whether it is a clinical or non-clinical app.

\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
``The IT Department maintains a list of allowed and unauthorised applications and makes them available to users on the intranet.''}]
'sans-policy' says 'it-department' can-say App isInstallable.
\end{lstlisting}
\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape ``%
Apps for work usage must not be downloaded onto corporately issued mobile devices (even if approved on the NHS apps store) unless they have been approved through the following Trust channels: 
    Clinical apps;
        at the time of writing there are no apps clinically approved by the Trust for use with patients/clients. 
        However, if a member of staff believes that there are clinical apps \dotdotdot{or other technologies that could benefit their patients/clients, this should be discussed with the clinical lead in the first instance and} ratification should be sought via the Care and Clinical Policies Group. 
        \dotdotdot{A clinical app should not be used if it has not been approved via this group.}
    Business apps; 
        at the time of writing there are no business (i.e., non-clinical) apps approved by the Trust for use other than those preloaded onto the device at the point of issue. 
        However, if a member of staff believes that there are apps \dotdotdot{or other technologies that could benefit their non-clinical work,} ratification of the app must be sought via the Management of Information Group (MIG). 
        \dotdotdot{An app should not be used if it has not been approved via this group.}
Following approval through Care and Clinical Policies and/or MIG, final approval will be required through Integrated Governance Committee.''}]
'nhs-policy' says App isInstallable if App isUsable, App isFinallyApproved.
'nhs-policy' says 'cacpg' can-say App isUsable if App isUsableClinically.
'nhs-policy' says 'mig' can-say App isUsable if App isUsableNonClinically.
'nhs-policy' says 'igc' can-say App isFinallyApproved.
\end{lstlisting}

In the case of the NHS requirement the entire requirement seems redundant as it states there no apps have been approved.
What the rule gives us, and what is made explicit in the formalization is who would need to approve the app if we were going to get an approved one.
As well as providing improvements to readability, the formalization in AppPAL makes explicit who has approved each app.
It also helps us spot when a rule is duplicated, for instance the SANS policy later has a requirement that:

\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
``Only approved third party applications can be
installed on handhelds. The approved list can be
obtained by contacting the IT department, or should be
available on the intranet.''}]
'sans-policy' says 'it-department' can-say App isInstallable.
\end{lstlisting}

When reading the natural language requirement it might not be immediately obvious that there is duplication, but with the formalization it is immediately apparent.  This suggests that the policy could be improved by removing, or combining the requirements to remove ambiguity and prevent problems later if the requirement was to be updated.

% 2 pages
\section{My Idea}
\label{sec:idea}

Writing policies in a formal language gives us two major benefits. 
Firstly it gives us a precise framework for describing what each policy actually means and how it can be satisfied and allowing comparison between policies from different sources.  Secondly it allows the use of policy checking tools that can infer when a policy may have problems.

Translating natural language policies into fromal language is not trivial, however.
Care should be taken to preserve not only the meaning of the policy but also the spirit and style of the original policy.
Not all policies are specified in the same manner however. 
Both the SANS and NHS policies, for example, specify when it is appropriate to use the camera.
The SANS policy tends toward being prohibitive, describing when features should be disabled.
\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape
``Digital camera embedded on handheld devices might be
disabled in restricted environments, according to Company risk analysis.}]
'company' says Device mustDisable('camera') 
  if 'restricted-environment' isWhereIs(Device).
\end{lstlisting}
The NHS policy is expressed differently describing when it is appropriate to use the camera.
\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape
  ``Some mobile devices have the ability to take photographs~/~videos. This
  function should not be used for photographs~/~videos of an individual's
  care and treatment unless the device has encryption enabled and it is
  clinically appropriate to do so.''}]
'nhs-trust' says Device canPhotograph(Patient)
  if Device isEncrypted, Patient canBePhotographed.
\end{lstlisting}

Both approaches can be expressed using AppPAL, and even if the two
approaches were mixed into the same policy in this case it would not
cause problems from an enforcement perspective: a photo could not be
taken if the camera were disabled even if the the user had the
authorization.  In general though this can lead to confilicts, which
will need to be resolved.  When using AppPAL as a tool for comparison
and thinking about policies the difference in approach makes
comparison hard, especially when using automated tools.  It isn't
obvious, at least to a machine, that disabling the camera will prevent
photographs from being taken.

To help resolve issues with specification we present a generalized schema of BYOD predicates that can be used when describing BYOD policies (\autoref{tab:predicates}).  They represent some of the most common decisions and entities making decisions we have found when looking at BYOD policies.
AppPAL policies can use arbitrary predicates, but to help structure our policies we use standard prefixes.

\begin{minipage}{\textwidth}
  \footnotesize
  \centering
\newcommand{\PrefixT}[3]{\textit{#1}~\texttt{#2}\textit{#3}}
\begin{tabular}{l l}
  \toprule
  Predicate & Meaning \\
  \midrule
  \PrefixT{Subject}{can}{Action} & The \emph{subject} is allowed to perform the \emph{action}. \\
  \PrefixT{Subject}{has}{Attribute} & The \emph{subject} has the a certain \emph{attribute}. \\
  \PrefixT{Subject}{is}{Type} & The \emph{subject} is an instance of \emph{type}. \\
  \PrefixT{Subject}{must}{Task} & The \emph{subject} should try to satisfy the obligation. \\
  \bottomrule\\
\end{tabular}
\end{minipage}

\begin{table}
  \footnotesize
  \newcommand{\Obj}[1]{\textit{#1}}
  \newcommand{\Sbj}[1]{\texttt{#1}}
  \newcommand{\Sba}[2]{\Sbj{#1(}\Obj{#2}\Sbj{)}}
\begin{tabular}{p{0.35\textwidth} p{0.65\textwidth}}
  \toprule
  Predicate & Description \\
  \midrule
  \Obj{App} \Sbj{isInstallable} & An \emph{app} can be installed. \\
  \Obj{X} \Sba{hasManager}{Manager} & Specifies who \emph{X's}~manager is. \\
  \Obj{Device} \Sba{mustDisable}{Feature} & The \emph{device} should ensure that the \emph{feature} is disabled. \\
  TODO & Fill this in with most common ones and add longer list to appendix. \\
  \bottomrule \\
\end{tabular}
\label{tab:predicates}
\caption{Excerpt of Generalized predicates for BYOD policies.}
\end{table}

Using AppPAL we were able to translate 23 out of the 60 requriements in the NHS policy and XX out of the XX requirements in the SANS policy. \textbf{\large TODO: check these numbers properly}.  Sometimes this was because the requirement was not a policy rule but rather something the device owner should be aware of.  For example:

\begin{quote}\footnotesize
  \textbf{NHS}: \itshape 
``In receiving a mobile device from the Trust, the individual receiving and using
the device accepts that the device can be used to communicate through all
corporate channels including voice calls, emails and texts and where appropriate
enabled web applications, during working hours.''
\end{quote}

\begin{quote}\footnotesize
  \textbf{NHS}: \itshape 
``Individuals who have personal data of any kind stored on a corporately issued
mobile device must be aware that in the event of loss of the device the above
data wipe will include removal of all personal data.''
\end{quote}

In these cases no rule is needed as there is no behaviour we wish to control. 
If we wished to use AppPAL to enforce these policies we could require device owners issue a statement acknowledging these limitations, similar to an EULA agreement.

{\large \textbf{...actually this sounds like a good idea.  What percentage of these policies are acknowledgements.}}

% 5 pages
\section{The Details}
\label{sec:details}

\newcommand{\facts}[1]{\ensuremath \textsf{facts}\left(#1\right)}
\newcommand{\reachable}[2]{\ensuremath \textsf{reachable}\left(#1,~#2\right)}
\subsection{Checking Reachability}

Inference in AppPAL happens by collecting ground facts and constraints that satisfy rules. 
These rules are then combined to form a policy.
If a rule is reachable then we would expect there to be some combination of facts that could satisfy the rule.
If there are no facts that could satisfy the rule then this hints that the policy may be incomplete as there are policy rules that can never be satisfied.

\begin{equation}
  \reachable{R}{AC}\equiv\exists p\in\facts{R}:~p\theta\subseteq AC
\end{equation}
Where $\facts{R}$ is the set of sets of facts and constraints that would satisfy rule $R$, and $\theta$ is an alpha-renaming of any variables in the set of facts such that equivalent free variable have the same name.

We find all reachable assertions using the algorithm shown in~\autoref{alg:reachability} building a set of pairs of speakers and predicates.  If a pair is in the set then that predicate is said to be reachable from that speaker (i.e.~we can infer the speaker may say something about that predicate).  
We search over all the assertions in the assertion context.  If all of an assetion's conditionals (the facts in the \emph{if} part) are reachable (or it has none) then the speaker and predicate is added to the reachable set.  If the statement is a can-say statement then we additionally check if the delegated predicate is reachable from the delgated speaker, and if so mark the delegated statement as reachable from the speaker who made the can-say statement.

\begin{figure}
\begin{algorithmic}
\Procedure {Reachable}{AC}
  \State reachable $\gets \emptyset$
  \Repeat
    \State iterate $\gets$ False
    \ForAll {$a \in$ AC}
      \State $e \gets $Speaker($a$)
      \State $p \gets $Predicate($a$)
      \If {$p~$:: CanSay $\wedge$ $(e,p)\in$ reachable}
        \If {$($Delegator($p$)$,$ Delegation($p$)$)~\in$ reachable}
          \If {$\forall c \in$ Conditions($a$)$.~(e,$ Predicate($c$)$) \in$ reachable}
            \State reachable $\gets$ reachable $\cup$ ($e,$ Delegation($p$))
            \State iterate $\gets$ True
          \EndIf
        \EndIf
      \ElsIf {$(e,p)\not\in$ reachable}
        \If {$\forall c \in$ Conditions($a$)$.~(e,$ Predicate($c$)$) \in$ reachable}
          \State reachable $\gets$ reachable $\cup~(e,p)$
          \State iterate $\gets$ True
        \EndIf
      \EndIf
    \EndFor
  \Until {iterate = True}
  \State \textbf{return} reachable
\EndProcedure
\end{algorithmic}
\caption{Procedure for finding all reachable assertions.}
\label{alg:reachability}
\end{figure}

\subsection{Checking Redundancy}

When checking for redundency we are looking for rules that will need the same or a subset of the ground facts and constraints to satisfy it.
If one rule makes another redundant then the first rule will not only always be satisfied whenever the redundant one is, but may require less facts in order to satisfy it.

Formally for any two rules $R$ and $R^\prime$; $R$ is said to be \emph{made redundant} by $R^\prime$ ($R\geq R^\prime$) if:
\begin{equation}
  R \geq R^\prime \equiv \forall p \in \facts{R}.~\exists p^\prime \in \facts{R^\prime}:~p\theta \supseteq p^\prime\theta
\end{equation}
If $R\geq R^\prime$ and $R\leq R^\prime$ then the two rules are equivalent (i.e. proving one will always prove the other).  This is particularly interesting if the two rules are for different predicates as it indicates that the two predicates may be equivalent also.

% 1-2 pages
\section{Related Work}
\label{sec:related}

The policies present in the NHS organisation and hospitals have been an attractive target for the application of formal methods.
There have been multiple attempts to present formalizations of electronic health records in both the Ponder~\cite{blobel_authorization_2004} and Cassandra~\cite{becker_cassandra_2004} policy languages.

% 0.5 pages
\section{Conclusions}
\label{sec:conclusions}

\bibliography{paper}{}
\bibliographystyle{plain}
\end{document}

### 2016-06-30 paper-2.tex
adds a second version with more of a focus on medical byod

\documentclass[a4paper]{article}

\usepackage{microtype}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{acronym}
\acrodef{MDM}{mobile device management}
\acrodef{BYOD}{bring your own device}

\usepackage{algorithm}
\usepackage{algpseudocode}

% \apppal{} and Listing formatting
\usepackage{listings}
\lstdefinelanguage{AppPAL}{%
  morekeywords={if,inf,says,where,true,false},
  otherkeywords={can-act-as,can-say},
  sensitive=true,
  morestring=[b]',
  literate={\ inf\ }{{$\infty$}}5
}[keywords,strings]
\lstset{%
  basicstyle=\ttfamily\footnotesize{},
  stringstyle=\sffamily\footnotesize{},
  keywordstyle=\ttfamily\slshape\footnotesize{},
  language=AppPAL,
  columns=flexible,
  frame=single,
  framesep=0.5pt,
  framexleftmargin=2.5pt,
  framexrightmargin=2.5pt,
}

\newcommand{\dotdotdot}[1]{#1}
\newcommand{\etal}[0]{et~al{.}\@}
\newcommand{\apppal}[0]{App{P}{A}{L}}

\title{Case Study Reasoning about BYOD Policies in the Healthcare Sector}
\author{Joseph Hallett and David Aspinall}
\begin{document}
\maketitle

% 4 sentences
\begin{abstract}
  Implementing a \ac{BYOD} policy can be difficult especially in hospitals, where there is a need to balance privacy requirements against the need to treat patients.
  It is common to publish a \ac{BYOD} policy to describe how staff should use their devices in the workplace.
  These policies are written using natural language, and employees are expected to read and follow them: this can be tricky as the policies are often poorly specified.
  We show how \apppal{}, a policy language for mobile app preferences, can be used as a tool for reasoning about policies and how it found issues with BYOD policies used in hospitals and as guidelines.
\end{abstract}

\section{Introduction}
% 1 page
\label{sec:intro}
% 1. Describe the problem
%    - BYOD hard
%    - Hospitals are an interesting case
%    - Need tools for reasoning about policies not enforcing

Employees bringing their own devices to work has become the norm:
  around 70\% of all companies having some form of \ac{BYOD} scheme~\cite{schulze_byod_2016}.
In hospitals the use of personal devices as part of patient care is particularly common.
One survey from the UK found that around 80\% of surgical doctors were willing to use their personal devices for work, with 85\% regularly using them to look up information, and 50\% having medical apps already installed on their phones~\cite{patel_uk_2015}.  Another from the US found similar numbers of physicians regularly using their devices as part of their work~\cite{moyer_managing_2013}.

Several healthcare organisations have started looking into using mobile devices as part of patient care, and found that these devices may be helpful~\cite{prgomet_impact_2009,faulds_feasibility_2016}.  
Organisations, such as the US's HiMMS~\cite{seymour_mobile_2012}, have published guidelines helping hospital administrators implement BYOD policies in their workplaces.  
Additionally general advice for other organisations seeking to implement BYOD policies is available from SANS~\cite{nicholas_r._c._guerin_security_2008}, NIST~\cite{souppaya_guidelines_????}, HP~\cite{hp_byod_????} and the CESG~\cite{cesg_byod_2015} amongst others.
The NHS has even published guidelines for developers seeking to develop apps for use in hospitals~\cite{prgomet_impact_2009}.

Whilst there has been a push for bring \ac{BYOD} and mobile devices into the healthcare industry there has been push back against having mandated security policies automatically enforced.  
The use of \ac{MDM} software to enforce policies can be problematic:
  a survey found that even for companies that use \ac{MDM}, over half of them still have devices in their networks that do not meet their policies~\cite{mobileiron_security_labs_q4_2015}.
In the healthcare industry the circumvention of security controls by physicians is normal, and in some cases taught as good practice~\cite{koppel_workarounds_2015}.
The use of tools to automatically enforce policies in a hospital environment is not always the right approach:
\emph{Moyer} concluded that:
\emph{``Since mobile devices are already being used in hospitals, an approach toward educating the users instead of trying to control the technology may be more practical in the short term''}~\cite{moyer_managing_2013}.
Similarly \emph{Faulds}~\etal~echo this sentiment noting that \emph{``concerns about data protection are legitimate''}, and going on to \emph{``encourage the Information Commissioner’s Office, in consultation with professional medical organisations, to publish guidance specific to the use of BYOD technology''}~\cite{faulds_feasibility_2016}.

The conflict between patient care and data security, as well as \ac{MDM} software being relatively limited, has lead to \ac{BYOD} policies being published as natural language documents that employees are supposed to agree to follow.
Using natural language in policies can create ambiguities, and lead to poorly specified rules.
If a policy is poorly specified and confusing then its subjects may chose to ignore it or misinterpret it's rules.
If we wish to improve our policies we need to be able to reason about them clearly.
We need to be able to identify problems before they happen.

% 2. State your contribution
%    - Standard lexicon of predicates and idioms
%    - Formalisation of several BYOD policies
%    - Identification of problems in said policies
%    - Methods for inferring when \apppal{} policies are problematic

\apppal{} is a policy language for describing mobile app preferences~\cite{hallett_apppal_2016}.
We show how \apppal{} can also be used to formalise \ac{BYOD} policies by translating three \ac{BYOD} policies into \apppal{}.
One is taken from an NHS hospital trust~\cite{kennington_mobiles_2014}. 
Another is from HiMMS published as an example policy to help hospitals design their own policies~\cite{healthcare_information_and_management_systems_society_mobile_2012}.
The third is a  general policy from SANS to help organisations to develop their own \ac{BYOD} schemes~\cite{nicholas_r._c._guerin_security_2008}.
Using these policies we suggest a standard lexicon of BYOD predicates and idioms for using when reasoning about \ac{BYOD} policies. 
With the AppPAL formalisation we show how problems with reachability and redundancy can be identified automatically, and suggest where the policies could be improved.

\section{My Idea}
% 2 pages
\label{sec:idea}

% - Brief intro to apppal but mostly link to existing paper
% - Give an overview of the different policies we're looking at
% - Show differences and similarities between the policies
% - Make sure to tie it all into healthcare, but also be clear this is more general

\section{The Details}
% 5 pages
\label{sec:details}

% - Show the standard predicates and explain how they help to give structure to apppal policies
%   - show how they are tied into the trust relationships with patient care (i.e. consent)
% - Show the consistency checker
%   - Give algorithm and intuition
%   - Show examples where it has found problems
% - Show redundancy checks
%   - Algorithm and intuition
%   - Examples where problems were found

\section{Related Work}
% 1-2 pages
\label{sec:related}

\section{Conclusions}
% 0.5 pages
\label{sec:conclusions}

\bibliography{paper}{}
\bibliographystyle{plain}
\end{document}

### 2016-06-30 paper-2.tex
Added intro to the three different policies

@@
 \acrodef{BYOD}{bring your own device}
+\acrodef{HiMMS}{Healthcare Information and Management Systems Society}
 
@@
 Several healthcare organisations have started looking into using mobile devices as part of patient care, and found that these devices may be helpful~\cite{prgomet_impact_2009,faulds_feasibility_2016}.
-Organisations, such as the US's HiMMS~\cite{seymour_mobile_2012}, have published guidelines helping hospital administrators implement BYOD policies in their workplaces.  
+Organisations, such as the US's \ac{HiMMS}~\cite{seymour_mobile_2012}, have published guidelines helping hospital administrators implement BYOD policies in their workplaces.  
 Additionally general advice for other organisations seeking to implement BYOD policies is available from SANS~\cite{nicholas_r._c._guerin_security_2008}, NIST~\cite{souppaya_guidelines_????}, HP~\cite{hp_byod_????} and the CESG~\cite{cesg_byod_2015} amongst others.
@@
 One is taken from an NHS hospital trust~\cite{kennington_mobiles_2014}. 
-Another is from HiMMS published as an example policy to help hospitals design their own policies~\cite{healthcare_information_and_management_systems_society_mobile_2012}.
+Another is from \ac{HiMMS} published as an example policy to help hospitals design their own policies~\cite{healthcare_information_and_management_systems_society_mobile_2012}.
 The third is a  general policy from SANS to help organisations to develop their own \ac{BYOD} schemes~\cite{nicholas_r._c._guerin_security_2008}.
@@
 
-\section{My Idea}
+%\section{My Idea}
+\section{Formalising policies using AppPAL}
 % 2 pages
@@
 
+Each of the three policies we have chosen to look at has a different style in expressing their rules.
+Despite the different styles, however, there are areas where the policies overlap and make similar rules.
+We present the policies here as excerpts alongside their AppPAL translations.
+The full translations and original policies can be found online\footnote{\url{https://github.com/apppal/apppal-byod-policy-translations}}.
+
+The \ac{HiMMS} policy is taken from their \emph{Mobile Security Toolkit} for assisting healthcare organisations and security practitioners in managing the security of their mobile computing devices. 
+\ac{HiMMS} itself is an American non-profit company advising healthcare companies on IT.
+Their policy~\cite{healthcare_information_and_management_systems_society_mobile_2012} takes the form of a short user agreement that it's subjects are expected to sign.  
+This gives a different style to the other policies: policies are expressed from the perspective of the subject, rather than as decrees from the policy authors.
+It also congtains many rules where the subject is expected to acknowledge a risk or regulation.  
+For example in one rule the subject is required to acknowledge that networks can be unreliable:
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMMS}}:~\itshape 
+I understand and accept that synchronisation relies on one or more cellular network providers and the Internet, and that both are subject to slowdowns and outages of extended duration that are beyond the control of IT.}]
+'himms' says User can-say User hasAcknowledged('internet-unreliable')
+  if User isUser.
+
+'himms' says User hasAgreeedToPolicy
+  if ...,  User hasAcknowledged('internet-unreliable').
+\end{lstlisting}
+These rules are interesting as on the face of them they do not say anything about security or how the subject should use their device.
+We translate rules like this into acknowledgement statements in AppPAL, which can be collected into a check that the user has acknowledged everything they should have done.
+
+The NHS policy~\cite{kennington_mobiles_2014}, in contrast, is a more prescriptive document.
+It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series fo \emph{trusts}, each of which contains several hospitals.
+A doctor may work at several hospitals within a trust.}.
+It is interesting in part because it represents a \ac{BYOD} policy in actual use, but also because it is significantly more complex than most \ac{BYOD} policies as it goes beyond basic requirements such as requiring encryption by describing relationships between organisational groups within the trust and precise organisation specific requirements.
+Like the \ac{HiMMS} policy it has rules that just require an acknowledgement (for example Rule~3.2 requires that the device owner \emph{``accepts that the device can be used to communicate through all corporate channels''}) but there is a greater emphasis on the specifying the responsibilities of different groups and roles within the hospital.
+For example in Rule~2.4 a staff member who has applied for a phone needs it to be signed off by a finance director and a relevant assistant director:
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{NHS}}:~\itshape 
+Upon receipt of the completed application form, allocation will also need to be authorised by the relevant Assistant Director and the Director of Finance.}]
+'nhs-trust' says Staff hasAppliedForPhone(Form)
+  if Staff hasSubmitted(Form),
+     'assistant-director' hasAuthorized(Form),
+     'finance-director' hasAuthorized(Form).
+
+'nhs-trust' says X can-say 'assistant-director' hasAuthorized(Form)
+  if X can-act-as 'assistant-director',
+     Staff hasSubmitted(Form),
+     X isManagerOf(Staff).
+\end{lstlisting}
+
+The SANS Institute is a private US company that specialise in information security.
+Whilst their example \ac{BYOD} policy is not healthcare specific, it is still interesting as there advice is widely regarded, and contains more general security advice.
+The SANS \ac{BYOD} policy is split into sections that describe how device features, such as Bluetooth or OTA provisioning, should be configured.
+Some rules are acknowledgements: for example the use of internet services in Section~2.8 requires acknowledgement of the email security, acceptable use and an eCommerce security policy.
+Many rules describe when device features should be disabled, in contrast to the other two policies which tend to favor describing when features can be used.
+For example in Section~2.4 the SANS policy states that infrared support should be disabled if Blutooth is available: this is easilly expressed in AppPAL:
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
+Infrared supported MUST be disabled if Bluetooth connectivity is supported.}]
+'sans' says Device shouldDisable('irda')
+  if Device hasFeature('bluetooth').
+\end{lstlisting}
+
 % - Brief intro to apppal but mostly link to existing paper
+%   - Why apppal?
 % - Give an overview of the different policies we're looking at

### 2016-06-30 paper-2.tex
justifies why apppal

@@
 This gives a different style to the other policies: policies are expressed from the perspective of the subject, rather than as decrees from the policy authors.
-It also congtains many rules where the subject is expected to acknowledge a risk or regulation.  
+It also contains many rules where the subject is expected to acknowledge a risk or regulation.  
 For example in one rule the subject is required to acknowledge that networks can be unreliable:
@@
 The NHS policy~\cite{kennington_mobiles_2014}, in contrast, is a more prescriptive document.
-It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series fo \emph{trusts}, each of which contains several hospitals.
+It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series of \emph{trusts}, each of which contains several hospitals.
 A doctor may work at several hospitals within a trust.}.
@@
 Many rules describe when device features should be disabled, in contrast to the other two policies which tend to favor describing when features can be used.
-For example in Section~2.4 the SANS policy states that infrared support should be disabled if Blutooth is available: this is easilly expressed in AppPAL:
+For example in Section~2.4 the SANS policy states that infrared support should be disabled if Bluetooth is available: this is easily expressed in AppPAL:
 \begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
@@
 % - Make sure to tie it all into healthcare, but also be clear this is more general
+
+\subsection{Why AppPAL?}
+
+AppPAL is an authorization language designed to model app preferences~\cite{hallett_apppal_2016}.
+It was created by instantiating the SecPAL language (which was designed for access control in distributed computing environments~\cite{becker_secpal:_2010}) with predicates and checking functions for describing apps.
+It supports delegation, roles and constraints including arbitrary external checking programs to describe precisely what apps users might want to use.
+
+AppPAL (and SecPAL) is designed to be readable.
+It is powerful enough to describe complex policies and rules, whilst also being semantically simple having only three rules for evaluation.
+When reasoning about \ac{BYOD} policies, it simplicity and readability gives it an advantage over more powerful languages such as XACML~\cite{oasis_extensible_2013} as a rule written in AppPAL can be comprehended quickly.
+Whilst an end-user might not write a policy in a formal language, an IT administrator might wish to use a formal language to aid with reasoning and checking the policies they write.
+AppPAL gives these IT professionals an easy entry point to formal methods, without over complicating things.
+Early experiments showed that AppPAL can describe \ac{BYOD} policies, and help make clarify mistakes ahead of time~\cite{hallett_specifying_2016}.
+
+An AppPAL evaluation library has been developed that runs on Java and Android and is available online\footnote{\url{https://github.com/apppal/libapppal}}.
+As part of this work we have developed (in Section~\ref{sec:details}) a series of policy analysis tools for AppPAL.
+These are compiled alongside the main AppPAL interpreter and may be run from the command-line.
 

### 2016-07-01 paper-2.tex
adds description of reachability and redundancy algorithms

@@
 
+\subsection{Policy Idioms}
+
+\subsection{Checking Reachability}
+
+Inference in AppPAL happens by collecting ground facts and constraints that satisfy rules.
+These rules are then combined to form a policy. 
+If a rule is reachable then we would expect there to be some combination of facts that could satisfy the rule.
+If there are no facts that could satisfy the rule then this suggests that the policy may be incomplete as there are policy rules that can never be satisfied.
+
+We search for reachable assertions using the algorithm in Listing~\ref{alg:reachable}:
+it produces a set of pairs of predicates and speakers where the a pair of a speaker and a predicate indicates that that speaker may say something about that predicate.
+We search over all the assertions in the AppPAL assertion context.
+If all of an assertion's conditionals (the facts in the if part) are reachable (or it has none) then the speaker and predicate are added to the reachable set.
+If the statement is a can-say statement then we additionally check if the delegated predicate is reachable from the delegated speaker, and if so mark the delegated statement as reachable from the speaker who made the can-say statement.
+
+\begin{lstlisting}[language=Python,float,caption={Procedure for finding all reachable assertions.},label={alg:reachable}]
+def reachable(ac) -> set:
+  reachable = new set()
+  iterate = True
+  while iterate == True:
+    iterate = False
+    for assertion in ac:
+      e = a.speaker
+      p = a.predicate
+      if p.isCanSay() and (e, p) in reachable:
+        if (p.delegator, p.delegation) in reachable:
+          if for all c in a.conditions: (e, c.predicate) in reachable:
+            reachable.add((e, p.delegation))
+            iterate = True
+      else if not (e, p) in reachable:
+        if for all c in a.conditions; (e, c.predicate) in reachable:
+          reachable.add((e, p))
+          iterate = True
+  return reachable
+\end{lstlisting}
+
+\subsection{Checking Redundancy}
+
+To check redundancy we convert the policy into a graph structure.
+Each statement we might want to prove (a goal) becomes a node in the graph, its children are organised into sets of goals (a proof), where if all the goals in any set were proved true then the goal node would also become true.
+If a node has an empty set of goals to prove it is a fact.
+Once the graph has been constructed (taking into account delegation and unification with other statements), the graph is flattened by applying the flatten procedure (Listing~\ref{alg:flatten}) repeatedly until a fixed point is reached.
+When the graph has been flattened we look for redundancy by looking for goals with proofs that are subsets of their other proofs, implying that if the larger proof is true, then the shorter proof will always be true too;
+  and different goals with identical proofs, implying that the two decisions are not independent~(Listing~\ref{alg:redundancy}).
+
+\begin{lstlisting}[language=Python,float,caption={Procedure for flattening the redundancy graph.},label={alg:flatten}]
+def flatten(graph):
+  for goal in graph.goals:
+    hoist = true
+    for proof in graph[goal]:
+      for proof_goal in proof:
+        if not proof_goal.is_fact:
+          hoist = False
+          break
+    if hoist == True:
+      hoist(graph, goal)
+
+def hoist(graph, target):
+  for parent in graph[target].parents:
+    for proof in parent.proofs:
+      if proof.uses(target):
+        for replacements in graph[target].parents:
+          parent.proofs += proof.replace(target, replacements)
+        parent.proofs -= proof
+\end{lstlisting}
+
+\begin{lstlisting}[language=Python,float,caption={Procedure to check for redundancy.},label={alg:redundancy}]
+def check_redundancy(graph) -> boolean:
+  for goal in graph.goals:
+    for a in graph[goal]:
+      for b in graph[goal]:
+        if a >= b and a.goals.subset(b.goals):
+          if b.goals.subset(a.goals):
+            warn(a+" has multiple equivalent proofs") 
+          else
+            warn(a+" has a redundant proof")
+    for other_goal in graph.goals:
+      if goal > other_goal:
+        for a in graph[goal]:
+          for b in graph[other_goal]:
+            if not (a.goals.is_empty() or b.is_empty())
+               and a.goals == b.goals:
+              warn(a+" and "+b+" have equivalent proofs")
+\end{lstlisting}
+
 \section{Related Work}

### 2016-07-06 paper-2.tex
Added a bit on prefixes

@@
 For example in Rule~2.4 a staff member who has applied for a phone needs it to be signed off by a finance director and a relevant assistant director:
-\begin{lstlisting}[title={\footnotesize\textbf{\ac{NHS}}:~\itshape 
+\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape 
 Upon receipt of the completed application form, allocation will also need to be authorised by the relevant Assistant Director and the Director of Finance.}]
@@
 For example in Section~2.4 the SANS policy states that infrared support should be disabled if Bluetooth is available: this is easily expressed in AppPAL:
-\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
+\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
 Infrared supported MUST be disabled if Bluetooth connectivity is supported.}]
@@
 \subsection{Policy Idioms}
+
+When translating the \ac{BYOD} policies we found types of actions or checks being performed.
+There are checks if things are of a particular type (for~example in the NHS policy rule~2.1  a staff member is elligable for a mobile device if they are lone-working, work out-of-hours and work out-of-office).
+There are checks if an action is allowed (such~as whether they can connect to an wireless access point in the \ac{HiMMS} policy);  and checks to see if an action has already been performed (like the need to have acknowledged a policy in all three policies).
+We also find policy rules that state the user should perform an action (such as disabling phone functionality in the SANS policy, or an obligation to report a phone missing in the NHS policy).
+
+To help structure the formalisation we give the AppPAL predicates used to translate the policies prefixes.
+The prefixes help to clarify the meaning of each predicate and groups the predicates into four categories.
+
+\begin{tabular}{l l}
+  \toprule
+  Predicate & Meaning \\
+  \midrule
+  \textit{Subject} \texttt{can}\textit{Action}  & The \emph{subject} is allowed to perform the \emph{action}. \\
+  \textit{Subject} \texttt{has}\textit{Action}  & The \emph{subject} has performed the \emph{action}. \\
+  \textit{Subject} \texttt{is}\textit{Propery}  & The \emph{property} is true about the \emph{subject}. \\
+  \textit{Subject} \texttt{must}\textit{Action} & The \emph{subject} is obliged to perform the action. \\
+  \bottomrule \\
+\end{tabular}
 
@@
 
+Armando~\etal~created a BYODroid~\cite{armando_bring_2013}; \ac{BYOD} framework to enforce \ac{BYOD} policies by extending Featherweight Java~\cite{igarashi_featherweight_2001} with a type system that can describe how Android apps interact.
+Whilst this work gives a framework for enforcing policies on Android, it doesn't help policy authors decide what the policies should be to begin with.
+
 \section{Conclusions}

### 2016-07-13 paper.tex
adds notes on the acknowledgement idiom

@@
 \acrodef{BYOD}{bring your own device}
+\acrodef{HiMSS}{Healthcare Information and Management Systems Society}
 
@@
 
-% AppPAL and Listing formatting
+% \apppal{} and Listing formatting
 \usepackage{listings}
@@
 \newcommand{\dotdotdot}[1]{#1}
+\newcommand{\etal}[0]{et~al{.}\@}
+\newcommand{\apppal}[0]{App{P}{A}{L}}
 
-\title{Case Study Formalising Mobile BYOD Policies with AppPAL}
+\title{Case Study Reasoning about BYOD Policies in the Healthcare Sector}
 \author{Joseph Hallett and David Aspinall}
@@
 \begin{abstract}
-  \begin{itemize}
-  \item BYOD policies written infromally, with no real means to enforce.
-  \item Created standard language by analyzing \emph{X} real policies, using AppPAL
-  \item Using formal methods we found \emph{Y} problems.
-  \item AppPAL is a helpful language for thinking about policies in. 
-  \end{itemize}
+  Implementing a \ac{BYOD} policy can be difficult especially in hospitals, where there is a need to balance privacy requirements against the need to treat patients.
+  It is common to publish a \ac{BYOD} policy to describe how staff should use their devices in the workplace.
+  These policies are written using natural language, and employees are expected to read and follow them: this can be tricky as the policies are often poorly specified.
+  We show how \apppal{}, a policy language for mobile app preferences, can be used as a tool for reasoning about policies and how it found issues with BYOD policies used in hospitals and as guidelines.
 \end{abstract}
 
+\section{Introduction}
 % 1 page
-\section{Introduction}
 \label{sec:intro}
 % 1. Describe the problem
-% 2. State your contributions
-% ...and that is all
+%    - BYOD hard
+%    - Hospitals are an interesting case
+%    - Need tools for reasoning about policies not enforcing
 
-Employees bring their personal devices to work.
-This creates a problem for corporate IT departments:
-  they want to secure their company's networks, but they have limited control over the personal devices running on them.
-To ensure compliance IT departments pubilsh mobile device polices.
-A recent survey found that around 70\% of companies published some form of mobile device policy~\cite{schulze_byod_2016}.
-Employees must agree to follow these policies if they want to use their personal devices at work.
+Employees bringing their own devices to work has become the norm:
+  around 70\% of all companies having some form of \ac{BYOD} scheme~\cite{schulze_byod_2016}.
+In hospitals the use of personal devices as part of patient care is particularly common.
+One survey from the UK found that around 80\% of surgical doctors were willing to use their personal devices for work, with 85\% regularly using them to look up information, and 50\% having medical apps already installed on their phones~\cite{patel_uk_2015}.
+Another from the US found similar numbers of physicians regularly using their devices as part of their work~\cite{moyer_managing_2013}.
 
-Often the policies are specified with natural language.
-This makes compliance hard as the policies can contain ambiguities, contradictions and can change without the user's noticing there is a new policy to follow.
-To solve this some companies offer \ac{MDM} software that can help enforce some of the policies, however their capabilities are usually limited to banning specific apps and configuring email and filesharing services.
-Even when companies have \ac{MDM} software a survey found that over 50\% of companies have one or more devices on their network that did not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.
+Several healthcare organisations have started looking into using mobile devices as part of patient care, and found that these devices may be helpful~\cite{prgomet_impact_2009,faulds_feasibility_2016}.
+Organisations, such as the US's \ac{HiMSS}~\cite{seymour_mobile_2012}, have published guidelines helping hospital administrators implement BYOD policies in their workplaces.  
+Additionally general advice for other organisations seeking to implement BYOD policies is available from SANS~\cite{nicholas_r._c._guerin_security_2008}, NIST~\cite{souppaya_guidelines_????}, HP~\cite{hp_byod_????} and the CESG~\cite{cesg_byod_2015} amongst others.
+The NHS has even published guidelines for developers seeking to develop apps for use in hospitals~\cite{prgomet_impact_2009}.
 
-Formal languages can help to remove ambiguities in policies and identify when policies contain problems.
-We show how the AppPAL language~\cite{hallett_apppal_2016} can be used to formalise \ac{BYOD} policies, and present a standard set of predicates and idioms from translating real \ac{BYOD} policies from the SANS institute and NHS.
-By translating these policies into formal language we were able to identify cases where the policy rule were duplicated, and underspecified.
+Whilst there has been a push for bring \ac{BYOD} and mobile devices into the healthcare industry there has been push back against having mandated security policies automatically enforced.  
+The use of \ac{MDM} software to enforce policies can be problematic:
+  a survey found that even for companies that use \ac{MDM}, over half of them still have devices in their networks that do not meet their policies~\cite{mobileiron_security_labs_q4_2015}.
+In the healthcare industry the circumvention of security controls by physicians is normal, and in some cases taught as good practice~\cite{koppel_workarounds_2015}.
+The use of tools to automatically enforce policies in a hospital environment is not always the right approach:
+\emph{Moyer} concluded that:
+\emph{``Since mobile devices are already being used in hospitals, an approach toward educating the users instead of trying to control the technology may be more practical in the short term''}~\cite{moyer_managing_2013}.
+Similarly \emph{Faulds}~\etal~echo this sentiment noting that \emph{``concerns about data protection are legitimate''}, and going on to \emph{``encourage the Information Commissioner’s Office, in consultation with professional medical organisations, to publish guidance specific to the use of BYOD technology''}~\cite{faulds_feasibility_2016}.
 
-In this paper we present excerpts from the AppPAL translations of two BYOD policies.
-In particular we focus on the translations of a policy from an NHS Trust hospital, and from an exemplar policy from the SANS institute designed to help companies devise their own policies.
-These policies are more complex than a typical BYOD policies as they have rules which go beyond white and black listing.
-We illustrate how AppPAL's checking methods can identify when the policies are under-specified and when the policies contain redundency, either through policies being duplicated in the documents or by policies with simplified rules in the same document.
+The conflict between patient care and data security, as well as \ac{MDM} software being relatively limited, has lead to \ac{BYOD} policies being published as natural language documents that employees are supposed to agree to follow.
+Using natural language in policies can create ambiguities, and lead to poorly specified rules.
+If a policy is poorly specified and confusing then its subjects may chose to ignore it or misinterpret it's rules.
+If we wish to improve our policies we need to be able to reason about them clearly.
+We need to be able to identify problems before they happen.
 
-% 1 page
-\section{The Problem}
-\label{sec:problem}
+% 2. State your contribution
+%    - Standard lexicon of predicates and idioms
+%    - Formalisation of several BYOD policies
+%    - Identification of problems in said policies
+%    - Methods for inferring when \apppal{} policies are problematic
 
-Consider a policy for describing what apps can be installed on a mobile device for use at work.
-One approach might be to provide a list of authorized and unauthorized apps.
-Employees should cosult this list before running any apps, and any \ac{MDM} software ought to enforce the list automatically.
-Both the SANS and NHS policies give requirements implementing these policies.
-In the SANS policy the IT department is responsible for maintaining the lists, and in the NHS policy (which only uses whitelisting) an app is only approved if the Care and Clinical Policy group approves it as well as another group depending on whether it is a clinical or non-clinical app.
+\apppal{} is a policy language for describing mobile app preferences~\cite{hallett_apppal_2016}.
+We show how \apppal{} can also be used to formalise \ac{BYOD} policies by translating three \ac{BYOD} policies into \apppal{}.
+One is taken from an NHS hospital trust~\cite{kennington_mobiles_2014}. 
+Another is from \ac{HiMSS} published as an example policy to help hospitals design their own policies~\cite{healthcare_information_and_management_systems_society_mobile_2012}.
+The third is a  general policy from SANS to help organisations to develop their own \ac{BYOD} schemes~\cite{nicholas_r._c._guerin_security_2008}.
+Using these policies we suggest a standard lexicon of BYOD predicates and idioms for using when reasoning about \ac{BYOD} policies. 
+With the AppPAL formalisation we show how problems with reachability and redundancy can be identified automatically, and suggest where the policies could be improved.
 
-\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
-``The IT Department maintains a list of allowed and unauthorised applications and makes them available to users on the intranet.''}]
-'sans-policy' says 'it-department' can-say App isInstallable.
+%\section{My Idea}
+\section{Formalising policies using AppPAL}
+% 2 pages
+\label{sec:idea}
+
+Each of the three policies we have chosen to look at has a different style in expressing their rules.
+Despite the different styles, however, there are areas where the policies overlap and make similar rules.
+We present the policies here as excerpts alongside their AppPAL translations.
+The full translations and original policies can be found online\footnote{\url{https://github.com/apppal/apppal-byod-policy-translations}}.
+
+The \ac{HiMSS} policy is taken from their \emph{Mobile Security Toolkit} for assisting healthcare organisations and security practitioners in managing the security of their mobile computing devices. 
+\ac{HiMSS} itself is an American non-profit company advising healthcare companies on IT.
+Their policy~\cite{healthcare_information_and_management_systems_society_mobile_2012} takes the form of a short user agreement that it's subjects are expected to sign.  
+This gives a different style to the other policies: policies are expressed from the perspective of the subject, rather than as decrees from the policy authors.
+It also contains many rules where the subject is expected to acknowledge a risk or regulation.  
+For example in one rule the subject is required to acknowledge that networks can be unreliable:
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMSS}}:~\itshape 
+I understand and accept that synchronisation relies on one or more cellular network providers and the Internet, and that both are subject to slowdowns and outages of extended duration that are beyond the control of IT.}]
+'ac{HiMSS}' says User can-say User hasAcknowledged('internet-unreliable')
+  if User isUser.
+
+'ac{HiMSS}' says User hasAgreeedToPolicy
+  if ...,  User hasAcknowledged('internet-unreliable').
 \end{lstlisting}
-\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape ``%
-Apps for work usage must not be downloaded onto corporately issued mobile devices (even if approved on the NHS apps store) unless they have been approved through the following Trust channels: 
-    Clinical apps;
-        at the time of writing there are no apps clinically approved by the Trust for use with patients/clients. 
-        However, if a member of staff believes that there are clinical apps \dotdotdot{or other technologies that could benefit their patients/clients, this should be discussed with the clinical lead in the first instance and} ratification should be sought via the Care and Clinical Policies Group.
-        \dotdotdot{A clinical app should not be used if it has not been approved via this group.}
-    Business apps; 
-        at the time of writing there are no business (i.e., non-clinical) apps approved by the Trust for use other than those preloaded onto the device at the point of issue. 
-        However, if a member of staff believes that there are apps \dotdotdot{or other technologies that could benefit their non-clinical work,} ratification of the app must be sought via the Management of Information Group (MIG).
-        \dotdotdot{An app should not be used if it has not been approved via this group.}
-Following approval through Care and Clinical Policies and/or MIG, final approval will be required through Integrated Governance Committee.''}]
-'nhs-policy' says App isInstallable if App isUsable, App isFinallyApproved.
-'nhs-policy' says 'cacpg' can-say App isUsable if App isUsableClinically.
-'nhs-policy' says 'mig' can-say App isUsable if App isUsableNonClinically.
-'nhs-policy' says 'igc' can-say App isFinallyApproved.
+These rules are interesting as on the face of them they do not say anything about security or how the subject should use their device.
+We translate rules like this into acknowledgement statements in AppPAL, which can be collected into a check that the user has acknowledged everything they should have done.
+
+The NHS policy~\cite{kennington_mobiles_2014}, in contrast, is a more prescriptive document.
+It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series of \emph{trusts}, each of which contains several hospitals.
+A doctor may work at several hospitals within a trust.}.
+It is interesting in part because it represents a \ac{BYOD} policy in actual use, but also because it is significantly more complex than most \ac{BYOD} policies as it goes beyond basic requirements such as requiring encryption by describing relationships between organisational groups within the trust and precise organisation specific requirements.
+Like the \ac{HiMSS} policy it has rules that just require an acknowledgement (for example Rule~3.2 requires that the device owner \emph{``accepts that the device can be used to communicate through all corporate channels''}) but there is a greater emphasis on the specifying the responsibilities of different groups and roles within the hospital.
+For example in Rule~2.4 a staff member who has applied for a phone needs it to be signed off by a finance director and a relevant assistant director:
+\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape 
+Upon receipt of the completed application form, allocation will also need to be authorised by the relevant Assistant Director and the Director of Finance.}]
+'nhs-trust' says Staff hasAppliedForPhone(Form)
+  if Staff hasSubmitted(Form),
+     'assistant-director' hasAuthorized(Form),
+     'finance-director' hasAuthorized(Form).
+
+'nhs-trust' says X can-say 'assistant-director' hasAuthorized(Form)
+  if X can-act-as 'assistant-director',
+     Staff hasSubmitted(Form),
+     X isManagerOf(Staff).
 \end{lstlisting}
 
-In the case of the NHS requirement the entire requirement seems redundant as it states there no apps have been approved.
-What the rule gives us, and what is made explicit in the formalization is who would need to approve the app if we were going to get an approved one.
-As well as providing improvements to readability, the formalization in AppPAL makes explicit who has approved each app.
-It also helps us spot when a rule is duplicated, for instance the SANS policy later has a requirement that:
-
+The SANS Institute is a private US company that specialise in information security.
+Whilst their example \ac{BYOD} policy is not healthcare specific, it is still interesting as there advice is widely regarded, and contains more general security advice.
+The SANS \ac{BYOD} policy is split into sections that describe how device features, such as Bluetooth or OTA provisioning, should be configured.
+Some rules are acknowledgements: for example the use of internet services in Section~2.8 requires acknowledgement of the email security, acceptable use and an eCommerce security policy.
+Many rules describe when device features should be disabled, in contrast to the other two policies which tend to favor describing when features can be used.
+For example in Section~2.4 the SANS policy states that infrared support should be disabled if Bluetooth is available: this is easily expressed in AppPAL:
 \begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
-``Only approved third party applications can be
-installed on handhelds. The approved list can be
-obtained by contacting the IT department, or should be
-available on the intranet.''}]
-'sans-policy' says 'it-department' can-say App isInstallable.
+Infrared supported MUST be disabled if Bluetooth connectivity is supported.}]
+'sans' says Device shouldDisable('irda')
+  if Device hasFeature('bluetooth').
 \end{lstlisting}
 
-When reading the natural language requirement it might not be immediately obvious that there is duplication, but with the formalization it is immediately apparent.
-This suggests that the policy could be improved by removing, or combining the requirements to remove ambiguity and prevent problems later if the requirement was to be updated.
+% - Brief intro to apppal but mostly link to existing paper
+%   - Why apppal?
+% - Give an overview of the different policies we're looking at
+% - Show differences and similarities between the policies
+% - Make sure to tie it all into healthcare, but also be clear this is more general
 
-% 2 pages
-\section{My Idea}
-\label{sec:idea}
+\subsection{Why AppPAL?}
 
-Writing policies in a formal language gives us two major benefits. 
-Firstly it gives us a precise framework for describing what each policy actually means and how it can be satisfied and allowing comparison between policies from different sources.
-Secondly it allows the use of policy checking tools that can infer when a policy may have problems.
+AppPAL is an authorization language designed to model app preferences~\cite{hallett_apppal_2016}.
+It was created by instantiating the SecPAL language (which was designed for access control in distributed computing environments~\cite{becker_secpal:_2010}) with predicates and checking functions for describing apps.
+It supports delegation, roles and constraints including arbitrary external checking programs to describe precisely what apps users might want to use.
 
-Translating natural language policies into fromal language is not trivial, however.
-Care should be taken to preserve not only the meaning of the policy but also the spirit and style of the original policy.
-Not all policies are specified in the same manner however. 
-Both the SANS and NHS policies, for example, specify when it is appropriate to use the camera.
-The SANS policy tends toward being prohibitive, describing when features should be disabled.
-\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape
-``Digital camera embedded on handheld devices might be
-disabled in restricted environments, according to Company risk analysis.}]
-'company' says Device mustDisable('camera') 
-  if 'restricted-environment' isWhereIs(Device).
-\end{lstlisting}
-The NHS policy is expressed differently describing when it is appropriate to use the camera.
-\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape
-  ``Some mobile devices have the ability to take photographs~/~videos. This
-  function should not be used for photographs~/~videos of an individual's
-  care and treatment unless the device has encryption enabled and it is
-  clinically appropriate to do so.''}]
-'nhs-trust' says Device canPhotograph(Patient)
-  if Device isEncrypted, Patient canBePhotographed.
-\end{lstlisting}
+AppPAL (and SecPAL) is designed to be readable.
+It is powerful enough to describe complex policies and rules, whilst also being semantically simple having only three rules for evaluation.
+When reasoning about \ac{BYOD} policies, it simplicity and readability gives it an advantage over more powerful languages such as XACML~\cite{oasis_extensible_2013} as a rule written in AppPAL can be comprehended quickly.
+Whilst an end-user might not write a policy in a formal language, an IT administrator might wish to use a formal language to aid with reasoning and checking the policies they write.
+AppPAL gives these IT professionals an easy entry point to formal methods, without over complicating things.
+Early experiments showed that AppPAL can describe \ac{BYOD} policies, and help make clarify mistakes ahead of time~\cite{hallett_specifying_2016}.
 
-Both approaches can be expressed using AppPAL, and even if the two
-approaches were mixed into the same policy in this case it would not
-cause problems from an enforcement perspective: a photo could not be
-taken if the camera were disabled even if the the user had the
-authorization.  In general though this can lead to confilicts, which
-will need to be resolved.  When using AppPAL as a tool for comparison
-and thinking about policies the difference in approach makes
-comparison hard, especially when using automated tools.  It isn't
-obvious, at least to a machine, that disabling the camera will prevent
-photographs from being taken.
+An AppPAL evaluation library has been developed that runs on Java and Android and is available online\footnote{\url{https://github.com/apppal/libapppal}}.
+As part of this work we have developed (in Section~\ref{sec:details}) a series of policy analysis tools for AppPAL.
+These are compiled alongside the main AppPAL interpreter and may be run from the command-line.
 
-To help resolve issues with specification we present a generalized schema of BYOD predicates that can be used when describing BYOD policies (\autoref{tab:predicates}).
-They represent some of the most common decisions and entities making decisions we have found when looking at BYOD policies.
-AppPAL policies can use arbitrary predicates, but to help structure our policies we use standard prefixes.
+\section{The Details}
+% 5 pages
+\label{sec:details}
 
-\begin{minipage}{\textwidth}
-  \footnotesize
-  \centering
-\newcommand{\PrefixT}[3]{\textit{#1}~\texttt{#2}\textit{#3}}
+% - Show the standard predicates and explain how they help to give structure to apppal policies
+%   - show how they are tied into the trust relationships with patient care (i.e. consent)
+% - Show the consistency checker
+%   - Give algorithm and intuition
+%   - Show examples where it has found problems
+% - Show redundancy checks
+%   - Algorithm and intuition
+%   - Examples where problems were found
+
+\subsection{Policy Idioms}
+
+When translating the \ac{BYOD} policies we found types of actions or checks being performed.
+There are checks if things are of a particular type (for~example in the NHS policy rule~2.1  a staff member is elligable for a mobile device if they are lone-working, work out-of-hours and work out-of-office).
+There are checks if an action is allowed (such~as whether they can connect to an wireless access point in the \ac{HiMSS} policy);  and checks to see if an action has already been performed (like the need to have acknowledged a policy in all three policies).
+We also find policy rules that state the user should perform an action (such as disabling phone functionality in the SANS policy, or an obligation to report a phone missing in the NHS policy).
+
+To help structure the formalisation we give the AppPAL predicates used to translate the policies prefixes.
+The prefixes help to clarify the meaning of each predicate and groups the predicates into four categories.
+
 \begin{tabular}{l l}
@@
   \midrule
-  \PrefixT{Subject}{can}{Action} & The \emph{subject} is allowed to perform the \emph{action}. \\
-  \PrefixT{Subject}{has}{Attribute} & The \emph{subject} has the a certain \emph{attribute}. \\
-  \PrefixT{Subject}{is}{Type} & The \emph{subject} is an instance of \emph{type}. \\
-  \PrefixT{Subject}{must}{Task} & The \emph{subject} should try to satisfy the obligation. \\
-  \bottomrule\\
-\end{tabular}
-\end{minipage}
-
-\begin{table}
-  \footnotesize
-  \newcommand{\Obj}[1]{\textit{#1}}
-  \newcommand{\Sbj}[1]{\texttt{#1}}
-  \newcommand{\Sba}[2]{\Sbj{#1(}\Obj{#2}\Sbj{)}}
-\begin{tabular}{p{0.35\textwidth} p{0.65\textwidth}}
-  \toprule
-  Predicate & Description \\
-  \midrule
-  \Obj{App} \Sbj{isInstallable} & An \emph{app} can be installed. \\
-  \Obj{X} \Sba{hasManager}{Manager} & Specifies who \emph{X's}~manager is. \\
-  \Obj{Device} \Sba{mustDisable}{Feature} & The \emph{device} should ensure that the \emph{feature} is disabled. \\
-  TODO & Fill this in with most common ones and add longer list to appendix. \\
+  \textit{Subject} \texttt{can}\textit{Action}  & The \emph{subject} is allowed to perform the \emph{action}. \\
+  \textit{Subject} \texttt{has}\textit{Action}  & The \emph{subject} has performed the \emph{action}. \\
+  \textit{Subject} \texttt{is}\textit{Propery}  & The \emph{property} is true about the \emph{subject}. \\
+  \textit{Subject} \texttt{must}\textit{Action} & The \emph{subject} is obliged to perform the action. \\
   \bottomrule \\
 \end{tabular}
-\label{tab:predicates}
-\caption{Excerpt of Generalized predicates for BYOD policies.}
+
+\subsubsection{Acknowledgements}
+
+A common idiom in each of the policies is requiring an acknowledgement from the device owner that a policy exists, or that the company may have certain rights.  
+These rules are interesting as there is often not a technical means to enforce them.
+For example both the NHS and HiMSS policies have rules stating that they cannot be held liable if a user loses personal information if their device needs to be erased to protect personal information.
+
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMSS}}:~\itshape 
+I agree to hold XYZ Health System harmless for any loss relating to the administration of PDA/Smartphone connectivity to XYZ Health System systems including, but not limited to, loss of personal information stored on a PDA/Smartphone due to data deletion done to protect sensitive information related to XYZ Health System, its patients, members or partners.}]
+'user' says 'user' hasAcknowledged('data-loss-policy').
+\end{lstlisting}
+
+\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape 
+Individuals who have personal data of any kind stored on a corporately issued mobile device must be aware that in the event of loss of the device the above data wipe will include removal of all personal data.}]
+'nhs-trust' says Staff can-say
+	Staff hasAcknowledged('data-loss-policy') 
+	if Staff isStaff.
+\end{lstlisting}
+
+This idiom is particularly common in the \ac{HiMSS} policy where over the rules require some level of acknowledgement, and a third require an acknowledgement alone (\autoref{tab:idioms}).
+Similarly in the longer NHS policy almost a quarter of the rules require the subject to acknowledge a rule.
+In contrast the more technically focussed SANS policy has only one occurance of the idiom, when describing the other policies a user should be following in order to use a mobile device.
+
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
+Users MUST agree to the email security/acceptable use policy and eventually to the eCommerce security policy.}]
+'company' says User canUse(Device) 
+  if User hasDevice(Device),
+     User hasAcknowledged('email-security'),
+     User hasAcknowledged('acceptable-use'), 
+     User hasAcknowledged('ecommerce-security').
+\end{lstlisting}
+
+The reliance in both the NHS and \ac{HiMSS} policies suggests that \ac{BYOD} policies often go beyond technical rules; something ignored in previous work on policy formalisation.
+These policies often require that employees \emph{follow the rules}, and have an awareness of the systems surrounding them.
+
+\begin{table}\centering
+  \begin{tabular}{l c c c}
+    \toprule
+    Idiom                    & \ac{HiMSS}   & NHS      & SANS    \\
+    \midrule
+    Acknowledgement          & 5 (33\%) & 9 (18\%) & 1 (3\%) \\
+    Partial Acknowledgement  & 3 (20\%) & 2 (4\%)  & 0       \\
+    \bottomrule             \\
+  \end{tabular}
+  \label{tab:idioms}
+  \caption{Occurences of policy idioms in each of the \ac{BYOD} policies.}
 \end{table}
 
-Using AppPAL we were able to translate 23 out of the 60 requriements in the NHS policy and XX out of the XX requirements in the SANS policy.
-\textbf{\large TODO: check these numbers properly}.
-Sometimes this was because the requirement was not a policy rule but rather something the device owner should be aware of.
-For example:
-
-\begin{quote}\footnotesize
-  \textbf{NHS}: \itshape 
-``In receiving a mobile device from the Trust, the individual receiving and using
-the device accepts that the device can be used to communicate through all
-corporate channels including voice calls, emails and texts and where appropriate
-enabled web applications, during working hours.''
-\end{quote}
-
-\begin{quote}\footnotesize
-  \textbf{NHS}: \itshape 
-``Individuals who have personal data of any kind stored on a corporately issued
-mobile device must be aware that in the event of loss of the device the above
-data wipe will include removal of all personal data.''
-\end{quote}
-
-In these cases no rule is needed as there is no behaviour we wish to control. 
-If we wished to use AppPAL to enforce these policies we could require device owners issue a statement acknowledging these limitations, similar to an EULA agreement.
-
-{\large \textbf{...actually this sounds like a good idea.  What percentage of these policies are acknowledgements.}}
-
-% 5 pages
-\section{The Details}
-\label{sec:details}
-
-\newcommand{\facts}[1]{\ensuremath \textsf{facts}\left(#1\right)}
-\newcommand{\reachable}[2]{\ensuremath \textsf{reachable}\left(#1,~#2\right)}
 \subsection{Checking Reachability}
@@
 If a rule is reachable then we would expect there to be some combination of facts that could satisfy the rule.
-If there are no facts that could satisfy the rule then this hints that the policy may be incomplete as there are policy rules that can never be satisfied.
+If there are no facts that could satisfy the rule then this suggests that the policy may be incomplete as there are policy rules that can never be satisfied.
 
-\begin{equation}
-  \reachable{R}{AC}\equiv\exists p\in\facts{R}:~p\theta\subseteq AC
-\end{equation}
-Where $\facts{R}$ is the set of sets of facts and constraints that would satisfy rule $R$, and $\theta$ is an alpha-renaming of any variables in the set of facts such that equivalent free variable have the same name.
+We search for reachable assertions using the algorithm in Listing~\ref{alg:reachable}:
+it produces a set of pairs of predicates and speakers where the a pair of a speaker and a predicate indicates that that speaker may say something about that predicate.
+We search over all the assertions in the AppPAL assertion context.
+If all of an assertion's conditionals (the facts in the if part) are reachable (or it has none) then the speaker and predicate are added to the reachable set.
+If the statement is a can-say statement then we additionally check if the delegated predicate is reachable from the delegated speaker, and if so mark the delegated statement as reachable from the speaker who made the can-say statement.
 
-We find all reachable assertions using the algorithm shown in~\autoref{alg:reachability} building a set of pairs of speakers and predicates.
-If a pair is in the set then that predicate is said to be reachable from that speaker (i.e.~we can infer the speaker may say something about that predicate).
-We search over all the assertions in the assertion context.
-If all of an assetion's conditionals (the facts in the \emph{if} part) are reachable (or it has none) then the speaker and predicate is added to the reachable set.
-If the statement is a can-say statement then we additionally check if the delegated predicate is reachable from the delgated speaker, and if so mark the delegated statement as reachable from the speaker who made the can-say statement.
-
-\begin{figure}
-\begin{algorithmic}
-\Procedure {Reachable}{AC}
-  \State reachable $\gets \emptyset$
-  \Repeat
-    \State iterate $\gets$ False
-    \ForAll {$a \in$ AC}
-      \State $e \gets $Speaker($a$)
-      \State $p \gets $Predicate($a$)
-      \If {$p~$:: CanSay $\wedge$ $(e,p)\in$ reachable}
-        \If {$($Delegator($p$)$,$ Delegation($p$)$)~\in$ reachable}
-          \If {$\forall c \in$ Conditions($a$)$.~(e,$ Predicate($c$)$) \in$ reachable}
-            \State reachable $\gets$ reachable $\cup$ ($e,$ Delegation($p$))
-            \State iterate $\gets$ True
-          \EndIf
-        \EndIf
-      \ElsIf {$(e,p)\not\in$ reachable}
-        \If {$\forall c \in$ Conditions($a$)$.~(e,$ Predicate($c$)$) \in$ reachable}
-          \State reachable $\gets$ reachable $\cup~(e,p)$
-          \State iterate $\gets$ True
-        \EndIf
-      \EndIf
-    \EndFor
-  \Until {iterate = True}
-  \State \textbf{return} reachable
-\EndProcedure
-\end{algorithmic}
-\caption{Procedure for finding all reachable assertions.}
-\label{alg:reachability}
-\end{figure}
+\begin{lstlisting}[language=Python,float,caption={Procedure for finding all reachable assertions.},label={alg:reachable}]
+def reachable(ac) -> set:
+  reachable = new set()
+  iterate = True
+  while iterate == True:
+    iterate = False
+    for assertion in ac:
+      e = a.speaker
+      p = a.predicate
+      if p.isCanSay() and (e, p) in reachable:
+        if (p.delegator, p.delegation) in reachable:
+          if for all c in a.conditions: (e, c.predicate) in reachable:
+            reachable.add((e, p.delegation))
+            iterate = True
+      else if not (e, p) in reachable:
+        if for all c in a.conditions; (e, c.predicate) in reachable:
+          reachable.add((e, p))
+          iterate = True
+  return reachable
+\end{lstlisting}
 
@@
 
-When checking for redundency we are looking for rules that will need the same or a subset of the ground facts and constraints to satisfy it.
-If one rule makes another redundant then the first rule will not only always be satisfied whenever the redundant one is, but may require less facts in order to satisfy it.
+To check redundancy we convert the policy into a graph structure.
+Each statement we might want to prove (a goal) becomes a node in the graph, its children are organised into sets of goals (a proof), where if all the goals in any set were proved true then the goal node would also become true.
+If a node has an empty set of goals to prove it is a fact.
+Once the graph has been constructed (taking into account delegation and unification with other statements), the graph is flattened by applying the flatten procedure (Listing~\ref{alg:flatten}) repeatedly until a fixed point is reached.
+When the graph has been flattened we look for redundancy by looking for goals with proofs that are subsets of their other proofs, implying that if the larger proof is true, then the shorter proof will always be true too;
+  and different goals with identical proofs, implying that the two decisions are not independent~(Listing~\ref{alg:redundancy}).
 
-Formally for any two rules $R$ and $R^\prime$; $R$ is said to be \emph{made redundant} by $R^\prime$ ($R\geq R^\prime$) if:
-\begin{equation}
-  R \geq R^\prime \equiv \forall p \in \facts{R}.~\exists p^\prime \in \facts{R^\prime}:~p\theta \supseteq p^\prime\theta
-\end{equation}
-If $R\geq R^\prime$ and $R\leq R^\prime$ then the two rules are equivalent (i.e.
-proving one will always prove the other).
-This is particularly interesting if the two rules are for different predicates as it indicates that the two predicates may be equivalent also.
+\begin{lstlisting}[language=Python,float,caption={Procedure for flattening the redundancy graph.},label={alg:flatten}]
+def flatten(graph):
+  for goal in graph.goals:
+    hoist = true
+    for proof in graph[goal]:
+      for proof_goal in proof:
+        if not proof_goal.is_fact:
+          hoist = False
+          break
+    if hoist == True:
+      hoist(graph, goal)
 
+def hoist(graph, target):
+  for parent in graph[target].parents:
+    for proof in parent.proofs:
+      if proof.uses(target):
+        for replacements in graph[target].parents:
+          parent.proofs += proof.replace(target, replacements)
+        parent.proofs -= proof
+\end{lstlisting}
+
+\begin{lstlisting}[language=Python,float,caption={Procedure to check for redundancy.},label={alg:redundancy}]
+def check_redundancy(graph) -> boolean:
+  for goal in graph.goals:
+    for a in graph[goal]:
+      for b in graph[goal]:
+        if a >= b and a.goals.subset(b.goals):
+          if b.goals.subset(a.goals):
+            warn(a+" has multiple equivalent proofs") 
+          else
+            warn(a+" has a redundant proof")
+    for other_goal in graph.goals:
+      if goal > other_goal:
+        for a in graph[goal]:
+          for b in graph[other_goal]:
+            if not (a.goals.is_empty() or b.is_empty())
+               and a.goals == b.goals:
+              warn(a+" and "+b+" have equivalent proofs")
+\end{lstlisting}
+
+\section{Related Work}
 % 1-2 pages
-\section{Related Work}
 \label{sec:related}
 
-The policies present in the NHS organisation and hospitals have been an attractive target for the application of formal methods.
-There have been multiple attempts to present formalizations of electronic health records in both the Ponder~\cite{blobel_authorization_2004} and Cassandra~\cite{becker_cassandra_2004} policy languages.
+Armando~\etal~created a BYODroid~\cite{armando_bring_2013}; \ac{BYOD} framework to enforce \ac{BYOD} policies by extending Featherweight Java~\cite{igarashi_featherweight_2001} with a type system that can describe how Android apps interact.
+Whilst this work gives a framework for enforcing policies on Android, it doesn't help policy authors decide what the policies should be to begin with.
 
+\section{Conclusions}
 % 0.5 pages
-\section{Conclusions}
 \label{sec:conclusions}

### 2016-08-01 paper.tex
updates to the paper

@@
 
-\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
+\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
 Users MUST agree to the email security/acceptable use policy and eventually to the eCommerce security policy.}]
@@
 The reliance in both the NHS and \ac{HiMSS} policies suggests that \ac{BYOD} policies often go beyond technical rules; something ignored in previous work on policy formalisation.
-These policies often require that employees \emph{follow the rules}, and have an awareness of the systems surrounding them.
+These policies often require that employees \emph{follow the rules}, and have an awareness of the systems surrounding them;
+  something which cannot be checked by static analysis.
 
@@
     Partial Acknowledgement  & 3 (20\%)   & 2 (4\%)   & 0         \\
+    Delegation               & 6 (40\%)   & 15 (30\%) & 14 (42\%) \\
     \bottomrule             \\
@@
 \end{table}
+
+\subsection{Delegating Responsibility}
+
+When describing rules employees must follow, there are often aspects of the rules that require other employees and departments to sign off on.
+For example in the NHS policy for an employee to be able to use a \ac{BYOD} phone the finance director and relevant assistant director must both sign off on it.
+
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{NHS}}:\itshape
+Upon receipt of the completed application form, allocation will need to be authorised by the relevant Assistant Director and the Director of Finance.}]
+'nhs-trust' says Form isAuthorised
+  if Staff hasAppliedForPhone(Form),
+     FD hasAuthorised(Form),
+     FD isFinanceDirector,
+     AD hasAuthorized(Form),
+     AD isAssistantDirector,
+     AD isManagerOf(Staff).
+
+'nhs-trust' says X can-say X hasAuthorised(Form)
+  if X isEmployee,
+     Form isForm.
+\end{lstlisting}
+
+These delegations start to show some of the different responsibilities of the different employees.
+Later in the NHS policy we find a rule that it is an employee's line manager's responsibility to sign off on the use of a paid app:
+
+\begin{lstlisting}[title={\footnotesize\textbf{NHS}:\itshape
+Use of paid apps must be agreed in advance with the device holder's line manager and their should be a demonstrable benefit.}]
+'nhs-trust' says Manager can-say
+  Staff canInstall(App)
+  if App isPaidApp,
+     Manager isManagerOf(Staff).
+\end{lstlisting}
+
+This highlights that the policy is incomplete: different managers may have different criteria for what constitutes an \emph{demonstrable benefit}.
+It is not the policy's job to state precisely for the manager what is beneficial or not, but rather to indicate that it is the manager's call.
+The manager could specify their requirements as a formal series of rules expressed in AppPAL and enforced by their favorite \ac{MDM} tool, or they could specify apps on a case by case basis for each employee that asks.
+
+These relationships are often ignored by existing BYOD policy tools, which focus on how a manager's policy could be implemented by restricting the functionality of the individual apps.
+In this work we're more interested in who makes the decision and what decisions they make rather than the mechanism for enforcing their policies.
+When modelling these policies it is necessary to use a formalism that supports delegation, such as AppPAL, to capture these requirements.
+
+Using AppPAL we model these kinds of decisions using delegation with the \emph{can-say} statement.
+All three policies contain some form of delegation relationships: in the \ac{HiMSS} policy it is the responsibility of the management to decide is a device must be wiped:
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMSS}}:\itshape
+I agree that the PDA/Smartphone can be wiped by XYZ Health System upon the decision of XYZ Health System management.}]
+'user' says 'xyz-health-system' can-say
+  Device mustBeErased
+  if 'user' hasDevice(Device).
+\end{lstlisting}
+In the SANS policy only the devices identified by the IT-Department can connect to internal networks and decide the access control policies:
+\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMSS}}:\itshape
+All devices, including handhelds that have to connect to internal networks MUST be identified by IT department}]
+'company' says Device canConnectTo(Network)
+  if Network isInternalNetwork,
+     Device isIdentified.
+
+'company' says 'it-department' can-say
+  Device isIdentified
+  if Device isDevice.
+\end{lstlisting}
+All three policies make extensive use of delegation patterns. 
+In the NHS policy acknowledgements are implemented by delegating to the user to make the statement, where as in the \ac{HiMSS} policy they are given as straight assertions as the \ac{HiMSS} policy is implemented from the view of the user.
+The SANS policy supposes your company has several different departments each with their own specialities (such as IT, security and desktop support personnel), each of which is delegated to for approriate decisions.
 

### 2016-08-01 paper.tex
updates done in edinburgh

--- 2016-07-13 paper.tex
@@
 
+\subsection{Delegation}
+\textit{About a page to be merged from the other machine you forgot to push. }
+
 \subsection{Checking Reachability}
@@
 
-Armando~\etal~created a BYODroid~\cite{armando_bring_2013}; \ac{BYOD} framework to enforce \ac{BYOD} policies by extending Featherweight Java~\cite{igarashi_featherweight_2001} with a type system that can describe how Android apps interact.
-Whilst this work gives a framework for enforcing policies on Android, it doesn't help policy authors decide what the policies should be to begin with.
+Armando~\etal~created a BYODroid~\cite{armando_bring_2013}; \ac{BYOD} framework
+to enforce \ac{BYOD} policies by extending Featherweight
+Java~\cite{igarashi_featherweight_2001} with a type system that can describe how
+Android apps interact. Whilst this work gives a framework for enforcing policies
+on Android, it doesn't help policy authors decide what the policies should be to
+begin with. They went on to show how BYODroid could be used to enforce a {NATO}
+Communications and Information Agency policy~\cite{armando_developing_2016}.
+Armando~\etal also proposed creating curated app stores based on corporate BYOD
+policies~\cite{armando_enabling_2014}. To do this they analysed a BYOD policy
+taken from the US Government BYOD Security
+Guidelines~\cite{digital_services_advisory_group_and_federal_chief_information_officers_council_bring_????}~and
+translated it into rules in the CONspec~\cite{aktug_conspec_2008} language which
+allows a low-level description of app behavior based on Java code. They then
+used these rules to filter apps for their store from a selection of 860 apps
+taken from the \emph{top selling free} section of the Google Play store.
+
+Martinelli~\etal~applied the UCON security model to describe \ac{BYOD} policies
+for Android~\cite{martinelli_enhancing_2016}. Using this model they created an
+enforcement framework that could alter the permissions granted to Android apps
+at run time. This framework requires a rooted device for most Android versions,
+however, they do claim it should work without modification on the latest
+version. Again like the work of Armando~\etal, this gives a framework for
+enforcement but without consideration of what the policies should be enforcing.
+
+Hong~\etal~focussedon the network aspects of a \ac{BYOD} policy with their
+PBS-Droid tool~\cite{sungmin_hong_towards_2016}.  PBS-Droid allowed BYOD policy
+authors the ability to sandbox individual apps into their own software-defined
+networks giving finer control over how individual apps interact with a companies
+networks and allowing for dynamic policy enforcement and network configuration.
+
+For more general mobile policies, there has been extensive work in both
+modeling and enforcing various policies.  Sadeh~\etal~surveyed users to find
+their privacy preferences and found four different types of user each with
+different privacy policies~\cite{sadeh_understanding_2009}.  Various tools have
+been proposed that can enforce app behavior policies.  Aurasium works by
+repackaging apps to use modified APIs~\cite{xu_aurasium:_2012}.  These APIs can
+enforce policies to protect a user's personal information or prevent sending
+premium rate text messages.  Another tool, CRePE~\cite{conti_crepe:_2010}
+modifies Android to provide a framework that can enforce policies based
+on the contextual information, such as the device's location or battery level.
+The SEAR tool takes a different approach and actively scans apps for
+vulnerabilities over collections of apps.  It then outputs security policies
+that users might want to enforce in order to prevent attacks from occurring in
+the first place~\cite{bagheri_practical_2016}.
 

### 2016-10-06 paper.tex
ch ch ch changes

--- 2016-08-01 paper.tex
@@
-\documentclass[a4paper]{article}
+\documentclass{article}
 
 \usepackage{microtype}
-\usepackage{booktabs}
-\usepackage{hyperref}
+
 \usepackage{acronym}
-\acrodef{MDM}{mobile device management}
-\acrodef{BYOD}{bring your own device}
-\acrodef{HiMSS}{Healthcare Information and Management Systems Society}
+\acrodef{MDM}{Mobile Device Management}
 
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-
-% \apppal{} and Listing formatting
-\usepackage{listings}
-\lstdefinelanguage{AppPAL}{%
-  morekeywords={if,inf,says,where,true,false},
-  otherkeywords={can-act-as,can-say},
-  sensitive=true,
-  morestring=[b]',
-  literate={\ inf\ }{{$\infty$}}5
-}[keywords,strings]
-\lstset{%
-  basicstyle=\ttfamily\footnotesize{},
-  stringstyle=\sffamily\footnotesize{},
-  keywordstyle=\ttfamily\slshape\footnotesize{},
-  language=AppPAL,
-  columns=flexible,
-  frame=single,
-  framesep=0.5pt,
-  framexleftmargin=2.5pt,
-  framexrightmargin=2.5pt,
-}
-
-\newcommand{\dotdotdot}[1]{#1}
-\newcommand{\etal}[0]{et~al{.}\@}
-\newcommand{\apppal}[0]{App{P}{A}{L}}
-
-\title{Case Study Reasoning about BYOD Policies in the Healthcare Sector}
+\title{Capturing Policies for BYOD}
 \author{Joseph Hallett and David Aspinall}
@@
 
-% 4 sentences
 \begin{abstract}
-  Implementing a \ac{BYOD} policy can be difficult especially in hospitals, where there is a need to balance privacy requirements against the need to treat patients.
-  It is common to publish a \ac{BYOD} policy to describe how staff should use their devices in the workplace.
-  These policies are written using natural language, and employees are expected to read and follow them: this can be tricky as the policies are often poorly specified.
-  We show how \apppal{}, a policy language for mobile app preferences, can be used as a tool for reasoning about policies and how it found issues with BYOD policies used in hospitals and as guidelines.
+  % 4 Sentences
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:intro}
+% Describe the problem
+% State my contribution
 % 1 page
-\label{sec:intro}
-% 1. Describe the problem
-%    - BYOD hard
-%    - Hospitals are an interesting case
-%    - Need tools for reasoning about policies not enforcing
 
-Employees bringing their own devices to work has become the norm:
-  around 70\% of all companies having some form of \ac{BYOD} scheme~\cite{schulze_byod_2016}.
-In hospitals the use of personal devices as part of patient care is particularly common.
-One survey from the UK found that around 80\% of surgical doctors were willing to use their personal devices for work, with 85\% regularly using them to look up information, and 50\% having medical apps already installed on their phones~\cite{patel_uk_2015}.
-Another from the US found similar numbers of physicians regularly using their devices as part of their work~\cite{moyer_managing_2013}.
+Employees bring their own devices to work.  
+In the past employees might have had a dedicated company device. 
+Today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016} and, in some fields,
+  85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}. 
+This creates a challenge for IT departments.
+They need to control access to company resources, but have limited control over the devices used to access them.
 
-Several healthcare organisations have started looking into using mobile devices as part of patient care, and found that these devices may be helpful~\cite{prgomet_impact_2009,faulds_feasibility_2016}.
-Organisations, such as the US's \ac{HiMSS}~\cite{seymour_mobile_2012}, have published guidelines helping hospital administrators implement BYOD policies in their workplaces.  
-Additionally general advice for other organisations seeking to implement BYOD policies is available from SANS~\cite{nicholas_r._c._guerin_security_2008}, NIST~\cite{souppaya_guidelines_????}, HP~\cite{hp_byod_????} and the CESG~\cite{cesg_byod_2015} amongst others.
-The NHS has even published guidelines for developers seeking to develop apps for use in hospitals~\cite{prgomet_impact_2009}.
+One common solution to control devices is to require users to agree to follow a policy.
+These policies take the form of user agreements, written in natural language, which describe how the devices should be used and configured.
+Various guides are available for companies wishing to implement a policy from governments, standards bodies, and companies seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
+On top of user agreements companies may also use \ac{MDM} software, which can enforce some policies automatically.
+The use of \ac{MDM} software does not guarantee compliance however.
+One survey found over 50\% of companies with \ac{MDM} software still have uncompliant devices in their networks~\cite{mobileiron_security_labs_q4_2015}.
 
-Whilst there has been a push for bring \ac{BYOD} and mobile devices into the healthcare industry there has been push back against having mandated security policies automatically enforced.  
-The use of \ac{MDM} software to enforce policies can be problematic:
-  a survey found that even for companies that use \ac{MDM}, over half of them still have devices in their networks that do not meet their policies~\cite{mobileiron_security_labs_q4_2015}.
-In the healthcare industry the circumvention of security controls by physicians is normal, and in some cases taught as good practice~\cite{koppel_workarounds_2015}.
-The use of tools to automatically enforce policies in a hospital environment is not always the right approach:
-\emph{Moyer} concluded that:
-\emph{``Since mobile devices are already being used in hospitals, an approach toward educating the users instead of trying to control the technology may be more practical in the short term''}~\cite{moyer_managing_2013}.
-Similarly \emph{Faulds}~\etal~echo this sentiment noting that \emph{``concerns about data protection are legitimate''}, and going on to \emph{``encourage the Information Commissioner’s Office, in consultation with professional medical organisations, to publish guidance specific to the use of BYOD technology''}~\cite{faulds_feasibility_2016}.
+BYOD policies are becoming more complex and \ac{MDM} software is becoming more proficient.
+There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
+Enforcing a policy is only part of the problem, however.
+BYOD policies are specified informally using natural language, and they contain more than just access control decisions.
+These policies describe trust relationships inside the company between the IT departments, users, and HR each who may be delegated to make decisions and provide rules.
+Policies contain rules that require employees to acknowledge risks, and regulations.
+An antivirus or \ac{MDM} program may be used to \emph{implement} part of a policy, but it is the policy which \emph{specifies} which software to use and when; there is no automatic way to check how a policy has been implemented and by what.
 
-The conflict between patient care and data security, as well as \ac{MDM} software being relatively limited, has lead to \ac{BYOD} policies being published as natural language documents that employees are supposed to agree to follow.
-Using natural language in policies can create ambiguities, and lead to poorly specified rules.
-If a policy is poorly specified and confusing then its subjects may chose to ignore it or misinterpret it's rules.
-If we wish to improve our policies we need to be able to reason about them clearly.
-We need to be able to identify problems before they happen.
+This paper describes a formalization of five BYOD policies using AppPAL---a policy language for mobile device privacy preferences~\cite{hallett_apppal_2016}.
+Using our formalization we identify the common concerns and trust relationships in these policies.
+We show how AppPAL can be used to implement these policies, and describe precisely the different trust relationships.
+AppPAL is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies.
+We have found, however, that it is a good fit for other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
 
-% 2. State your contribution
-%    - Standard lexicon of predicates and idioms
-%    - Formalisation of several BYOD policies
-%    - Identification of problems in said policies
-%    - Methods for inferring when \apppal{} policies are problematic
-
-\apppal{} is a policy language for describing mobile app preferences~\cite{hallett_apppal_2016}.
-We show how \apppal{} can also be used to formalise \ac{BYOD} policies by translating three \ac{BYOD} policies into \apppal{}.
-One is taken from an NHS hospital trust~\cite{kennington_mobiles_2014}. 
-Another is from \ac{HiMSS} published as an example policy to help hospitals design their own policies~\cite{healthcare_information_and_management_systems_society_mobile_2012}.
-The third is a  general policy from SANS to help organisations to develop their own \ac{BYOD} schemes~\cite{nicholas_r._c._guerin_security_2008}.
-Using these policies we suggest a standard lexicon of BYOD predicates and idioms for using when reasoning about \ac{BYOD} policies. 
-With the AppPAL formalisation we show how problems with reachability and redundancy can be identified automatically, and suggest where the policies could be improved.
-
-%\section{My Idea}
-\section{Formalising policies using AppPAL}
+\section{My Idea}
+\label{sec:idea}
 % 2 pages
-\label{sec:idea}
-
-Each of the three policies we have chosen to look at has a different style in expressing their rules.
-Despite the different styles, however, there are areas where the policies overlap and make similar rules.
-We present the policies here as excerpts alongside their AppPAL translations.
-The full translations and original policies can be found online\footnote{\url{https://github.com/apppal/apppal-byod-policy-translations}}.
-
-The \ac{HiMSS} policy is taken from their \emph{Mobile Security Toolkit} for assisting healthcare organisations and security practitioners in managing the security of their mobile computing devices. 
-\ac{HiMSS} itself is an American non-profit company advising healthcare companies on IT.
-Their policy~\cite{healthcare_information_and_management_systems_society_mobile_2012} takes the form of a short user agreement that it's subjects are expected to sign.  
-This gives a different style to the other policies: policies are expressed from the perspective of the subject, rather than as decrees from the policy authors.
-It also contains many rules where the subject is expected to acknowledge a risk or regulation.  
-For example in one rule the subject is required to acknowledge that networks can be unreliable:
-\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMSS}}:~\itshape 
-I understand and accept that synchronisation relies on one or more cellular network providers and the Internet, and that both are subject to slowdowns and outages of extended duration that are beyond the control of IT.}]
-'ac{HiMSS}' says User can-say User hasAcknowledged('internet-unreliable')
-  if User isUser.
-
-'ac{HiMSS}' says User hasAgreeedToPolicy
-  if ...,  User hasAcknowledged('internet-unreliable').
-\end{lstlisting}
-These rules are interesting as on the face of them they do not say anything about security or how the subject should use their device.
-We translate rules like this into acknowledgement statements in AppPAL, which can be collected into a check that the user has acknowledged everything they should have done.
-
-The NHS policy~\cite{kennington_mobiles_2014}, in contrast, is a more prescriptive document.
-It is taken from the \emph{Torbay and Southern Devon NHS trust}\footnote{The NHS is organised into a series of \emph{trusts}, each of which contains several hospitals.
-A doctor may work at several hospitals within a trust.}.
-It is interesting in part because it represents a \ac{BYOD} policy in actual use, but also because it is significantly more complex than most \ac{BYOD} policies as it goes beyond basic requirements such as requiring encryption by describing relationships between organisational groups within the trust and precise organisation specific requirements.
-Like the \ac{HiMSS} policy it has rules that just require an acknowledgement (for example Rule~3.2 requires that the device owner \emph{``accepts that the device can be used to communicate through all corporate channels''}) but there is a greater emphasis on the specifying the responsibilities of different groups and roles within the hospital.
-For example in Rule~2.4 a staff member who has applied for a phone needs it to be signed off by a finance director and a relevant assistant director:
-\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape 
-Upon receipt of the completed application form, allocation will also need to be authorised by the relevant Assistant Director and the Director of Finance.}]
-'nhs-trust' says Staff hasAppliedForPhone(Form)
-  if Staff hasSubmitted(Form),
-     'assistant-director' hasAuthorized(Form),
-     'finance-director' hasAuthorized(Form).
-
-'nhs-trust' says X can-say 'assistant-director' hasAuthorized(Form)
-  if X can-act-as 'assistant-director',
-     Staff hasSubmitted(Form),
-     X isManagerOf(Staff).
-\end{lstlisting}
-
-The SANS Institute is a private US company that specialise in information security.
-Whilst their example \ac{BYOD} policy is not healthcare specific, it is still interesting as there advice is widely regarded, and contains more general security advice.
-The SANS \ac{BYOD} policy is split into sections that describe how device features, such as Bluetooth or OTA provisioning, should be configured.
-Some rules are acknowledgements: for example the use of internet services in Section~2.8 requires acknowledgement of the email security, acceptable use and an eCommerce security policy.
-Many rules describe when device features should be disabled, in contrast to the other two policies which tend to favor describing when features can be used.
-For example in Section~2.4 the SANS policy states that infrared support should be disabled if Bluetooth is available: this is easily expressed in AppPAL:
-\begin{lstlisting}[title={\footnotesize\textbf{SANS}:~\itshape 
-Infrared supported MUST be disabled if Bluetooth connectivity is supported.}]
-'sans' says Device shouldDisable('irda')
-  if Device hasFeature('bluetooth').
-\end{lstlisting}
-
-% - Brief intro to apppal but mostly link to existing paper
-%   - Why apppal?
-% - Give an overview of the different policies we're looking at
-% - Show differences and similarities between the policies
-% - Make sure to tie it all into healthcare, but also be clear this is more general
-
-\subsection{Why AppPAL?}
-
-AppPAL is an authorization language designed to model app preferences~\cite{hallett_apppal_2016}.
-It was created by instantiating the SecPAL language (which was designed for access control in distributed computing environments~\cite{becker_secpal:_2010}) with predicates and checking functions for describing apps.
-It supports delegation, roles and constraints including arbitrary external checking programs to describe precisely what apps users might want to use.
-
-AppPAL (and SecPAL) is designed to be readable.
-It is powerful enough to describe complex policies and rules, whilst also being semantically simple having only three rules for evaluation.
-When reasoning about \ac{BYOD} policies, it simplicity and readability gives it an advantage over more powerful languages such as XACML~\cite{oasis_extensible_2013} as a rule written in AppPAL can be comprehended quickly.
-Whilst an end-user might not write a policy in a formal language, an IT administrator might wish to use a formal language to aid with reasoning and checking the policies they write.
-AppPAL gives these IT professionals an easy entry point to formal methods, without over complicating things.
-Early experiments showed that AppPAL can describe \ac{BYOD} policies, and help make clarify mistakes ahead of time~\cite{hallett_specifying_2016}.
-
-An AppPAL evaluation library has been developed that runs on Java and Android and is available online\footnote{\url{https://github.com/apppal/libapppal}}.
-As part of this work we have developed (in Section~\ref{sec:details}) a series of policy analysis tools for AppPAL.
-These are compiled alongside the main AppPAL interpreter and may be run from the command-line.
 
 \section{The Details}
+\label{sec:details}
 % 5 pages
-\label{sec:details}
-
-% - Show the standard predicates and explain how they help to give structure to apppal policies
-%   - show how they are tied into the trust relationships with patient care (i.e. consent)
-% - Show the consistency checker
-%   - Give algorithm and intuition
-%   - Show examples where it has found problems
-% - Show redundancy checks
-%   - Algorithm and intuition
-%   - Examples where problems were found
-
-\subsection{Policy Idioms}
-
-When translating the \ac{BYOD} policies we found types of actions or checks being performed.
-There are checks if things are of a particular type (for~example in the NHS policy rule~2.1  a staff member is elligable for a mobile device if they are lone-working, work out-of-hours and work out-of-office).
-There are checks if an action is allowed (such~as whether they can connect to an wireless access point in the \ac{HiMSS} policy);  and checks to see if an action has already been performed (like the need to have acknowledged a policy in all three policies).
-We also find policy rules that state the user should perform an action (such as disabling phone functionality in the SANS policy, or an obligation to report a phone missing in the NHS policy).
-
-To help structure the formalisation we give the AppPAL predicates used to translate the policies prefixes.
-The prefixes help to clarify the meaning of each predicate and groups the predicates into four categories.
-
-\begin{tabular}{l l}
-  \toprule
-  Predicate & Meaning \\
-  \midrule
-  \textit{Subject} \texttt{can}\textit{Action}  & The \emph{subject} is allowed to perform the \emph{action}. \\
-  \textit{Subject} \texttt{has}\textit{Action}  & The \emph{subject} has performed the \emph{action}. \\
-  \textit{Subject} \texttt{is}\textit{Propery}  & The \emph{property} is true about the \emph{subject}. \\
-  \textit{Subject} \texttt{must}\textit{Action} & The \emph{subject} is obliged to perform the action. \\
-  \bottomrule \\
-\end{tabular}
-
-\subsubsection{Acknowledgements}
-
-A common idiom in each of the policies is requiring an acknowledgement from the device owner that a policy exists, or that the company may have certain rights.  
-These rules are interesting as there is often not a technical means to enforce them.
-For example both the NHS and HiMSS policies have rules stating that they cannot be held liable if a user loses personal information if their device needs to be erased to protect personal information.
-
-\begin{lstlisting}[title={\footnotesize\textbf{\ac{HiMSS}}:~\itshape 
-I agree to hold XYZ Health System harmless for any loss relating to the administration of PDA/Smartphone connectivity to XYZ Health System systems including, but not limited to, loss of personal information stored on a PDA/Smartphone due to data deletion done to protect sensitive information related to XYZ Health System, its patients, members or partners.}]
-'user' says 'user' hasAcknowledged('data-loss-policy').
-\end{lstlisting}
-
-\begin{lstlisting}[title={\footnotesize\textbf{NHS}:~\itshape 
-Individuals who have personal data of any kind stored on a corporately issued mobile device must be aware that in the event of loss of the device the above data wipe will include removal of all personal data.}]
-'nhs-trust' says Staff can-say
-	Staff hasAcknowledged('data-loss-policy') 
-	if Staff isStaff.
-\end{lstlisting}
-
-This idiom is particularly common in the \ac{HiMSS} policy where over the rules require some level of acknowledgement, and a third require an acknowledgement alone (\autoref{tab:idioms}).
-Similarly in the longer NHS policy almost a quarter of the rules require the subject to acknowledge a rule.
-In contrast the more technically focussed SANS policy has only one occurance of the idiom, when describing the other policies a user should be following in order to use a mobile device.
-
-\begin{lstlisting}[title={\footnotesize\textbf{\ac{SANS}}:~\itshape 
-Users MUST agree to the email security/acceptable use policy and eventually to the eCommerce security policy.}]
-'company' says User canUse(Device) 
-  if User hasDevice(Device),
-     User hasAcknowledged('email-security'),
-     User hasAcknowledged('acceptable-use'), 
-     User hasAcknowledged('ecommerce-security').
-\end{lstlisting}
-
-The reliance in both the NHS and \ac{HiMSS} policies suggests that \ac{BYOD} policies often go beyond technical rules; something ignored in previous work on policy formalisation.
-These policies often require that employees \emph{follow the rules}, and have an awareness of the systems surrounding them.
-
-\begin{table}\centering
-  \begin{tabular}{l c c c}
-    \toprule
-    Idiom                    & \ac{HiMSS}   & NHS      & SANS    \\
-    \midrule
-    Acknowledgement          & 5 (33\%) & 9 (18\%) & 1 (3\%) \\
-    Partial Acknowledgement  & 3 (20\%) & 2 (4\%)  & 0       \\
-    \bottomrule             \\
-  \end{tabular}
-  \label{tab:idioms}
-  \caption{Occurences of policy idioms in each of the \ac{BYOD} policies.}
-\end{table}
-
-\subsection{Delegation}
-\textit{About a page to be merged from the other machine you forgot to push. }
-
-\subsection{Checking Reachability}
-
-Inference in AppPAL happens by collecting ground facts and constraints that satisfy rules.
-These rules are then combined to form a policy. 
-If a rule is reachable then we would expect there to be some combination of facts that could satisfy the rule.
-If there are no facts that could satisfy the rule then this suggests that the policy may be incomplete as there are policy rules that can never be satisfied.
-
-We search for reachable assertions using the algorithm in Listing~\ref{alg:reachable}:
-it produces a set of pairs of predicates and speakers where the a pair of a speaker and a predicate indicates that that speaker may say something about that predicate.
-We search over all the assertions in the AppPAL assertion context.
-If all of an assertion's conditionals (the facts in the if part) are reachable (or it has none) then the speaker and predicate are added to the reachable set.
-If the statement is a can-say statement then we additionally check if the delegated predicate is reachable from the delegated speaker, and if so mark the delegated statement as reachable from the speaker who made the can-say statement.
-
-\begin{lstlisting}[language=Python,float,caption={Procedure for finding all reachable assertions.},label={alg:reachable}]
-def reachable(ac) -> set:
-  reachable = new set()
-  iterate = True
-  while iterate == True:
-    iterate = False
-    for assertion in ac:
-      e = a.speaker
-      p = a.predicate
-      if p.isCanSay() and (e, p) in reachable:
-        if (p.delegator, p.delegation) in reachable:
-          if for all c in a.conditions: (e, c.predicate) in reachable:
-            reachable.add((e, p.delegation))
-            iterate = True
-      else if not (e, p) in reachable:
-        if for all c in a.conditions; (e, c.predicate) in reachable:
-          reachable.add((e, p))
-          iterate = True
-  return reachable
-\end{lstlisting}
-
-\subsection{Checking Redundancy}
-
-To check redundancy we convert the policy into a graph structure.
-Each statement we might want to prove (a goal) becomes a node in the graph, its children are organised into sets of goals (a proof), where if all the goals in any set were proved true then the goal node would also become true.
-If a node has an empty set of goals to prove it is a fact.
-Once the graph has been constructed (taking into account delegation and unification with other statements), the graph is flattened by applying the flatten procedure (Listing~\ref{alg:flatten}) repeatedly until a fixed point is reached.
-When the graph has been flattened we look for redundancy by looking for goals with proofs that are subsets of their other proofs, implying that if the larger proof is true, then the shorter proof will always be true too;
-  and different goals with identical proofs, implying that the two decisions are not independent~(Listing~\ref{alg:redundancy}).
-
-\begin{lstlisting}[language=Python,float,caption={Procedure for flattening the redundancy graph.},label={alg:flatten}]
-def flatten(graph):
-  for goal in graph.goals:
-    hoist = true
-    for proof in graph[goal]:
-      for proof_goal in proof:
-        if not proof_goal.is_fact:
-          hoist = False
-          break
-    if hoist == True:
-      hoist(graph, goal)
-
-def hoist(graph, target):
-  for parent in graph[target].parents:
-    for proof in parent.proofs:
-      if proof.uses(target):
-        for replacements in graph[target].parents:
-          parent.proofs += proof.replace(target, replacements)
-        parent.proofs -= proof
-\end{lstlisting}
-
-\begin{lstlisting}[language=Python,float,caption={Procedure to check for redundancy.},label={alg:redundancy}]
-def check_redundancy(graph) -> boolean:
-  for goal in graph.goals:
-    for a in graph[goal]:
-      for b in graph[goal]:
-        if a >= b and a.goals.subset(b.goals):
-          if b.goals.subset(a.goals):
-            warn(a+" has multiple equivalent proofs") 
-          else
-            warn(a+" has a redundant proof")
-    for other_goal in graph.goals:
-      if goal > other_goal:
-        for a in graph[goal]:
-          for b in graph[other_goal]:
-            if not (a.goals.is_empty() or b.is_empty())
-               and a.goals == b.goals:
-              warn(a+" and "+b+" have equivalent proofs")
-\end{lstlisting}
 
 \section{Related Work}
+\label{sec:related}
 % 1-2 pages
-\label{sec:related}
-
-Armando~\etal~created a BYODroid~\cite{armando_bring_2013}; \ac{BYOD} framework
-to enforce \ac{BYOD} policies by extending Featherweight
-Java~\cite{igarashi_featherweight_2001} with a type system that can describe how
-Android apps interact. Whilst this work gives a framework for enforcing policies
-on Android, it doesn't help policy authors decide what the policies should be to
-begin with. They went on to show how BYODroid could be used to enforce a {NATO}
-Communications and Information Agency policy~\cite{armando_developing_2016}.
-Armando~\etal also proposed creating curated app stores based on corporate BYOD
-policies~\cite{armando_enabling_2014}. To do this they analysed a BYOD policy
-taken from the US Government BYOD Security
-Guidelines~\cite{digital_services_advisory_group_and_federal_chief_information_officers_council_bring_????}~and
-translated it into rules in the CONspec~\cite{aktug_conspec_2008} language which
-allows a low-level description of app behavior based on Java code. They then
-used these rules to filter apps for their store from a selection of 860 apps
-taken from the \emph{top selling free} section of the Google Play store.
-
-Martinelli~\etal~applied the UCON security model to describe \ac{BYOD} policies
-for Android~\cite{martinelli_enhancing_2016}. Using this model they created an
-enforcement framework that could alter the permissions granted to Android apps
-at run time. This framework requires a rooted device for most Android versions,
-however, they do claim it should work without modification on the latest
-version. Again like the work of Armando~\etal, this gives a framework for
-enforcement but without consideration of what the policies should be enforcing.
-
-Hong~\etal~focussedon the network aspects of a \ac{BYOD} policy with their
-PBS-Droid tool~\cite{sungmin_hong_towards_2016}.  PBS-Droid allowed BYOD policy
-authors the ability to sandbox individual apps into their own software-defined
-networks giving finer control over how individual apps interact with a companies
-networks and allowing for dynamic policy enforcement and network configuration.
-
-For more general mobile policies, there has been extensive work in both
-modeling and enforcing various policies.  Sadeh~\etal~surveyed users to find
-their privacy preferences and found four different types of user each with
-different privacy policies~\cite{sadeh_understanding_2009}.  Various tools have
-been proposed that can enforce app behavior policies.  Aurasium works by
-repackaging apps to use modified APIs~\cite{xu_aurasium:_2012}.  These APIs can
-enforce policies to protect a user's personal information or prevent sending
-premium rate text messages.  Another tool, CRePE~\cite{conti_crepe:_2010}
-modifies Android to provide a framework that can enforce policies based
-on the contextual information, such as the device's location or battery level.
-The SEAR tool takes a different approach and actively scans apps for
-vulnerabilities over collections of apps.  It then outputs security policies
-that users might want to enforce in order to prevent attacks from occurring in
-the first place~\cite{bagheri_practical_2016}.
 
 \section{Conclusions}
-% 0.5 pages
 \label{sec:conclusions}
+% 1/2 page
 

### 2016-10-07 paper.tex
did barely anything yesterday. shame

@@
 
+Corporate BYOD policies are tricky to fully enforce.
+Part of their policies are prescriptive:  if you configure your device in this way you will mitigate that threat.
+The policies contain more than just configuration, however.
+Consider this rule taken from the \emph{Security Policy for the use of handheld devices in corporate environments} by SANS~\cite{nicholas_r._c._guerin_security_2008}:
+
+\begin{quote}
+  Digital camera embedded on handheld devices \emph{might} be disabled in restricted environments, according to $\langle$COMPANY NAME$\rangle$ risk analysis. 
+  In sensitive facilities, information can be stolen using pictures and possibly sent using MMS or E-mail services.
+
+  In high-security facilities such as R&D labs or design manufacturers, camera MUST be disabled.
+  Furthermore, MMS messages should be disabled as well, to prevent malicious users from sending proprietary pictures.
+\end{quote}
+
 \section{The Details}

### 2016-10-12 paper.tex
updates

@@
 \usepackage{microtype}
+\usepackage{amsmath,amssymb,mathtools}
 
@@
 \acrodef{MDM}{Mobile Device Management}
+
+\usepackage{listings}
+
+\usepackage{framed}
+\newcommand{\todo}[1]{\marginpar{\begin{framed}\footnotesize\sffamily\raggedright #1\end{framed}}}
 
@@
 We have found, however, that it is a good fit for other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
+Using these papers we identify a set of standard decisions and trust relationships used in BYOD policies.
+These are BYOD idioms that are capture decisions at common requirements in BYOD policies.
+These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.
 
@@
 
+\newcommand{\textbra}[1]{\ensuremath{\left\langle \text{\sffamily #1} \right\rangle}}
 \begin{quote}
-  Digital camera embedded on handheld devices \emph{might} be disabled in restricted environments, according to $\langle$COMPANY NAME$\rangle$ risk analysis. 
+  Digital camera embedded on handheld devices \emph{might} be disabled in restricted environments, according to \textbra{COMPANY NAME} risk analysis. 
   In sensitive facilities, information can be stolen using pictures and possibly sent using MMS or E-mail services.
 
-  In high-security facilities such as R&D labs or design manufacturers, camera MUST be disabled.
+  In high-security facilities such as R\&D labs or design manufacturers, camera MUST be disabled.
   Furthermore, MMS messages should be disabled as well, to prevent malicious users from sending proprietary pictures.
 \end{quote}
+
+To implement this policy an \ac{MDM} program, such as IBM's MaaS360 could be used~\cite{_ibm_????}.
+This software supports app wrapping (where an app is recompiled to use guarded APIs), and supports security ``checkboxes'' to enable security controls and access control policies.
+Using techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration to implement.  
+It talks of \emph{restricted environments} decided by \emph{company risk analysis}; but how is this communicated to the device?
+Does it access the list of restricted environments once, from a server?
+Does it understand how the risk analysis was performed?  
+Can it decide on the basis of a policy if a location is restricted?
+The rule also gives a security objective: \emph{prevent malicious users from sending proprietary pictures}.
+The guidelines are given, however, for the specific cases of a compliant user using MMS or email.
+This may not be sufficient to stop a sufficiently motivated \emph{malicious} user.
+
+Our approach does not try to implement the policy.
+Rather we attempt to capture the goals of the policy rule so that the delegations of trust, tools used, and their configuration is made explicit.
+This gives us greater clarity as to which tool is being trusted to implement what policy.
+It allows us to see who is being trusted to make which decisions.
+It also lets us use automatic tools to uncover aspects of the policy that may be unimplemented~\cite{hallett_specifying_2016}.
+Continuing with the example above, this can be expressed in AppPAL as:
+
+\begin{lstlisting}
+'company' says 'risk-analysis' can-say 
+  Location:L isHighSecurityFacility.
+
+'company' says Device:D mustDisableIn(Location, 'camera')
+  if Location isHighSecurityFacility.
+
+'company' says Device:D mustDisableIn(Location, 'mms') 
+  if Location isHighSecurityFacility.
+
+'company' says User:U hasSatisfied('proprietary pictures policy')
+  if U hasDevice(D),
+     D mustDisableIn(Location, 'camera'),
+     D mustDisableIn(Location, 'mms'),
+     Location isHighSecurityFacility.
+\end{lstlisting}
+
+\todo{%
+  Why is the apppal better?
+  Clearer, explicit, can identify problems automatically.
+
+  Explain why I'm trying to build a schema for byod policies and why this is worth a paper.
+
+  How is this falsifiable?
+  Examples of policy idioms we can't express.
+}
 
@@
 % 5 pages
+
+\subsection{BYOD Policies}
+Policies were taken from 5 sources:
+\begin{itemize}
+  \item SANS
+  \item NHS
+  \item HiMMS
+  \item Edinburgh
+  \item Code3SPE
+\end{itemize}
+\todo{Describe differences and similaritites at a high level between these policies.}
+
+\subsection{BYOD Idioms}
+\todo{Describe common idioms and patterns found in each of the policies.  Check the earlier draft for some starting work}
 

### 2016-10-13 paper.tex
updates

@@
 \usepackage{listings}
+\lstdefinelanguage{AppPAL}{%
+  morekeywords={if,inf,says,where,true,false,with,is},
+  otherkeywords={can-act-as,can-say},
+  sensitive=true,
+  morestring=[b]',
+  literate={\ inf\ }{{$\infty$}}5
+}[keywords,strings]
+\lstset{%
+  basicstyle=\ttfamily,
+  keywordstyle=\ttfamily\slshape,
+  stringstyle=\sffamily,
+  columns=flexible,
+  mathescape,
+  tabsize=2,
+  showstringspaces=true,
+  %numbers=left,
+  breaklines=true,
+  numberstyle=\tiny\ttfamily\color{Gray},
+  commentstyle=\ttfamily\color{Gray}
+}
+\newcommand{\code}[2][]{\lstinline[breaklines=true, #1]{#2}}
 
@@
 \subsection{BYOD Policies}
-Policies were taken from 5 sources:
+
+We analysed five different policies to find common idioms.
+The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
+This is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices. 
+Companies are expected to modify the document to suit their needs.  The policy is general; not specific to any particular industry, device, or country's legislation.
+
+The second is taken from HiMMS.
+
 \begin{itemize}

### 2016-10-21 paper.tex
updates

@@
 Companies are expected to modify the document to suit their needs.  The policy is general; not specific to any particular industry, device, or country's legislation.
-
-The second is taken from HiMMS.
+The second is taken from Healthcare Information Management System Society~\cite{healthcare_information_and_management_systems_society_mobile_2012}; a US non-profit company trying to improve healthcare through IT.
+The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
+The third is taken from an british hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital. 
+Finally two simpler policies are taken from Edinburgh university~\cite{david_williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}. 
+These policies are simpler than the other comprised of much more general rules and only a page or two in length.
 

### 2016-10-25 paper.tex
updates

@@
 \usepackage{amsmath,amssymb,mathtools}
+\usepackage{hyperref}
+\usepackage{booktabs}
 
@@
   %numbers=left,
+  frame=single,
   breaklines=true,
@@
 \newcommand{\todo}[1]{\marginpar{\begin{framed}\footnotesize\sffamily\raggedright #1\end{framed}}}
+
+\newenvironment{policyrule}[1]{%
+  \begin{framed}
+      \noindent\textbf{\sffamily #1}:~\itshape
+}{
+  \end{framed}
+}
 
@@
 
+\subsection{Changes to AppPAL}
+
+We have used AppPAL to formalise the BYOD policies.
+AppPAL is an instantiation of Becker's SecPAL~\cite{becker_secpal:_2010} we have previously used to describe app privacy preferences~\cite{hallett_apppal_2016}.
+To describe BYOD policies we have extended AppPAL with types.
+This allows our policies to be more succinct and readable whilst losing none of the expressive power.
+When describing variables in the head of an AppPAL assertion.
+
 \subsection{BYOD Policies}
@@
 This is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices. 
-Companies are expected to modify the document to suit their needs.  The policy is general; not specific to any particular industry, device, or country's legislation.
-The second is taken from Healthcare Information Management System Society~\cite{healthcare_information_and_management_systems_society_mobile_2012}; a US non-profit company trying to improve healthcare through IT.
+Companies are expected to modify the document to suit their needs.  
+The policy is general; not specific to any particular industry, device, or country's legislation.
+The second is taken from Healthcare Information Management System Society~\cite{healthcare_information_and_management_systems_society_mobile_2012}; 
+  a US non-profit company trying to improve healthcare through IT.  
+It is relatively short but interesting because it contains concerns specific to healthcare scenarios, and is written as a contract the users are agreeing to follow. 
+In contrast every other policy we looked at is written as a series of rules users should follow to ensure compliance.
 The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
@@
 
-\begin{itemize}
-  \item SANS
-  \item NHS
-  \item HiMMS
-  \item Edinburgh
-  \item Code3SPE
-\end{itemize}
-\todo{Describe differences and similaritites at a high level between these policies.}
+\begin{table}\centering\footnotesize\sffamily
+  \begin{tabular}{l c c c c c}
+    \toprule
+                             & {SANS}       & {HiMSS}     & {NHS}       & {Edinburgh} & {Sirens}    \\
+    \midrule
+    Number of rules          & 33           & 15          & 56          &             & 25          \\
+    SecPAL assertions        & 71           & 21          & 58          & 10          & 39          \\
+    Policy coverage          & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) &             & 22 (88$\%$) \\
+    \midrule
+    Delegations              & 23           & 5           & 33          & 2           & 13          \\
+    Acknowledgement          &              &             &             &             &             \\
+    Partial Acknowledgement  &              &             &             &             &             \\
+    \bottomrule             \\
+  \end{tabular}
+  \label{tab:summary}
+  \caption{Summary of the contents of each of the BYOD policies.}
+\end{table}
+
+The policies are summarised in \autoref{tab:summary}.
+Each policy contains a series of \emph{rules}, which are implemented by one or more \emph{AppPAL assertions}.
+The \emph{policy coverage} represents the number of rules that have an AppPAL description attached.
+Not all the rules in a policy require an AppPAL translation.
+For example rule 9.1 in the NHS policy states:
+\begin{policyrule}{NHS}
+  Downloading of personal apps onto a corporately issued mobile device should be avoided where possible. 
+  The Trust would not encourage staff members to download apps for personal use onto a corporately issued mobile device.
+  All staff are reminded that they must adhere to the guidance outlined in the Social Media Policy.
+\end{policyrule}
+The rule does not prohibit any action, and doesn't require the staff to do anything different.
+It does remind staff to be aware of a rule which we could implement with an acknowledegement; a statement that the party has acknowledged a rule.
+\begin{lstlisting}
+'nhs-trust' says Staff:S mustAcknowledge('avoid-installing-personal-apps').
+'nhs-trust' says Staff:S mustAcknowledge('social-media-policy').
+\end{lstlisting}
+In this case, however, since nothing is required even an acknowledgement seems excessive.
 

### 2016-10-25 paper.tex
describes types

@@
 
-We have used AppPAL to formalise the BYOD policies.
-AppPAL is an instantiation of Becker's SecPAL~\cite{becker_secpal:_2010} we have previously used to describe app privacy preferences~\cite{hallett_apppal_2016}.
-To describe BYOD policies we have extended AppPAL with types.
-This allows our policies to be more succinct and readable whilst losing none of the expressive power.
-When describing variables in the head of an AppPAL assertion.
+When using AppPAL to describe policies, we need to describe what domain each variable belongs to.
+AppPAL inherits from SecPAL (and Datalog) a safety condition that all variables in the head of an assertion must be referenced in the body.
+This leads to policies which can be noisy to read.
+For example in the SANS policy there is a rule that states that devices can only connect to wireless access points owned by the company.
+This is written in AppPAL as:
+\begin{lstlisting}
+'company' says Device canConnectToAP(AP)
+  if AP isOwnedByCompany,
+     Device isDevice,
+     AP isAccessPoint.
+\end{lstlisting}
+To simplify the presentation we allow a special sugared notation for conditionals expressing the type of a variable (\emph{variable \emph{is}Type}).
+Any variable in the head of the rule of the form \texttt{\emph{Type}:Variable} is replaced by the variable and a conditional that \textt{Variable is\emph{Type}} is added to the body.
+The rule can be re-written as shown bellow hiding the typing statements.
+\begin{lstlisting}
+'company' says Device:D canConnectToAP(AP:X)
+  if X isOwnedByCompany.
+\end{lstlisting}
 

### 2016-10-25 paper.tex
wrote section on changes to apppal

@@
 }[keywords,strings]
+\newcommand{\listingsize}[0]{\footnotesize}
 \lstset{%
-  basicstyle=\ttfamily,
-  keywordstyle=\ttfamily\slshape,
-  stringstyle=\sffamily,
+  basicstyle=\ttfamily\listingsize,
+  keywordstyle=\ttfamily\slshape\listingsize,
+  stringstyle=\sffamily\listingsize,
   columns=flexible,
@@
 
-When using AppPAL to describe policies, we need to describe what domain each variable belongs to.
+In order to describe BYOD specific conditions we instantiate AppPAL with four kinds of facts: \emph{can, has, is} and \emph{must}.
+Like other SecPAL-based instantiations~\cite{becker_framework_2009,aziz_secpal4dsa:_2011} we extend the syntax of facts to support these constructs.
+Internally, however, they are treated as standard predicates.
+
+\begin{center}
+  \footnotesize\sffamily
+  \newcommand{\predicate}[3]{#1 \texttt{#2\textit{#3}}}
+  \begin{tabular}{l l}
+    \toprule
+    Fact                              & Meaning                                         \\
+    \midrule
+    \predicate{subject}{can}{Action}  & The subject is permitted to perform the action. \\
+    \predicate{subject}{has}{Action}  & The subject has performed the action.           \\
+    \predicate{subject}{is}{Type}     & The subject is a member of the type.            \\
+    \predicate{subject}{must}{Action} & The subject must perform the action.            \\
+    \bottomrule
+  \end{tabular}
+\end{center}
+
+Facts of the \emph{must}-kind represent obligations, actions that should be completed if a particular scenario presents itself.
+For these facts we add a rule to ensure the obligation is performed.
+This rule should be checked periodically to ensure compliance.
+\begin{lstlisting}
+$\langle\text{speaker}\rangle$ says $\langle\text{subject}\rangle$ hasSatisfiedObligation$\langle\text{Action}\rangle$
+  if $\langle\text{subject}\rangle$ must$\langle\text{Action}\rangle$,
+     $\langle\text{subject}\rangle$ has$\langle\text{Action}\rangle$.
+\end{lstlisting}
+
+When using AppPAL to describe policies, we also need to describe what domain each variable belongs to.
+This is performed through the \emph{is}-kind facts.
 AppPAL inherits from SecPAL (and Datalog) a safety condition that all variables in the head of an assertion must be referenced in the body.
@@
 To simplify the presentation we allow a special sugared notation for conditionals expressing the type of a variable (\emph{variable \emph{is}Type}).
-Any variable in the head of the rule of the form \texttt{\emph{Type}:Variable} is replaced by the variable and a conditional that \textt{Variable is\emph{Type}} is added to the body.
+Any variable in the head of the rule of the form \texttt{\emph{Type}:Variable} is replaced by the variable and a conditional that \texttt{Variable is\emph{Type}} is added to the body.
 The rule can be re-written as shown bellow hiding the typing statements.

### 2016-10-25 paper.tex
Added a bit on app selection

@@
 \usepackage{booktabs}
+
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
 
@@
 \newenvironment{policyrule}[1]{%
-  \begin{framed}
-      \noindent\textbf{\sffamily #1}:~\itshape
-}{
+  \begin{framed}\footnotesize
+      \noindent\textbf{\sffamily #1}:~\itshape%
+}{%
   \end{framed}
@@
 The \emph{policy coverage} represents the number of rules that have an AppPAL description attached.
+
 Not all the rules in a policy require an AppPAL translation.
@@
 
+Table~\ref{tab:byod-predicates} summarises the facts used by multiple policies.
+This shows what concerns and controls and are shared between different BYOD policies.
+All five of the policies make use of acknowledgements, for example.
+This could be because not all the rules in a policy can be enforced through technical means, 
+  or it could indicate policy authors care more that the subjects are aware of the rules than they do for rigorous enforcement.
+All but the HiMMS policy have rules that include locking down a device by disabling features.
+All but the Edinburgh policy have rules that look at what should happen when a device is lost,
+  and have rules that require employees inform someone when something happens.
+These common concerns suggest where future \ac{MDM} software should focus their efforts.
+
+Only the NHS and SANS, the two most complex policies, describe when a device can install an app and what kinds of apps are installable.
+In both policies this is predominantly limited to a delegation to the appropriate groups to authorize an app.
+For example, in the SANS policy the IT-Department are responsible for deciding what apps can be installed.
+\begin{policyrule}{SANS}
+  The IT Department maintains lists of allowed and unauthorized applications and makes them available to users on the intranet.
+  \normalfont
+  \begin{lstlisting}
+'company' says 'it-department' can-say App:A isInstallable.
+  \end{lstlisting}
+\end{policyrule}
+The NHS policy, however, is significantly more complicated.
+Apps have to be approved by three different groups (the Integrated Governance Comitee, the Employee's manager, and the relevant group within the trust depending on where the employee works) before the trust will say that an app can be installed.
+\begin{policyrule}{NHS}
+  Apps for work usage must not be downloaded onto corporately issued
+  mobile devices (even if approved on the NHS apps store) unless they have
+  been approved through the following Trust channels:
+
+  Clinical apps; at the time of writing there are no apps clinically
+  approved by the Trust for use with patients~/~clients. However, if a
+  member of staff believes that there are clinical apps or other
+  technologies that could benefit their patients~/~clients, this should be
+  discussed with the clinical lead in the first instance and ratification
+  should be sought via the Care and Clinical Policies Group. A clinical
+  app should not be used if it has not been approved via this group.
+
+  Business apps; at the time of writing there are no business (i.e.,
+  non-clinical) apps approved by the Trust for use other than those
+  preloaded onto the device at the point of issue. However, if a member of
+  staff believes that there are apps or other technologies that could
+  benefit their non-clinical work, ratification of the app must be sought
+  via the Management of Information Group (MIG). An app should not be used
+  if it has not been approved via this group.
+
+  Following approval through Care and Clinical Policies and~/~or MIG, final
+  approval will be required through Integrated Governance Committee.
+
+  Use of paid apps must be agreed in advance with the device holder's line
+  manager and there should be a demonstrable benefit.
+  \normalfont
+  \begin{lstlisting}
+'nhs-trust' says App isInstallable if App hasMet('clinical-use-case').
+'nhs-trust' says App isInstallable if App hasMet('business-use-case').
+'nhs-trust' says 'cacpg' can-say App:A hasMet('clinical-use-case').
+'nhs-trust' says 'mig' can-say App:A hasMet('business-use-case').
+'nhs-trust' says App isInstallable
+  if App hasMet('final-app-approval').
+'nhs-trust' says 'igc' can-say App hasMet('final-app-approval').
+'nhs-trust' says Device canInstall(App)
+  if App isInstallable,
+     App isApprovedFor(Device).
+'nhs-trust' says Employee:Manager can-say
+  App:A isApprovedFor(Device)
+  if Manager isResponsibleFor(Device).
+  \end{lstlisting}
+\end{policyrule}
+Given that user's have privacy preferences expressable as permissions~\cite{lin_modeling_2014}, we might expect corporate policies to also express preferences about what permissions an app can have.
+This does not appear to be the case, however. 
+As part of selecting the apps an IT department or group may chose to use advanced instrumentation and policies~\cite{armando_enabling_2014}, or they may manually chose apps to form a curated app store as some \ac{MDM} vendors allow.
+But from the perspective of the policy itself it is more important \emph{who} makes the decision rather than \emph{what} they chose, if they have an opinion at all.
+
+\begin{table}\centering\footnotesize
+  \newcommand{\angledtitle}[1]{\rlap{\rotatebox{30}{#1}}}
+  \sffamily
+  \begin{tabular}{l c c c c c p{0.45\linewidth} }
+    \toprule
+    Predicate                                   & \angledtitle{Sirens}   & \angledtitle{Edinburgh} & \angledtitle{HiMMS} & \angledtitle{NHS} & \angledtitle{SANS}
+    & Description                                                                          \\
+    \midrule
+    Device \texttt{canBackupTo}(Server)         &                        & \cmark                  & \cmark              &                   &
+    & Says the device may send backups to a server.                                        \\
+    Device \texttt{canCall}(Number)             &                        &                         &                     & \cmark            & \cmark
+    & Says the telephone numbers a device can call.                                        \\
+    Device \texttt{canConnectToAP}(AP)          &                        &                         & \cmark              &                   & \cmark
+    & Says a device may associate with an access point.                                    \\
+    Device \texttt{canConnectToNetwork}(Subnet) & \cmark                 &                         &                     &                   & \cmark
+    & Says the device may connect to a network, for example computers all within a subnet. \\
+    Device \texttt{canConnectToServer}(IP)      &                        &                         &                     & \cmark            & \cmark
+    & Says the subject (a device) can connect to a given server (identitfied by a URL).    \\
+    Device \texttt{canInstall}(App)             &                        &                         &                     & \cmark            & \cmark
+    & Says the device may install an app.                                                  \\
+    User \texttt{canMonitor}(Device)            & \cmark                 &                         &                     & \cmark            & \cmark
+    & Says the subject (a user) can monitor and unlock a device.                           \\
+    Device \texttt{canStore}(File)              &                        &                         & \cmark              &                   & \cmark
+    & Says a device may store some data in permanent (non-transient) storage.              \\
+    User \texttt{canUse}(Device)                &                        &                         &                     & \cmark            & \cmark
+    & Says a user may use a device.                                                        \\
+    User \texttt{hasAcknowledged}(Policy)       & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
+    & Says an individual has acknoweldged a policy.                                        \\
+    Device \texttt{hasFeature}(Feature)         &                        & \cmark                  &                     & \cmark            &
+    & Says a device has a feature enabled.                                                 \\
+    Something \texttt{hasMet}(Policy)           &                        & \cmark                  &                     & \cmark            & \cmark
+    & Says a something has met a given set of requirements.                                \\
+    Device \texttt{isActivated}                 &                        &                         & \cmark              & \cmark            & \cmark
+    & Specifies a device has been activated for BYOD usage.                                \\
+    App \texttt{isApp}                          & \cmark                 &                         & \cmark              & \cmark            & \cmark
+    & Specifies an app.                                                                    \\
+    Something \texttt{isApproved}               &                        &                         & \cmark              &                   & \cmark
+    & Specifies that something has been vetted and approved of.                            \\
+    Data \texttt{isData}                        & \cmark                 &                         &                     &                   & \cmark
+    & Specifies data.                                                                      \\
+    Device \texttt{isDevice}                    & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
+    & Specifies a device.                                                                  \\
+    User \texttt{isEmployee}                    & \cmark                 & \cmark                  &                     & \cmark            & \cmark
+    & Specifies that someone is an employee.                                               \\
+    Device \texttt{isEncrypted}                 &                        & \cmark                  &                     & \cmark            & \cmark
+    & Specifies a device is encrypted.                                                     \\
+    Feature \texttt{isFeature}                  & \cmark                 &                         &                     &                   & \cmark
+    & Specifies a feature.                                                                 \\
+    App \texttt{isInstallable}                  &                        &                         &                     & \cmark            & \cmark
+    & Specifies an app is installable.                                                     \\
+    Device \texttt{isLost}                      & \cmark                 &                         & \cmark              & \cmark            & \cmark
+    & Specifies a device is missing.                                                       \\
+    Device \texttt{isOwnedBy}(User)             & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
+    & Specifies something's owner.                                                         \\
+    File \texttt{isSecurityLevel}(Level)        &                        &                         & \cmark              &                   & \cmark
+    & Specifies some data as having business sensitive information.                        \\
+    Something \texttt{isString}                 & \cmark                 &                         &                     & \cmark            &
+    & Specifies a string.                                                                  \\
+    Number \texttt{isTelephoneNumber}           &                        &                         &                     & \cmark            & \cmark
+    & Specifies a telephone number.                                                        \\
+    Device \texttt{mustDisable}(Feature)        & \cmark                 & \cmark                  &                     & \cmark            & \cmark
+    & Forces the disablement of a feature (or a device).                                   \\
+    User \texttt{mustInform}(User, Something)   & \cmark                 &                         & \cmark              & \cmark            & \cmark
+    & Forces someone to disclose something to someone.                                     \\
+    Device \texttt{mustWipe}                    & \cmark                 &                         & \cmark              & \cmark            &
+    & Forces a device to be erased.                                                        \\
+    \bottomrule
+  \end{tabular}
+  \caption{Summary of AppPAL facts used in multiple BYOD policies.}
+  \label{tab:byod-predicates}
+\end{table}
+
 \subsection{BYOD Idioms}

### 2016-10-25 paper.tex
spelling

@@
 The use of \ac{MDM} software does not guarantee compliance however.
-One survey found over 50\% of companies with \ac{MDM} software still have uncompliant devices in their networks~\cite{mobileiron_security_labs_q4_2015}.
+One survey found over 50\% of companies with \ac{MDM} software still have un-compliant devices in their networks~\cite{mobileiron_security_labs_q4_2015}.
 
@@
 \todo{%
-  Why is the apppal better?
+  Why is the AppPAL better?
   Clearer, explicit, can identify problems automatically.
 
-  Explain why I'm trying to build a schema for byod policies and why this is worth a paper.
+  Explain why I'm trying to build a schema for BYOD policies and why this is worth a paper.
 
@@
 The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
-The third is taken from an british hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital.
+The third is taken from an British hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital.
 Finally two simpler policies are taken from Edinburgh university~\cite{david_williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
@@
 The rule does not prohibit any action, and doesn't require the staff to do anything different.
-It does remind staff to be aware of a rule which we could implement with an acknowledegement; a statement that the party has acknowledged a rule.
+It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
 \begin{lstlisting}
@@
 The NHS policy, however, is significantly more complicated.
-Apps have to be approved by three different groups (the Integrated Governance Comitee, the Employee's manager, and the relevant group within the trust depending on where the employee works) before the trust will say that an app can be installed.
+Apps have to be approved by three different groups (the Integrated Governance Committee, the Employee's manager, and the relevant group within the trust depending on where the employee works) before the trust will say that an app can be installed.
 \begin{policyrule}{NHS}
@@
 \end{policyrule}
-Given that user's have privacy preferences expressable as permissions~\cite{lin_modeling_2014}, we might expect corporate policies to also express preferences about what permissions an app can have.
+Given that user's have privacy preferences expressible as permissions~\cite{lin_modeling_2014}, we might expect corporate policies to also express preferences about what permissions an app can have.
 This does not appear to be the case, however. 
@@
     Device \texttt{canConnectToServer}(IP)      &                        &                         &                     & \cmark            & \cmark
-    & Says the subject (a device) can connect to a given server (identitfied by a URL).    \\
+    & Says the subject (a device) can connect to a given server (identified by a URL).    \\
     Device \texttt{canInstall}(App)             &                        &                         &                     & \cmark            & \cmark
@@
     User \texttt{hasAcknowledged}(Policy)       & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
-    & Says an individual has acknoweldged a policy.                                        \\
+    & Says an individual has acknowledged a policy.                                        \\
     Device \texttt{hasFeature}(Feature)         &                        & \cmark                  &                     & \cmark            &

### 2016-11-08 paper.tex
work work work

@@
-\documentclass{article}
+\documentclass{easychair}
 
@@
 \newcommand{\xmark}{\ding{55}}%
+
+\newcommand{\etal}{et~al.\@}
 
@@
 \lstset{%
-  basicstyle=\ttfamily\listingsize,
-  keywordstyle=\ttfamily\slshape\listingsize,
-  stringstyle=\sffamily\listingsize,
+  basicstyle=\normalfont\ttfamily\listingsize,
+  keywordstyle=\normalfont\ttfamily\slshape\listingsize,
+  stringstyle=\normalfont\sffamily\listingsize,
   columns=flexible,
@@
 
-\usepackage{framed}
-\newcommand{\todo}[1]{\marginpar{\begin{framed}\footnotesize\sffamily\raggedright #1\end{framed}}}
+\usepackage{mdframed}
+\newcommand{\todo}[1]{\marginpar{\begin{mdframed}[userdefinedwidth=1cm]\tiny\sffamily\raggedright #1\end{mdframed}}}
 
 \newenvironment{policyrule}[1]{%
-  \begin{framed}\footnotesize
+  \begin{mdframed}\footnotesize
       \noindent\textbf{\sffamily #1}:~\itshape%
 }{%
-  \end{framed}
+  \end{mdframed}
 }
@@
 \author{Joseph Hallett and David Aspinall}
+\institute{University of Edinburgh}
 \begin{document}
@@
   % 4 Sentences
+  BYOD policies are informally specified using natural language.
+  Various tools and \ac{MDM} solutions can implement parts of the policies, but there isn't a means to tie the written policy to its enforcement method.
+  In this paper we show how AppPAL can be used to capture policies for BYOD.
+  We identify common BYOD idioms and delegation structures.
 \end{abstract}
@@
 \newcommand{\textbra}[1]{\ensuremath{\left\langle \text{\sffamily #1} \right\rangle}}
-\begin{quote}
+\begin{policyrule}{SANS}
   Digital camera embedded on handheld devices \emph{might} be disabled in restricted environments, according to \textbra{COMPANY NAME} risk analysis.
@@
   Furthermore, MMS messages should be disabled as well, to prevent malicious users from sending proprietary pictures.
-\end{quote}
+\end{policyrule}
 
@@
     & Specifies a telephone number.                                                        \\
+    User \texttt{mustAcknowledged}(Policy)      & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
+    & Requires the user to acknowledge a  policy.                                          \\
     Device \texttt{mustDisable}(Feature)        & \cmark                 & \cmark                  &                     & \cmark            & \cmark
@@
 
+\subsubsection{Acknowledgement}
+
+All of the policies looked at require their subjects to be aware and acknowledge certain rules or policies, 
+  or that the company may perform certain actions.
+For example the NHS and HiMSS policies state that the company will wipe devices remotely to protect confidential information if the device is lost.
+Alongside these rules both policies also state that if an employee has personal information on the device then this will be lost also.
+The employee is required to be aware of this, and in the case of the HiMSS policy, agree to hold the company harmless for the loss.
+
+\begin{center}
+  \noindent
+  %\begin{minipage}{0.49\linewidth}
+    \begin{policyrule}{HiMSS}
+      I agree to hold XYZ Health System harmless for any loss relating to the
+      administration of PDA/Smartphone connectivity to XYZ Health System systems
+      including, but not limited to, loss of personal information stored on a
+      PDA/Smartphone due to data deletion done to protect sensitive information
+      related to XYZ Health System, its patients, members or partners.
+      \normalfont
+      \begin{lstlisting}
+'xyz-health-system' says 'user' mustAcknowledged('data-loss-policy').
+      \end{lstlisting}
+    \end{policyrule}
+  %\end{minipage}
+  %\begin{minipage}{0.49\linewidth}
+    \begin{policyrule}{NHS}
+      Individuals who have personal data of any kind stored on a corporately
+      issued mobile device must be aware that in the event of loss of the device
+      the above data wipe will include removal of all personal data.
+      \normalfont
+      \begin{lstlisting}
+'nhs-trust' says Staff:S can-say
+  S hasAcknowledged('data-loss-policy').
+'nhs-trust' says Staff:S mustAcknowledged('data-loss-policy').
+      \end{lstlisting}
+    \end{policyrule}   
+  %\end{minipage}
+\end{center}
+
+Both the SANS and the siren-company policies use acknowledgements to link to other sets of rules that employees should be aware of.
+These policies aren't further specified, and in the case of an acceptable use policy may be hard to enforce automatically.
+
+\begin{policyrule}{SANS}
+  Users MUST agree to the email security/acceptable use policy and eventually to the eCommerce security policy.
+  \begin{lstlisting}
+'company' says Employee:U mustAcknowledged('email-security'). 
+'company' says Employee:U mustAcknowledged('acceptable-use'). 
+'company' says Employee:U mustAcknowledged('ecommerce-security').
+  \end{lstlisting}
+\end{policyrule}
+\begin{policyrule}{Sirens}
+  The employee is expected to use his or her devices in an ethical manner at all times and adhere to the company's acceptable use policy.
+  \begin{lstlisting}
+'department' says Employee:E mustAcknowledged('acceptable-use').
+  \end{lstlisting}
+\end{policyrule}
+
+\todo{Provide a template for what kinds of policy should be implemented like this and template AppPAL.}
+
+\subsection{Delegation and Roles within Policies}
+
+AppPAL requires all assertions be made by an explicit speaker.
+When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:principals}),
+  with the exception of the HiMSS policy that is written as a contract the user is agreeing to.
+All the policies delegate also describe other entities that might also make assertions.
+With AppPAL policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
+This might be to a user to acknowledge a policy, or it might be to other groups in the company who are responsible for certain decisions.
+
+In all of the policies we looked at the majority of the decisions are made by three speakers: 
+  the company, the IT-department, and the users or employees.
+All of the policies also delegate to a user (apart from HiMSS where the user is the primary speaker).
+The user is typically responsible for providing information, such as agreements to policies, reporting devices missing, and updating passcodes.
+
+The Sirens, SANS and NHS policy each describe an IT-department who are delegated to make some decisions.
+The HiMSS policy also describes an \emph{xyz-health-system}, who act similarly to an IT-department.
+These decisions are more varied and can overlap with the responsibilities of the company.
+In the NHS and SANS policies the IT department is responsible for maintaining lists of activated devices.
+In the Sirens and SANS policies the IT department maintains a list of what is installable on a device or not.
+
+\begin{table}\centering\footnotesize
+  \begin{tabular}{l l}
+    \toprule
+    Policy      & Primary Speaker           \\
+    \midrule
+    {SANS}      & \emph{company}            \\
+    {HiMSS}     & \emph{user}               \\
+    {NHS}       & \emph{nhs-trust}          \\
+    {Edinburgh} & \emph{records-management} \\
+    {Sirens}    & \emph{department}         \\
+    \bottomrule
+  \end{tabular}
+  \label{tab:principals}
+  \caption{Principal Speakers from each of the policies examined.}
+\end{table}
+
 \section{Related Work}
@@
 % 1-2 pages
+
+Companies have very little visibility into how their policies are implemented.
+Morrow notes that \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it it's current''}~\cite{morrow_byod_2012}.
+Even when devices have means to implement policies correctly it is difficult to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
+Our work aims to address these problem directly: by using formal languages we can link the policy to the implementation.
+
+Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_enabling_2014}.
+Their tool allows companies to distribute a secure app~store where through a combination of static and dynamic tools they can verify that apps follow fine-grained permission policies.
+Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
+Using their tool they implemented parts of a policy relating to personal networks, and data management for the NATO Communications and Information Agency~\cite{armando_developing_2016}.
+Their work shows how parts of a BYOD policy can be checked for and enforced using tools.
+An AppPAL policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
+AppPAL makes explicit how the natural language policy is implemented and by which tools.
+
+Costantino~\etal~created a role-based access control scheme for BYOD devices that allowed a company to verify the devices within their network and what servers and services they had access to~\cite{costantino_towards_2013}.
+The scheme relied on devices having a TPM however to ensure policy violations were reported.
+
+Other work has looked at using app wrapping to enforce more general policies.
+Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
+This has the advantage that the device's underlying OS needn't be modified as the apps can be modified at the source code level.
+Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}, however app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.
 

### 2016-11-17 paper.tex
switched to llncs and some trivial things

@@
-\documentclass{easychair}
+\documentclass{llncs}
 
@@
 \newcommand{\etal}{et~al.\@}
+
+\newcommand{\AppPAL}[0]{SecPAL4BYOD}
 
@@
   Various tools and \ac{MDM} solutions can implement parts of the policies, but there isn't a means to tie the written policy to its enforcement method.
-  In this paper we show how AppPAL can be used to capture policies for BYOD.
+  In this paper we show how \AppPAL~can be used to capture policies for BYOD.
   We identify common BYOD idioms and delegation structures.
@@
 
-This paper describes a formalization of five BYOD policies using AppPAL---a policy language for mobile device privacy preferences~\cite{hallett_apppal_2016}.
+This paper describes a formalization of five BYOD policies using a variation of SecPAL called \AppPAL.
+This has been implemented atop of AppPAL---a policy language for mobile device privacy preferences~\cite{hallett_apppal_2016}.
 Using our formalization we identify the common concerns and trust relationships in these policies.
-We show how AppPAL can be used to implement these policies, and describe precisely the different trust relationships.
-AppPAL is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies.
+We show how \AppPAL\AppPALcan be used to implement these policies, and describe precisely the different trust relationships.
+\AppPAL\AppPALis an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies.
 We have found, however, that it is a good fit for other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
@@
 It also lets us use automatic tools to uncover aspects of the policy that may be unimplemented~\cite{hallett_specifying_2016}.
-Continuing with the example above, this can be expressed in AppPAL as:
+Continuing with the example above, this can be expressed in \AppPAL~as:
 
@@
 \todo{%
-  Why is the AppPAL better?
+  Why is the \AppPAL~better?
   Clearer, explicit, can identify problems automatically.
@@
 
-\subsection{Changes to AppPAL}
+\subsection{Changes to \AppPAL}
 
-In order to describe BYOD specific conditions we instantiate AppPAL with four kinds of facts: \emph{can, has, is} and \emph{must}.
+In order to describe BYOD specific conditions we instantiate \AppPAL~with four kinds of facts: \emph{can, has, is} and \emph{must}.
 Like other SecPAL-based instantiations~\cite{becker_framework_2009,aziz_secpal4dsa:_2011} we extend the syntax of facts to support these constructs.
@@
 
-When using AppPAL to describe policies, we also need to describe what domain each variable belongs to.
+When using \AppPAL~to describe policies, we also need to describe what domain each variable belongs to.
 This is performed through the \emph{is}-kind facts.
-AppPAL inherits from SecPAL (and Datalog) a safety condition that all variables in the head of an assertion must be referenced in the body.
+\AppPAL~inherits from SecPAL (and Datalog) a safety condition that all variables in the head of an assertion must be referenced in the body.
 This leads to policies which can be noisy to read.
 For example in the SANS policy there is a rule that states that devices can only connect to wireless access points owned by the company.
-This is written in AppPAL as:
+This is written in \AppPAL~as:
 \begin{lstlisting}
@@
 The third is taken from an British hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital.
-Finally two simpler policies are taken from Edinburgh university~\cite{david_williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
+Finally two simpler policies are taken from The University of Edinburgh~\cite{david_williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
 These policies are simpler than the other comprised of much more general rules and only a page or two in length.
@@
 The policies are summarised in \autoref{tab:summary}.
-Each policy contains a series of \emph{rules}, which are implemented by one or more \emph{AppPAL assertions}.
-The \emph{policy coverage} represents the number of rules that have an AppPAL description attached.
+Each policy contains a series of \emph{rules}, which are implemented by one or more \emph{\AppPAL~assertions}.
+The \emph{policy coverage} represents the number of rules that have a \AppPAL~description attached.
 
-Not all the rules in a policy require an AppPAL translation.
+Not all the rules in a policy require a \AppPAL~translation.
 For example rule 9.1 in the NHS policy states:
@@
 \begin{table}\centering\footnotesize
-  \newcommand{\angledtitle}[1]{\rlap{\rotatebox{30}{#1}}}
+  \newcommand{\angledtitle}[1]{\rlap{\rotatebox{45}{#1}}\hspace{1em}}
   \sffamily
@@
   \end{tabular}
-  \caption{Summary of AppPAL facts used in multiple BYOD policies.}
+  \caption{Summary of \AppPAL~facts used in multiple BYOD policies.}
   \label{tab:byod-predicates}
@@
 
-\todo{Provide a template for what kinds of policy should be implemented like this and template AppPAL.}
+\todo{Provide a template for what kinds of policy should be implemented like this and template \AppPAL.}
 
@@
 
-AppPAL requires all assertions be made by an explicit speaker.
+\AppPAL~requires all assertions be made by an explicit speaker.
 When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:principals}),
@@
 All the policies delegate also describe other entities that might also make assertions.
-With AppPAL policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
+With \AppPAL~policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
 This might be to a user to acknowledge a policy, or it might be to other groups in the company who are responsible for certain decisions.
@@
 Their work shows how parts of a BYOD policy can be checked for and enforced using tools.
-An AppPAL policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
-AppPAL makes explicit how the natural language policy is implemented and by which tools.
+A \AppPAL~policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
+\AppPAL~makes explicit how the natural language policy is implemented and by which tools.
 

### 2016-11-25 paper.tex
working 9-3 what a way to make a living

@@
 \usepackage{booktabs}
+\usepackage{graphicx}
 
@@
 \lstset{%
-  basicstyle=\normalfont\ttfamily\listingsize,
-  keywordstyle=\normalfont\ttfamily\slshape\listingsize,
-  stringstyle=\normalfont\sffamily\listingsize,
+  basicstyle=\normalfont\ttfamily,%\listingsize,
+  keywordstyle=\normalfont\ttfamily\slshape,%\listingsize,
+  stringstyle=\normalfont\sffamily,%\listingsize,
   columns=flexible,
@@
 \maketitle
-
 \begin{abstract}
@@
 \end{abstract}
-
 \section{Introduction}
@@
 Using our formalization we identify the common concerns and trust relationships in these policies.
-We show how \AppPAL\AppPALcan be used to implement these policies, and describe precisely the different trust relationships.
-\AppPAL\AppPALis an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies.
+We show how \AppPAL~can be used to implement these policies, and describe precisely the different trust relationships.
+\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies.
 We have found, however, that it is a good fit for other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
@@
 
+When checking a policy written in a formal language we generate a certificate that shows how the policy was satisfied.
+These certificates not only act as proof of how the policy was followed but also provide an audit trail.
+In a company decisions may be delegated to different departments.
+Knowing who made the decision in each department, and whether that was on the basis of policy or a one-off, allows auditors to see what happened when things go wrong, and builds trust that a system is working correctly.
+
 \todo{%
@@
 
-In order to describe BYOD specific conditions we instantiate \AppPAL~with four kinds of facts: \emph{can, has, is} and \emph{must}.
+To create \AppPAL~we instantiate SecPAL~with four kinds of facts specific to BYOD: \emph{can, has, is} and \emph{must}.
 Like other SecPAL-based instantiations~\cite{becker_framework_2009,aziz_secpal4dsa:_2011} we extend the syntax of facts to support these constructs.
@@
 \subsection{BYOD Policies}
+\label{ssec:byod-policies}
 
@@
     Number of rules          & 33           & 15          & 56          &             & 25          \\
-    SecPAL assertions        & 71           & 21          & 58          & 10          & 39          \\
+    \AppPAL~assertions       & 71           & 21          & 58          & 10          & 39          \\
     Policy coverage          & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) &             & 22 (88$\%$) \\
@@
   \begin{lstlisting}
-'nhs-trust' says App isInstallable if App hasMet('clinical-use-case').
-'nhs-trust' says App isInstallable if App hasMet('business-use-case').
+'nhs-trust' says App isUsable if App hasMet('clinical-use-case').
+'nhs-trust' says App isUsable if App hasMet('business-use-case').
 'nhs-trust' says 'cacpg' can-say App:A hasMet('clinical-use-case').
@@
 'nhs-trust' says App isInstallable
-  if App hasMet('final-app-approval').
+if App hasMet('final-app-approval'),
+   App isUsable.
 'nhs-trust' says 'igc' can-say App hasMet('final-app-approval').
@@
 But from the perspective of the policy itself it is more important \emph{who} makes the decision rather than \emph{what} they chose, if they have an opinion at all.
+
+\subsection{Worked Examples}
+
+Using the NHS rules for finding approved apps (\autoref{ssec:byod-policies}) as an example, suppose an employee, \emph{Alice}, wished to get an app, \emph{\ttfamily com.microsoft.office}, approved and installed on their device.
+To do so Alice would have to convince the device that \lstinline{'nhs-trust' says 'alices-phone' canInstall('com.microsoft.office').}
+Alice wishes to use the app for business so in order to satisfy the policy Alice must collect the following statements:
+\begin{itemize}
+    \newcommand{\weitemsize}[0]{\footnotesize}
+  \item {\weitemsize \lstinline{'nhs-trust' says 'com.microsoft.office' isInstallable.}\newline}
+    To do this she needs a statement from the Management of Information Group that it has a business use-case, and from the Integrated Governance Comitee that it is finally approved.
+    \begin{enumerate}\setcounter{enumi}{0}
+      \item {\weitemsize \lstinline{'mig' says 'com.microsoft.office' hasMet('business-use-case').}}
+      \item {\weitemsize \lstinline{'igc' says 'com.microsoft.office' hasMet('final-app-approval').}}
+    \end{enumerate}
+  \item {\weitemsize \lstinline{'nhs-trust' says 'com.microsoft.office' isApprovedFor('alices-device').}}
+    To get this she needs a statement from the manager responsible for Alice's device (\emph{Bob}) approving the app.
+    \begin{enumerate}\setcounter{enumi}{2}
+      \item {\weitemsize \lstinline{'bob' says 'com.microsoft.office' isApprovedFor('alices-device').}}
+      \item {\weitemsize \lstinline{'nhs-trust' says 'bob' isResponsibleFor('alices-device').}}
+    \end{enumerate}
+  \item Additionally she needs the following typing statements.
+    \begin{enumerate}\setcounter{enumi}{4}
+      \item {\weitemsize \lstinline{'nhs-trust' says 'com.microsoft.office' isApp.}} \label{item:isapp}
+      \item {\weitemsize \lstinline{'nhs-trust' says 'bob' isEmployee.}}
+    \end{enumerate}
+\end{itemize}
+
+Alice obtains the statements by contacting each of the speakers. 
+Each may either give her the statement she needs or may give her additional rules.
+For example the MIG and IGC may be happy to state their assertions (after a review).
+When checking if the app is an App in \autoref{item:isapp}, the NHS trust may be instead inclined to delegate further.
+They could reply that if the App is in the Google Play store then they're convinced it is an app.
+Alice would then have to obtain additional statements if she wanted to prove this statement.
+As with SecPAL, all assertions should be accompanied by a signature from their speaker proving they said the assertion.
+Alternatively the speaker could refuse to give the statement, either because they don't believe it to be true, or they cannot give an answer.
+In this case Alice would have to look for an alternative means to prove the statement or accept that they cannot install the app.
+
+Once all the statements have been collected into an \emph{assertion context}.
+Alice can use a SecPAL inference tool (such as AppPAL~\footnote{https://github.com/apppal/libapppal}) to check the policy has been satisfied.
 
@@
 
+\section{Comparison with Existing MDM Software}
+
+To evaluate \AppPAL~we compare it to existing MDM~tools: IBM's MaaS360, MobileIron, and VMWare's AirWatch; each leading MDM packages~\cite{rob_smith_magic_2016}.
+We used the Android versions of each of these tools as they all supported Android.
+Each of these tools support enforcing and checking compliance policies. 
+They do not, however, use policy languages to specify policies; rather they supply canned policies and rules that admins can tweak.
+
+\subsection{Acknowledgement}
+
+Acknowledgements form a large portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
+MaaS360 can associate users with devices, and can assign them into groups but there is currently no means to track what policies a user has agreed to or should be agreeing to.
+Users can be added to groups, which can have separate policies.
+One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a manual process for an administrator to set up and manage.
+
+More generally, MaaS360 does not generally provide a means to apply policies on the basis of policies.
+Consider the following rule taken from the SANS policy:
+\begin{policyrule}{SANS}
+Any non business-owned (that, is private) device must be able to connect to Company network MUST first be approved by technical personnel such as those from the Company IT department or desktop support.
+If allowed, privately-owned handheld devices MUST comply with this security policy and MUST be inventoried along with corporate handhelds, but identified as private.
+This is in order to prevent theft of corporate data with unmanaged handhelds (i.e.
+owner of device is not identified).
+\end{policyrule}
+Administrators can use MaaS360 to mark devices as privately owned.
+A group could be set up for approved privately-owend devices.
+MaaS360 allows the creation of groups based on search criteria, (\autoref{fig:maas360search}), though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
+Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically.
+
+\begin{figure}\centering
+  \includegraphics[width=\linewidth]{figures/maas-groupsearch.png}
+  \caption{MaaS360's search interface for devices.}
+  \label{fig:maas360search}
+\end{figure}
+
 \section{Related Work}

### 2016-11-27 paper.tex
updates

@@
 
-\subsection{Acknowledgement}
+Acknowledgements form a large portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
+MaaS360 can associate users with devices, and can assign them into groups.
+It can distribute documents (which could include written policies) but there is currently no means to track what policies a user has agreed to or should be agreeing to.
+Administrators could add user's who've acknowledged a policy to a group
+One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a semi-manual process for an administrator to set up and manage.
+In contrast \AppPAL{} is more flexible. 
+It allows a user to state which policies they have acknowledged, and for policies to delegate the acknowledgement of the policy to the user.
+We can require that a user must acknowledge certain policies if they meet certain criteria;
+and once the assertion has been made it can be imported wherever and used in making other decisions.
 
-Acknowledgements form a large portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
-MaaS360 can associate users with devices, and can assign them into groups but there is currently no means to track what policies a user has agreed to or should be agreeing to.
-Users can be added to groups, which can have separate policies.
-One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a manual process for an administrator to set up and manage.
-
-More generally, MaaS360 does not generally provide a means to apply policies on the basis of policies.
+This lack of flexibility extends to trying to compose policies. 
 Consider the following rule taken from the SANS policy:
@@
 \end{policyrule}
-Administrators can use MaaS360 to mark devices as privately owned.
-A group could be set up for approved privately-owend devices.
+Implementing this in an MDM solution such as MaaS360 administrators can mark some devices as being privately owned.
+A group could be set up for approved privately-owned devices.
 MaaS360 allows the creation of groups based on search criteria, (\autoref{fig:maas360search}), though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
-Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically.
+Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically, and could pass them along for inventorying.
+Again the management of \emph{who} is \emph{technical personnel} is effectively manual, with optional searches to define groups.
+Once we find those devices that we should allow to connect to a network we can add an access control rule to as well as enforcing the general security policy on.
+
+\begin{lstlisting}
+'company' says Device:D canConnectTo(Network:N)
+  if N isCompanyOwned,
+     D isPrivatelyOwned,
+     D isApproved,
+     D hasMet('security-policy').
+
+'company' says 'technical-personnel' can-say
+  Device:D isApproved.
+
+'company' says 'it-department' can-act-as 'technical-personnel'.
+'company' says' 'desktop-support' can-act-as 'technical-personnel'.
+\end{lstlisting}
+
+In contrast using \AppPAL~all relationships are specified by the policy.
+We are free to use conformance to one policy as a conditional for applying a second.
+Roles such as being \emph{technical personnel} can be given to users holding other roles without requiring manual intervention.
+Furthermore the \AppPAL~rule is presented in much the same form as the rule in the policy document:  device can do action if checks are satisfied.
 

### 2016-11-28 paper.tex
Tweaks

@@
 
-This paper describes a formalization of five BYOD policies using a variation of SecPAL called \AppPAL.
-This has been implemented atop of AppPAL---a policy language for mobile device privacy preferences~\cite{hallett_apppal_2016}.
-Using our formalization we identify the common concerns and trust relationships in these policies.
-We show how \AppPAL~can be used to implement these policies, and describe precisely the different trust relationships.
-\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies.
-We have found, however, that it is a good fit for other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
-Using these papers we identify a set of standard decisions and trust relationships used in BYOD policies.
+Using a formalization of five BYOD policies written in \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
+Our formalizations help identify the common concerns and trust relationships in these policies.
+We identify a set of standard decisions and trust relationships used in BYOD policies.
 These are BYOD idioms that are capture decisions at common requirements in BYOD policies.
 These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.
+
+We show how \AppPAL~can be used to implement these policies, and describe precisely the different trust relationships.
+\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies, and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
+We have found, however, that SecPAL is a good tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}, and our AppPAL implementation can be extended to support new types of policies.
+
+\section{Related Work}
+\label{sec:related}
+% 1-2 pages
+
+Companies have very little visibility into how their policies are implemented.
+Morrow notes that \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it it's current''}~\cite{morrow_byod_2012}.
+Even when devices have means to implement policies correctly it is difficult to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
+Our work aims to address these problem directly: by using formal languages we can link the policy to the implementation.
+
+Martinelli~\etal{}'s work has looked at creating a dynamic permissions manager, called UC-Droid,
+that can alter what Android permissions an app is given at run time based on policies~\cite{martinelli_enhancing_2016}.
+This allows companies to reconfigure an apps permissions based on whether the employee is at work, in a highly sensitive lab, or working out-of-hours.
+These kinds of policies are far more configurable than the geo-fenced base policies existing MDM packages provide.
+Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
+This allowed a company to verify the devices within their network and what servers and services they had access to, as well as providing different user's with different policies.
+
+Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_enabling_2014}.
+Their tool allows companies to distribute a secure app~store where through a combination of static and dynamic tools they can verify that apps follow fine-grained permission policies.
+Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
+Using their tool they implemented parts of a policy relating to personal networks, and data management for the NATO Communications and Information Agency~\cite{armando_developing_2016}.
+Their work shows how parts of a BYOD policy can be checked for and enforced using tools.
+A \AppPAL~policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
+\AppPAL~makes explicit how the natural language policy is implemented and by which tools.
+
+Other work has looked at using app wrapping to enforce more general policies.
+Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
+This has the advantage that the device's underlying OS needn't be modified as the apps can be modified at the source code level.
+Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}, however app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.
 
@@
 \begin{lstlisting}
-'company' says 'risk-analysis' can-say
+'company' says 'risk-analyst' can-say
   Location:L isHighSecurityFacility.
@@
 
-\todo{%
-  Why is the \AppPAL~better?
-  Clearer, explicit, can identify problems automatically.
-
-  Explain why I'm trying to build a schema for BYOD policies and why this is worth a paper.
-
-  How is this falsifiable?
-  Examples of policy idioms we can't express.
-}
-
 \section{The Details}
@@
 
-\subsection{Changes to \AppPAL}
+\subsection{Changes from SecPAL}
+\label{ssec:changes}
 
@@
 It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
-\begin{lstlisting}
-'nhs-trust' says Staff:S mustAcknowledge('avoid-installing-personal-apps').
-'nhs-trust' says Staff:S mustAcknowledge('social-media-policy').
-\end{lstlisting}
+%\begin{lstlisting}
+%'nhs-trust' says Staff:S mustAcknowledge('avoid-installing-personal-apps').
+%'nhs-trust' says Staff:S mustAcknowledge('social-media-policy').
+%\end{lstlisting}
 In this case, however, since nothing is required even an acknowledgement seems excessive.
@@
 For example, in the SANS policy the IT-Department are responsible for deciding what apps can be installed.
-\begin{policyrule}{SANS}
-  The IT Department maintains lists of allowed and unauthorized applications and makes them available to users on the intranet.
-  \normalfont
-  \begin{lstlisting}
-'company' says 'it-department' can-say App:A isInstallable.
-  \end{lstlisting}
-\end{policyrule}
+%\begin{policyrule}{SANS}
+%  The IT Department maintains lists of allowed and unauthorized applications and makes them available to users on the intranet.
+%  \normalfont
+%  \begin{lstlisting}
+%'company' says 'it-department' can-say App:A isInstallable.
+%  \end{lstlisting}
+%\end{policyrule}
 The NHS policy, however, is significantly more complicated.
@@
 
-\begin{table}\centering\footnotesize
-  \newcommand{\angledtitle}[1]{\rlap{\rotatebox{45}{#1}}\hspace{1em}}
-  \sffamily
-  \begin{tabular}{l c c c c c p{0.45\linewidth} }
-    \toprule
-    Predicate                                   & \angledtitle{Sirens}   & \angledtitle{Edinburgh} & \angledtitle{HiMMS} & \angledtitle{NHS} & \angledtitle{SANS}
-    & Description                                                                          \\
-    \midrule
-    Device \texttt{canBackupTo}(Server)         &                        & \cmark                  & \cmark              &                   &
-    & Says the device may send backups to a server.                                        \\
-    Device \texttt{canCall}(Number)             &                        &                         &                     & \cmark            & \cmark
-    & Says the telephone numbers a device can call.                                        \\
-    Device \texttt{canConnectToAP}(AP)          &                        &                         & \cmark              &                   & \cmark
-    & Says a device may associate with an access point.                                    \\
-    Device \texttt{canConnectToNetwork}(Subnet) & \cmark                 &                         &                     &                   & \cmark
-    & Says the device may connect to a network, for example computers all within a subnet. \\
-    Device \texttt{canConnectToServer}(IP)      &                        &                         &                     & \cmark            & \cmark
-    & Says the subject (a device) can connect to a given server (identified by a URL).    \\
-    Device \texttt{canInstall}(App)             &                        &                         &                     & \cmark            & \cmark
-    & Says the device may install an app.                                                  \\
-    User \texttt{canMonitor}(Device)            & \cmark                 &                         &                     & \cmark            & \cmark
-    & Says the subject (a user) can monitor and unlock a device.                           \\
-    Device \texttt{canStore}(File)              &                        &                         & \cmark              &                   & \cmark
-    & Says a device may store some data in permanent (non-transient) storage.              \\
-    User \texttt{canUse}(Device)                &                        &                         &                     & \cmark            & \cmark
-    & Says a user may use a device.                                                        \\
-    User \texttt{hasAcknowledged}(Policy)       & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
-    & Says an individual has acknowledged a policy.                                        \\
-    Device \texttt{hasFeature}(Feature)         &                        & \cmark                  &                     & \cmark            &
-    & Says a device has a feature enabled.                                                 \\
-    Something \texttt{hasMet}(Policy)           &                        & \cmark                  &                     & \cmark            & \cmark
-    & Says a something has met a given set of requirements.                                \\
-    Device \texttt{isActivated}                 &                        &                         & \cmark              & \cmark            & \cmark
-    & Specifies a device has been activated for BYOD usage.                                \\
-    App \texttt{isApp}                          & \cmark                 &                         & \cmark              & \cmark            & \cmark
-    & Specifies an app.                                                                    \\
-    Something \texttt{isApproved}               &                        &                         & \cmark              &                   & \cmark
-    & Specifies that something has been vetted and approved of.                            \\
-    Data \texttt{isData}                        & \cmark                 &                         &                     &                   & \cmark
-    & Specifies data.                                                                      \\
-    Device \texttt{isDevice}                    & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
-    & Specifies a device.                                                                  \\
-    User \texttt{isEmployee}                    & \cmark                 & \cmark                  &                     & \cmark            & \cmark
-    & Specifies that someone is an employee.                                               \\
-    Device \texttt{isEncrypted}                 &                        & \cmark                  &                     & \cmark            & \cmark
-    & Specifies a device is encrypted.                                                     \\
-    Feature \texttt{isFeature}                  & \cmark                 &                         &                     &                   & \cmark
-    & Specifies a feature.                                                                 \\
-    App \texttt{isInstallable}                  &                        &                         &                     & \cmark            & \cmark
-    & Specifies an app is installable.                                                     \\
-    Device \texttt{isLost}                      & \cmark                 &                         & \cmark              & \cmark            & \cmark
-    & Specifies a device is missing.                                                       \\
-    Device \texttt{isOwnedBy}(User)             & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
-    & Specifies something's owner.                                                         \\
-    File \texttt{isSecurityLevel}(Level)        &                        &                         & \cmark              &                   & \cmark
-    & Specifies some data as having business sensitive information.                        \\
-    Something \texttt{isString}                 & \cmark                 &                         &                     & \cmark            &
-    & Specifies a string.                                                                  \\
-    Number \texttt{isTelephoneNumber}           &                        &                         &                     & \cmark            & \cmark
-    & Specifies a telephone number.                                                        \\
-    User \texttt{mustAcknowledged}(Policy)      & \cmark                 & \cmark                  & \cmark              & \cmark            & \cmark
-    & Requires the user to acknowledge a  policy.                                          \\
-    Device \texttt{mustDisable}(Feature)        & \cmark                 & \cmark                  &                     & \cmark            & \cmark
-    & Forces the disablement of a feature (or a device).                                   \\
-    User \texttt{mustInform}(User, Something)   & \cmark                 &                         & \cmark              & \cmark            & \cmark
-    & Forces someone to disclose something to someone.                                     \\
-    Device \texttt{mustWipe}                    & \cmark                 &                         & \cmark              & \cmark            &
-    & Forces a device to be erased.                                                        \\
-    \bottomrule
-  \end{tabular}
-  \caption{Summary of \AppPAL~facts used in multiple BYOD policies.}
-  \label{tab:byod-predicates}
-\end{table}
-
 \subsection{BYOD Idioms}
 \todo{Describe common idioms and patterns found in each of the policies.  Check the earlier draft for some starting work}
-
-\subsubsection{Acknowledgement}
-
-All of the policies looked at require their subjects to be aware and acknowledge certain rules or policies, 
-  or that the company may perform certain actions.
-For example the NHS and HiMSS policies state that the company will wipe devices remotely to protect confidential information if the device is lost.
-Alongside these rules both policies also state that if an employee has personal information on the device then this will be lost also.
-The employee is required to be aware of this, and in the case of the HiMSS policy, agree to hold the company harmless for the loss.
-
-\begin{center}
-  \noindent
-  %\begin{minipage}{0.49\linewidth}
-    \begin{policyrule}{HiMSS}
-      I agree to hold XYZ Health System harmless for any loss relating to the
-      administration of PDA/Smartphone connectivity to XYZ Health System systems
-      including, but not limited to, loss of personal information stored on a
-      PDA/Smartphone due to data deletion done to protect sensitive information
-      related to XYZ Health System, its patients, members or partners.
-      \normalfont
-      \begin{lstlisting}
-'xyz-health-system' says 'user' mustAcknowledged('data-loss-policy').
-      \end{lstlisting}
-    \end{policyrule}
-  %\end{minipage}
-  %\begin{minipage}{0.49\linewidth}
-    \begin{policyrule}{NHS}
-      Individuals who have personal data of any kind stored on a corporately
-      issued mobile device must be aware that in the event of loss of the device
-      the above data wipe will include removal of all personal data.
-      \normalfont
-      \begin{lstlisting}
-'nhs-trust' says Staff:S can-say
-  S hasAcknowledged('data-loss-policy').
-'nhs-trust' says Staff:S mustAcknowledged('data-loss-policy').
-      \end{lstlisting}
-    \end{policyrule}   
-  %\end{minipage}
-\end{center}
-
-Both the SANS and the siren-company policies use acknowledgements to link to other sets of rules that employees should be aware of.
-These policies aren't further specified, and in the case of an acceptable use policy may be hard to enforce automatically.
-
-\begin{policyrule}{SANS}
-  Users MUST agree to the email security/acceptable use policy and eventually to the eCommerce security policy.
-  \begin{lstlisting}
-'company' says Employee:U mustAcknowledged('email-security'). 
-'company' says Employee:U mustAcknowledged('acceptable-use'). 
-'company' says Employee:U mustAcknowledged('ecommerce-security').
-  \end{lstlisting}
-\end{policyrule}
-\begin{policyrule}{Sirens}
-  The employee is expected to use his or her devices in an ethical manner at all times and adhere to the company's acceptable use policy.
-  \begin{lstlisting}
-'department' says Employee:E mustAcknowledged('acceptable-use').
-  \end{lstlisting}
-\end{policyrule}
-
-\todo{Provide a template for what kinds of policy should be implemented like this and template \AppPAL.}
 
@@
 \end{table}
+
+\subsubsection{Acknowledgement}
+
+All of the policies looked at require their subjects to be aware and acknowledge certain rules or policies, 
+  or that the company may perform certain actions.
+For example the NHS and HiMSS policies state that the company will wipe devices remotely to protect confidential information if the device is lost.
+Alongside these rules both policies also state that if an employee has personal information on the device then this will be lost also.
+The employee is required to be aware of this, and in the case of the HiMSS policy, agree to hold the company harmless for the loss.
+
+%\begin{center}
+%  \noindent
+%  %\begin{minipage}{0.49\linewidth}
+%    \begin{policyrule}{HiMSS}
+%      I agree to hold XYZ Health System harmless for any loss relating to the
+%      administration of PDA/Smartphone connectivity to XYZ Health System systems
+%      including, but not limited to, loss of personal information stored on a
+%      PDA/Smartphone due to data deletion done to protect sensitive information
+%      related to XYZ Health System, its patients, members or partners.
+%      \normalfont
+%      \begin{lstlisting}
+%'xyz-health-system' says 'user' mustAcknowledged('data-loss-policy').
+%      \end{lstlisting}
+%    \end{policyrule}
+%  %\end{minipage}
+%  %\begin{minipage}{0.49\linewidth}
+%    \begin{policyrule}{NHS}
+%      Individuals who have personal data of any kind stored on a corporately
+%      issued mobile device must be aware that in the event of loss of the device
+%      the above data wipe will include removal of all personal data.
+%      \normalfont
+%      \begin{lstlisting}
+%'nhs-trust' says Staff:S can-say
+%  S hasAcknowledged('data-loss-policy').
+%'nhs-trust' says Staff:S mustAcknowledged('data-loss-policy').
+%      \end{lstlisting}
+%    \end{policyrule}   
+%  %\end{minipage}
+%\end{center}
+
+Both the SANS and the siren-company policies use acknowledgements to link to other sets of rules that employees should be aware of.
+These policies aren't further specified, and in the case of an acceptable use policy may be hard to enforce automatically.
+The SANS policy requires that all employees follow an email security, acceptable use, and an ecomerce-security policy.
+The Sirens policy expects an employee to use their devices in an ethical manner and abide by an acceptable use policy.
+%\begin{policyrule}{SANS}
+%  Users MUST agree to the email security/acceptable use policy and eventually to the eCommerce security policy.
+%  \begin{lstlisting}
+%'company' says Employee:U mustAcknowledged('email-security'). 
+%'company' says Employee:U mustAcknowledged('acceptable-use'). 
+%'company' says Employee:U mustAcknowledged('ecommerce-security').
+%  \end{lstlisting}
+%\end{policyrule}
+%\begin{policyrule}{Sirens}
+%  The employee is expected to use his or her devices in an ethical manner at all times and adhere to the company's acceptable use policy.
+%  \begin{lstlisting}
+%'department' says Employee:E mustAcknowledged('acceptable-use').
+%  \end{lstlisting}
+%\end{policyrule}
+
+Rules of this form are used when there is a (usually separate) document that contains rules employees should be aware of and follow, but which the company may not have a specific set of rules they wish to enforce.
+For instance a company may have an ethical policy that says employees shouldn't use devices for criminal purposes.
+The company isn't interested in, or capable of, defining precisely what is criminal;
+  but they do trust their employees to make the right decision, and they do expect their employees to be aware of the rules.
+
+To implement these in \AppPAL~a policy author creates two rules: 
+  the first stating their employees must have acknowledged the policy,
+  the second delegating the acceptance of the policy to the employee themselves.
+\begin{lstlisting}
+'company' says Employee:E mustAcknowledged('policy').
+'company' says Employee:E can-say
+  E hasAcknowledged('policy').
+\end{lstlisting}
+
+Because \AppPAL's predicates have a predictible naming conventions (\autoref{ssec:changes}), our tooling can also generate additional rules to check if the acknowledgement has been made, and check whether there is anyone to acknowledge the policy.
+\begin{lstlisting}
+'company' says E hasSatisfiedObligationToMustAcknowledged('policy')
+  if E mustAcknowledged('policy').
+     E hasAcknowledged('policy').
+\end{lstlisting}
 
@@
 
-\section{Related Work}
-\label{sec:related}
-% 1-2 pages
-
-Companies have very little visibility into how their policies are implemented.
-Morrow notes that \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it it's current''}~\cite{morrow_byod_2012}.
-Even when devices have means to implement policies correctly it is difficult to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
-Our work aims to address these problem directly: by using formal languages we can link the policy to the implementation.
-
-Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_enabling_2014}.
-Their tool allows companies to distribute a secure app~store where through a combination of static and dynamic tools they can verify that apps follow fine-grained permission policies.
-Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
-Using their tool they implemented parts of a policy relating to personal networks, and data management for the NATO Communications and Information Agency~\cite{armando_developing_2016}.
-Their work shows how parts of a BYOD policy can be checked for and enforced using tools.
-A \AppPAL~policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
-\AppPAL~makes explicit how the natural language policy is implemented and by which tools.
-
-Costantino~\etal~created a role-based access control scheme for BYOD devices that allowed a company to verify the devices within their network and what servers and services they had access to~\cite{costantino_towards_2013}.
-The scheme relied on devices having a TPM however to ensure policy violations were reported.
-
-Other work has looked at using app wrapping to enforce more general policies.
-Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
-This has the advantage that the device's underlying OS needn't be modified as the apps can be modified at the source code level.
-Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}, however app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.
-
 \section{Conclusions}

### 2016-11-28 paper.tex
added bit on proof graphs

@@
 This rule should be checked periodically to ensure compliance.
+Our implementation contains tooling to generate these rules automatically.
 \begin{lstlisting}
@@
 
-\subsection{Worked Examples}
+\subsection{Worked Example}
 
@@
 Once all the statements have been collected into an \emph{assertion context}.
-Alice can use a SecPAL inference tool (such as AppPAL~\footnote{https://github.com/apppal/libapppal}) to check the policy has been satisfied.
+Alice can use a SecPAL inference tool (such as AppPAL\footnote{https://github.com/apppal/libapppal}) to check the policy has been satisfied.
+As well as making the decision we can output a proof tree (such as \autoref{fig:proof}).
+
+\begin{figure}\centering
+  \includegraphics[width=\linewidth]{figures/proof.png}
+  \caption{Proof tree for Alice installing her app.}
+  \label{fig:proof}
+\end{figure}
 
 \subsection{BYOD Idioms}
-\todo{Describe common idioms and patterns found in each of the policies.  Check the earlier draft for some starting work}
 
-\subsection{Delegation and Roles within Policies}
+\subsubsection{Delegation and Roles within Policies}
 
@@
 
-Because \AppPAL's predicates have a predictible naming conventions (\autoref{ssec:changes}), our tooling can also generate additional rules to check if the acknowledgement has been made, and check whether there is anyone to acknowledge the policy.
-\begin{lstlisting}
-'company' says E hasSatisfiedObligationToMustAcknowledged('policy')
-  if E mustAcknowledged('policy').
-     E hasAcknowledged('policy').
-\end{lstlisting}
-
 \section{Comparison with Existing MDM Software}

### 2016-11-30 paper.tex
i did work today

@@
 
-\section{My Idea}
+\section{Capturing BYOD Policies}
 \label{sec:idea}
@@
 To implement this policy an \ac{MDM} program, such as IBM's MaaS360 could be used~\cite{_ibm_????}.
-This software supports app wrapping (where an app is recompiled to use guarded APIs), and supports security ``checkboxes'' to enable security controls and access control policies.
-Using techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration to implement.
+This software supports app wrapping (where an app is recompiled to use guarded APIs), and supports security ``checkboxes'' to enable security controls and access control policies, geofencing could be used to apply additional policies in the area around a lab.
+Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration to implement.
 It talks of \emph{restricted environments} decided by \emph{company risk analysis}; but how is this communicated to the device?
-Does it access the list of restricted environments once, from a server?
+Does it access the list of restricted environments once from a server, are they fixed or can a device decide them for itself?
 Does it understand how the risk analysis was performed?
@@
 
-Our approach does not try to implement the policy.
-Rather we attempt to capture the goals of the policy rule so that the delegations of trust, tools used, and their configuration is made explicit.
+To answer these questions we present \AppPAL{}~as a tool for linking policies to the tools we use to implement them.
+Our approach does not try to implement the policy:
+  rather we act as a \emph{``glue-layer''} between the high-level policy and the tools and trust relationships used to implement them.
+We capture the goals of the policy rules so that the delegations of trust, tools implementing the policy and their configuration is made explicit.
 This gives us greater clarity as to which tool is being trusted to implement what policy.
-It allows us to see who is being trusted to make which decisions.
-It also lets us use automatic tools to uncover aspects of the policy that may be unimplemented~\cite{hallett_specifying_2016}.
-Continuing with the example above, this can be expressed in \AppPAL~as:
+It allows us to see who is being trusted to make which decisions,
+and to use automatic tools to uncover aspects of the policy that may be unimplemented~\cite{hallett_specifying_2016}.
+Continuing with the example above, we can write this in \AppPAL~as:
 
@@
 
-\section{The Details}
-\label{sec:details}
-% 5 pages
+\todo{How is this better than logging?}
 
-\subsection{Changes from SecPAL}
+%\section{The Details}
+%\label{sec:details}
+%% 5 pages
+
+\section{Changes from SecPAL}
 \label{ssec:changes}
@@
 
-\subsection{BYOD Policies}
+\section{BYOD Policies}
 \label{ssec:byod-policies}
@@
     Partial Acknowledgement  &               &                  &               &                    &               \\
+    \midrule
+    Principal Speaker        & company       & user             & nhs-trust     & records-management & department    \\
+    Technical Speaker        & it-department & xyz-healh-system & it-department &                    & it-department \\
     \bottomrule             \\
@@
 
-\subsection{Worked Example}
+\section{Worked Example}
 
@@
 
-\subsection{BYOD Idioms}
+\section{BYOD Idioms}
 
-\subsubsection{Delegation and Roles within Policies}
+\subsection{Delegation and Roles within Policies}
+
+Delegation is an important part of each of the policies.
+Each of the policies describe how separate entities may be responsible for making some decisions.
+This can be a delegation to an employee's manager to authorize a decision (as in the NHS policy), or to technical staff to decide what apps are part of a standard install (as in the sirens and SANS policies).
 
 \AppPAL~requires all assertions be made by an explicit speaker.
-When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:principals}),
+Other speakers can be delegated to by making a statement about what they \emph{can-say}.
+When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:summary}),
   with the exception of the HiMSS policy that is written as a contract the user is agreeing to.
@@
 
-In all of the policies we looked at the majority of the decisions are made by three speakers: 
+In all of the policies we looked at the majority of the decisions are made by three groups of speakers: 
   the company, the IT-department, and the users or employees.
@@
 The user is typically responsible for providing information, such as agreements to policies, reporting devices missing, and updating passcodes.
-
-The Sirens, SANS and NHS policy each describe an IT-department who are delegated to make some decisions.
-The HiMSS policy also describes an \emph{xyz-health-system}, who act similarly to an IT-department.
+In the Sirens, SANS and NHS policy each describe an IT-department who are delegated to make some decisions;
+  the HiMSS policy also describes an \emph{xyz-health-system}, who act similarly to an IT-department.
 These decisions are more varied and can overlap with the responsibilities of the company.
@@
 
-\begin{table}\centering\footnotesize
-  \begin{tabular}{l l}
-    \toprule
-    Policy      & Primary Speaker           \\
-    \midrule
-    {SANS}      & \emph{company}            \\
-    {HiMSS}     & \emph{user}               \\
-    {NHS}       & \emph{nhs-trust}          \\
-    {Edinburgh} & \emph{records-management} \\
-    {Sirens}    & \emph{department}         \\
-    \bottomrule
-  \end{tabular}
-  \label{tab:principals}
-  \caption{Principal Speakers from each of the policies examined.}
-\end{table}
+%\begin{table}\centering\footnotesize
+  %\begin{tabular}{l l}
+    %\toprule
+    %Policy      & Primary Speaker           \\
+    %\midrule
+    %{SANS}      & \emph{company}            \\
+    %{HiMSS}     & \emph{user}               \\
+    %{NHS}       & \emph{nhs-trust}          \\
+    %{Edinburgh} & \emph{records-management} \\
+    %{Sirens}    & \emph{department}         \\
+    %\bottomrule
+  %\end{tabular}
+  %\label{tab:principals}
+  %\caption{Principal Speakers from each of the policies examined.}
+%\end{table}
 
-\subsubsection{Acknowledgement}
+Delegations are used when a policy decision requires input from a third-party.
+An example might be that an employee's manager has to authorise an app install.
+The SecPAL \emph{can-say} statement is the basis for a delegation: we can to the HR department to state who is someones manager.
+\begin{lstlisting}
+'company' says 'hr-department' can-say Employee:E hasManager(Employee:M).
+\end{lstlisting}
+If we wish to delegate to someone where they have a specific relationship to what they're delegating we add a conditional to the can-say statement that enforces the relationship.
+\begin{lstlisting}
+'company' says Manager can-say 
+  Employee canInstall(App:A)
+  if Employee hasManager(Manager).
+\end{lstlisting}
+
+\subsection{Acknowledgement}
 
@@
 
+\begin{quote}
+Molestiae delectus suscipit aut et fugit et. Quos possimus fuga beatae ea
+quibusdam et reiciendis sunt. Neque molestias deleniti recusandae a et odit. Et
+ut quaerat libero. Delectus ipsum cupiditate sint. Molestiae sunt sunt quo
+neque et nostrum doloribus qui.
+
+At saepe itaque quam autem quae ea ea. Placeat enim sit quasi vero reiciendis
+sint tenetur. Saepe voluptatum voluptas mollitia. Rerum dolores debitis
+accusamus expedita soluta repellendus autem.
+
+Vero facere sit vitae qui aut consequatur aliquam quaerat. Fuga quos et
+excepturi voluptates architecto occaecati. Est enim quo doloribus harum.
+
+Aut doloremque voluptatem quidem est recusandae totam facere. Aut quia et
+fugiat rerum sed laborum ducimus. Culpa ut doloremque autem.
+
+Repudiandae autem eaque ea qui. Hic hic quidem laudantium fugiat eos. Omnis
+reprehenderit beatae autem. Ad voluptate quo qui omnis ea. Sit deleniti atque
+quis tenetur quam quis a voluptatem.
+\end{quote}
+
 \bibliography{paper}{}
 \bibliographystyle{plain}
+
 \end{document}

### 2016-12-07 paper.tex
updates

@@
 
-\newcommand{\AppPAL}[0]{SecPAL4BYOD}
+\newcommand{\AppPAL}[0]{SP4BYOD}
 
@@
 On top of user agreements companies may also use \ac{MDM} software, which can enforce some policies automatically.
-The use of \ac{MDM} software does not guarantee compliance however.
-One survey found over 50\% of companies with \ac{MDM} software still have un-compliant devices in their networks~\cite{mobileiron_security_labs_q4_2015}.
+The use of \ac{MDM} software does not guarantee compliance;
+  one survey found over 50\% of companies with \ac{MDM} software still have devices that do not comply with their policies in their networks~\cite{mobileiron_security_labs_q4_2015}.
 
-BYOD policies are becoming more complex and \ac{MDM} software is becoming more proficient.
+The policies are becoming more complex and \ac{MDM} software is becoming more proficient.
 There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
@@
 Our formalizations help identify the common concerns and trust relationships in these policies.
-We identify a set of standard decisions and trust relationships used in BYOD policies.
+We look at what decisions and trust relationships used in BYOD policies.
 These are BYOD idioms that are capture decisions at common requirements in BYOD policies.
@@
 Companies have very little visibility into how their policies are implemented.
-Morrow notes that \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it it's current''}~\cite{morrow_byod_2012}.
+When considering what tools a company may use Morrow notes that \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it it's current''}~\cite{morrow_byod_2012}.
 Even when devices have means to implement policies correctly it is difficult to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
@@
 Using their tool they implemented parts of a policy relating to personal networks, and data management for the NATO Communications and Information Agency~\cite{armando_developing_2016}.
-Their work shows how parts of a BYOD policy can be checked for and enforced using tools.
-A \AppPAL~policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
-\AppPAL~makes explicit how the natural language policy is implemented and by which tools.
+Their work shows how the app-specific parts of a BYOD policy can be checked for and enforced using tools but they do not look at where the checks or polices come from.
+An \AppPAL~policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
+One advantage of \AppPAL, however, is that it can be distributed. 
+We can delegate to different marketplaces (assuming they provided signed \AppPAL{} statements) to decide if an app meets different aspects of a policy, and create new ones combining them without having to rerun any checks they performed.
+This is very useful when using some static analysis tools which can take a long time to run (e.g.~TaintDroid~\cite{enck_taintdroid:_2014}).
 
-Other work has looked at using app wrapping to enforce more general policies.
+Other work has looked at using app wrapping (where an app is recompiled to use guarded APIs) to enforce more general policies.
 Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
@@
 
-Corporate BYOD policies are tricky to fully enforce.
+BYOD policies have grown increasingly complex.
 Part of their policies are prescriptive:  if you configure your device in this way you will mitigate that threat.
@@
 To implement this policy an \ac{MDM} program, such as IBM's MaaS360 could be used~\cite{_ibm_????}.
-This software supports app wrapping (where an app is recompiled to use guarded APIs), and supports security ``checkboxes'' to enable security controls and access control policies, geofencing could be used to apply additional policies in the area around a lab.
+This software supports app wrapping, and supports security ``checkboxes'' to enable security controls and access control policies, geofencing could be used to apply additional policies in the area around a lab.
 Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration to implement.
@@
 
-To answer these questions we present \AppPAL{}~as a tool for linking policies to the tools we use to implement them.
-Our approach does not try to implement the policy:
+To answer these questions we present \AppPAL{}~as a tool for linking policies to the tools we use to implement them and distributing decisions about apps.
+Our approach does not try to implement the policy by checking the apps code for programming errors:
   rather we act as a \emph{``glue-layer''} between the high-level policy and the tools and trust relationships used to implement them.
-We capture the goals of the policy rules so that the delegations of trust, tools implementing the policy and their configuration is made explicit.
+We capture the goals of the policy rules so that the delegations of trust, tools implementing the policy and their configuration are made explicit.
 This gives us greater clarity as to which tool is being trusted to implement what policy.
@@
 
-\todo{How is this better than logging?}
-
 %\section{The Details}
@@
 Like other SecPAL-based instantiations~\cite{becker_framework_2009,aziz_secpal4dsa:_2011} we extend the syntax of facts to support these constructs.
-Internally, however, they are treated as standard predicates.
 
@@
 
-When using \AppPAL~to describe policies, we also need to describe what domain each variable belongs to.
+We often need to describe the type of variables.
 This is performed through the \emph{is}-kind facts.
-\AppPAL~inherits from SecPAL (and Datalog) a safety condition that all variables in the head of an assertion must be referenced in the body.
+\AppPAL~inherits from SecPAL's (and Datalog's) safety condition that all variables in the head of an assertion must be referenced in the body.
 This leads to policies which can be noisy to read.
-For example in the SANS policy there is a rule that states that devices can only connect to wireless access points owned by the company.
-This is written in \AppPAL~as:
+In one policy we looked at (SANS) there is a rule that states that devices can only connect to wireless access points owned by the company.
 \begin{lstlisting}
@@
      Device isDevice,
-     AP isAccessPoint.
+     AP isAP.
 \end{lstlisting}
@@
 Any variable in the head of the rule of the form \texttt{\emph{Type}:Variable} is replaced by the variable and a conditional that \texttt{Variable is\emph{Type}} is added to the body.
-The rule can be re-written as shown bellow hiding the typing statements.
+The rule can be re-written as shown below hiding the typing statements, and emphasising the policy rule rather than SecPAL boiler-plate.
 \begin{lstlisting}
@@
 
-We analysed five different policies to find common idioms.
+We analysed five different policies looking for common idioms.
 The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
@@
     \midrule
-    Number of rules          & 33            & 15               & 56            &                    & 25            \\
+    Number of rules          & 33            & 15               & 56            & 20                 & 25            \\
     \AppPAL~assertions       & 71            & 21               & 58            & 10                 & 39            \\
@@
 
-Not all the rules in a policy require a \AppPAL~translation.
+Not all the rules in a policy require implementation.
 For example rule 9.1 in the NHS policy states:
@@
 
-Table~\ref{tab:byod-predicates} summarises the facts used by multiple policies.
-This shows what concerns and controls and are shared between different BYOD policies.
-All five of the policies make use of acknowledgements, for example.
+%Table~\ref{tab:byod-predicates} summarises the facts used by multiple policies.
+%This shows what concerns and controls and are shared between different BYOD policies.
+\todo{Add common policy idioms such as device configuration, restrictions, and the like to the table}
+All five of the policies make use of acknowledgements.
 This could be because not all the rules in a policy can be enforced through technical means, 
@@
 Only the NHS and SANS, the two most complex policies, describe when a device can install an app and what kinds of apps are installable.
-In both policies this is predominantly limited to a delegation to the appropriate groups to authorize an app.
+In both policies this is limited to a delegation to the appropriate groups to authorize an app.
 For example, in the SANS policy the IT-Department are responsible for deciding what apps can be installed.
@@
   been approved through the following Trust channels:
-
+% 
   Clinical apps; at the time of writing there are no apps clinically
-  approved by the Trust for use with patients~/~clients. However, if a
+  approved by the Trust for use with patients/clients. However, if a
   member of staff believes that there are clinical apps or other
-  technologies that could benefit their patients~/~clients, this should be
+  technologies that could benefit their patients/clients, this should be
   discussed with the clinical lead in the first instance and ratification
@@
   app should not be used if it has not been approved via this group.
-
+% 
   Business apps; at the time of writing there are no business (i.e.,
@@
   if it has not been approved via this group.
-
-  Following approval through Care and Clinical Policies and~/~or MIG, final
+%
+  Following approval through Care and Clinical Policies and/or MIG, final
   approval will be required through Integrated Governance Committee.
-
+%
   Use of paid apps must be agreed in advance with the device holder's line
@@
 
-\subsection{Delegation and Roles within Policies}
+\subsubsection{Delegation and Roles within Policies.}
 
@@
   with the exception of the HiMSS policy that is written as a contract the user is agreeing to.
-All the policies delegate also describe other entities that might also make assertions.
+All the policies describe multiple entities that might make assertions and be delegated to.
 With \AppPAL~policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
@@
 
-Delegations are used when a policy decision requires input from a third-party.
+In general delegations are used when a policy decision requires input from a third-party.
 An example might be that an employee's manager has to authorise an app install.
@@
 \begin{lstlisting}
-'company' says 'hr-department' can-say Employee:E hasManager(Employee:M).
+'company' says 'hr-department' can-say 
+  Employee:E hasManager(Employee:M).
 \end{lstlisting}
@@
 
-\subsection{Acknowledgement}
+\subsubsection{Acknowledgement.}
 
@@
 
-To evaluate \AppPAL~we compare it to existing MDM~tools: IBM's MaaS360, MobileIron, and VMWare's AirWatch; each leading MDM packages~\cite{rob_smith_magic_2016}.
-We used the Android versions of each of these tools as they all supported Android.
+It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and BlackBerry's Enterprise Mobility Management, both leading MDM packages~\cite{rob_smith_magic_2016}.
 Each of these tools support enforcing and checking compliance policies. 
@@
 'company' says 'it-department' can-act-as 'technical-personnel'.
-'company' says' 'desktop-support' can-act-as 'technical-personnel'.
+'company' says 'desktop-support' can-act-as 'technical-personnel'.
 \end{lstlisting}
@@
 
-\begin{quote}
-Molestiae delectus suscipit aut et fugit et. Quos possimus fuga beatae ea
-quibusdam et reiciendis sunt. Neque molestias deleniti recusandae a et odit. Et
-ut quaerat libero. Delectus ipsum cupiditate sint. Molestiae sunt sunt quo
-neque et nostrum doloribus qui.
+We have presented \AppPAL: an instantiation of SecPAL for BYOD policies.
+Using an \AppPAL~formalization of 5 BYOD policies we have identified several concerns that existing BYOD tools ignore.
+BYOD policies contain delegation and trust relationships that define who is responsible for making different decisions in a company.
+Sometimes that is administrators and technical staff deciding what is permitted with company property, and sometimes it is the user's themselves agreeing to follow a policy.
+Previous work has focussed on the technical staff's decisions and developing new ways to automate their decisions.
+Our work has instead looked at the policies at a higher level, tracking, managing and authorizing policies based on what people have said and what tools were run.
 
-At saepe itaque quam autem quae ea ea. Placeat enim sit quasi vero reiciendis
-sint tenetur. Saepe voluptatum voluptas mollitia. Rerum dolores debitis
-accusamus expedita soluta repellendus autem.
+Future work will look at integrating BYOD and static analysis tools into \AppPAL~so that they can be incorporated as part of the policy checks.
+Whilst earlier work on SecPAL based languages~\cite{becker_secpal:_2010,moritz_y_becker_secpal:_2009} has described how assertions can be distributed there hasn't been a well defined protocol for doing so.
+Future work will also work on tying down the distribution mechanisms and defining the distribution protocol and how it should integrate with the public keys signing the assertions.
 
-Vero facere sit vitae qui aut consequatur aliquam quaerat. Fuga quos et
-excepturi voluptates architecto occaecati. Est enim quo doloribus harum.
-
-Aut doloremque voluptatem quidem est recusandae totam facere. Aut quia et
-fugiat rerum sed laborum ducimus. Culpa ut doloremque autem.
-
-Repudiandae autem eaque ea qui. Hic hic quidem laudantium fugiat eos. Omnis
-reprehenderit beatae autem. Ad voluptate quo qui omnis ea. Sit deleniti atque
-quis tenetur quam quis a voluptatem.
-\end{quote}
-
+\todo{Nice ending paragraph.}
+\todo{Sort the bibliography.}
 \bibliography{paper}{}

### 2016-12-09 paper.tex
editing

@@
 \usepackage{graphicx}
+\usepackage{tablefootnote}
+\usepackage[backend=bibtex,firstinits=true]{biblatex}
+\addbibresource{paper.bib}
+\makeatletter
+\def\blx@maxline{77}
+\makeatother
 
@@
 \acrodef{MDM}{Mobile Device Management}
+\acrodef{BES}{Blackberry Enterprise Services}
 
@@
 \lstset{%
-  basicstyle=\normalfont\ttfamily,%\listingsize,
-  keywordstyle=\normalfont\ttfamily\slshape,%\listingsize,
-  stringstyle=\normalfont\sffamily,%\listingsize,
+  basicstyle=\normalfont\ttfamily\listingsize,
+  keywordstyle=\normalfont\ttfamily\slshape\listingsize,
+  stringstyle=\normalfont\sffamily\listingsize,
   columns=flexible,
@@
 On top of user agreements companies may also use \ac{MDM} software, which can enforce some policies automatically.
-The use of \ac{MDM} software does not guarantee compliance;
-  one survey found over 50\% of companies with \ac{MDM} software still have devices that do not comply with their policies in their networks~\cite{mobileiron_security_labs_q4_2015}.
+The use of \ac{MDM} software does not guarantee compliance.
+One survey found over 50\% of companies with \ac{MDM} software have devices that do not comply with their policies in their networks~\cite{mobileiron_security_labs_q4_2015}.
 
@@
 Policies contain rules that require employees to acknowledge risks, and regulations.
-An antivirus or \ac{MDM} program may be used to \emph{implement} part of a policy, but it is the policy which \emph{specifies} which software to use and when; there is no automatic way to check how a policy has been implemented and by what.
+An antivirus or \ac{MDM} program may be used to \emph{implement} part of a policy.
+It is the policy, however, that \emph{specifies} which software to use and when. 
+There is no automatic way to check how a policy has been implemented and by what.
 
@@
 \AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies, and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
-We have found, however, that SecPAL is a good tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}, and our AppPAL implementation can be extended to support new types of policies.
+We have found, however, that SecPAL is a good tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
+Our AppPAL implementation can also be easily extended to support new types of policies.
 
@@
 Companies have very little visibility into how their policies are implemented.
-When considering what tools a company may use Morrow notes that \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it it's current''}~\cite{morrow_byod_2012}.
-Even when devices have means to implement policies correctly it is difficult to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
-Our work aims to address these problem directly: by using formal languages we can link the policy to the implementation.
+When considering what tools a company may use Morrow notes \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it's current''}~\cite{morrow_byod_2012}.
+Even when devices can implement policies correctly, it is hard to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
+Our work aims to address these problems directly: by using formal languages we can link the policy to the implementation.
 
-Martinelli~\etal{}'s work has looked at creating a dynamic permissions manager, called UC-Droid,
-that can alter what Android permissions an app is given at run time based on policies~\cite{martinelli_enhancing_2016}.
+Martinelli~\etal{}'s work has looked at creating a dynamic permissions manager, called UC-Droid.
+Their tool can alter what Android permissions an app is given at run time based on policies~\cite{martinelli_enhancing_2016}.
 This allows companies to reconfigure an apps permissions based on whether the employee is at work, in a highly sensitive lab, or working out-of-hours.
@@
 Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
-This allowed a company to verify the devices within their network and what servers and services they had access to, as well as providing different user's with different policies.
+This allowed a company to verify the devices within their network and what servers and services they had access to.
+It also describe a mechanism for providing different users with different policies.
 
 Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_enabling_2014}.
-Their tool allows companies to distribute a secure app~store where through a combination of static and dynamic tools they can verify that apps follow fine-grained permission policies.
+Their tool allows companies to distribute apps through a secure app~store.
+The store ensures apps meet policies through a combination of static analysis and app rewriting with dynamic enforcement.
 Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
 Using their tool they implemented parts of a policy relating to personal networks, and data management for the NATO Communications and Information Agency~\cite{armando_developing_2016}.
-Their work shows how the app-specific parts of a BYOD policy can be checked for and enforced using tools but they do not look at where the checks or polices come from.
-An \AppPAL~policy might use BYODroid to ensure that a parts of a policy are enforced (as well as other tools for other parts).
+Their work shows how the app-specific parts of a BYOD policy can be check and enforced using tools.
+They do not look at where the checks or polices come from, however.
+An \AppPAL~policy might use BYODroid to ensure that parts of a policy are enforced (as well as other tools for other parts).
 One advantage of \AppPAL, however, is that it can be distributed. 
-We can delegate to different marketplaces (assuming they provided signed \AppPAL{} statements) to decide if an app meets different aspects of a policy, and create new ones combining them without having to rerun any checks they performed.
+We can delegate to different marketplaces to decide if an app meets different aspects of a policy.
+We can even create new stores by composing their policies and using the stores own statements about the apps.
 This is very useful when using some static analysis tools which can take a long time to run (e.g.~TaintDroid~\cite{enck_taintdroid:_2014}).
@@
 To implement this policy an \ac{MDM} program, such as IBM's MaaS360 could be used~\cite{_ibm_????}.
-This software supports app wrapping, and supports security ``checkboxes'' to enable security controls and access control policies, geofencing could be used to apply additional policies in the area around a lab.
+This software supports app wrapping, and supports security ``checkboxes'' to enable security controls and access control policies.
+Geofencing could also be used to apply additional policies in the area around a lab.
 Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration to implement.
@@
 
-To answer these questions we present \AppPAL{}~as a tool for linking policies to the tools we use to implement them and distributing decisions about apps.
+To answer these questions we present \AppPAL{}: 
+  a tool for linking policies to the tools used to implement them and distributing decisions about apps.
 Our approach does not try to implement the policy by checking the apps code for programming errors:
@@
 It allows us to see who is being trusted to make which decisions,
-and to use automatic tools to uncover aspects of the policy that may be unimplemented~\cite{hallett_specifying_2016}.
+  and use automatic-tools to uncover problematic aspects of the policy~\cite{hallett_specifying_2016}.
 Continuing with the example above, we can write this in \AppPAL~as:
@@
 
-When checking a policy written in a formal language we generate a certificate that shows how the policy was satisfied.
-These certificates not only act as proof of how the policy was followed but also provide an audit trail.
+After checking a policy written in \AppPAL, we generate a proof tree that shows how the policy was satisfied.
+These proof trees not only show how the policy was followed but also provide an audit trail.
 In a company decisions may be delegated to different departments.
-Knowing who made the decision in each department, and whether that was on the basis of policy or a one-off, allows auditors to see what happened when things go wrong, and builds trust that a system is working correctly.
+Auditors can see what happened when things go wrong.
+They know who made what decision, and whether they made it on the basis of a policy or as a stated fact.
 
@@
 \end{lstlisting}
-To simplify the presentation we allow a special sugared notation for conditionals expressing the type of a variable (\emph{variable \emph{is}Type}).
-Any variable in the head of the rule of the form \texttt{\emph{Type}:Variable} is replaced by the variable and a conditional that \texttt{Variable is\emph{Type}} is added to the body.
+To simplify the policies we add syntactic sugar for statements giving the type of a variable (\emph{variable \emph{is}Type}).
+Variables in the head of the rule of the form \texttt{\emph{Type}:Variable} are replaced by the variable and a condition \texttt{Variable is\emph{Type}} is added to the body.
 The rule can be re-written as shown below hiding the typing statements, and emphasising the policy rule rather than SecPAL boiler-plate.
@@
   a US non-profit company trying to improve healthcare through IT.
-It is relatively short but interesting because it contains concerns specific to healthcare scenarios, and is written as a contract the users are agreeing to follow.
-In contrast every other policy we looked at is written as a series of rules users should follow to ensure compliance.
+It is relatively short but interesting because it contains concerns specific to healthcare scenarios, and is written as a contract the users agree to follow.
+In contrast, every other policy we looked at is written as a series of rules users should follow to ensure compliance.
 The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
-The third is taken from an British hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital.
-Finally two simpler policies are taken from The University of Edinburgh~\cite{david_williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
-These policies are simpler than the other comprised of much more general rules and only a page or two in length.
+The third is taken from a British hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital.
+Finally, two simpler policies are taken from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
+These policies are simpler than the other policies we looked at comprised of much more general rules and around two pages in length.
 
 \begin{table}\centering\footnotesize\sffamily
+  \resizebox{\textwidth}{!}{%
+    \def\myfntx{}
     \begin{tabular}{l c c c c c}
@@
       \AppPAL~assertions              & 71           & 21          & 58          & 10                                & 39          \\
-    Policy coverage          & 33 (100$\%$)  & 14 (93$\%$)      & 40 (71$\%$)   &                     & 22 (88$\%$)   \\
+      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large nmuber of rules that whilst marked as rules are infact just descriptions of the document.
+All the policy rules that described restrictions or relationships were implemented in \AppPAL{}.} & 22 (88$\%$) \\
       \midrule
-    Delegations              & 23            & 5                & 33            & 2                  & 13            \\
-    Acknowledgement          &               &                  &               &                    &               \\
-    Partial Acknowledgement  &               &                  &               &                    &               \\
+      Rules using Acknowledgement     & 2            & 10          & 11          & 1                                 & 6           \\
+      Rules using Delegation          & 23           & 5           & 33          & 2                                 & 13          \\
+      Rules describing a restriction  & 18           & 3           & 8           & 1                                 & 5           \\
       \midrule
       Principal Speaker               & company      & user        & nhs-trust   & records-management                & department  \\
-    Technical Speaker        & it-department & xyz-healh-system & it-department &                    & it-department \\
+    %Technical Speaker            & it-department & xyz-healh-system & it-department &                    & it-department \\
       \bottomrule                    \\
     \end{tabular}
+  }
   \label{tab:summary}
@@
 %This shows what concerns and controls and are shared between different BYOD policies.
-\todo{Add common policy idioms such as device configuration, restrictions, and the like to the table}
 All five of the policies make use of acknowledgements.
-This could be because not all the rules in a policy can be enforced through technical means, 
-  or it could indicate policy authors care more that the subjects are aware of the rules than they do for rigorous enforcement.
+This could be because not all the rules in a policy can be enforced through technical means. 
+It could indicate policy authors care more that the subjects are aware of the rules than they do for rigorous enforcement.
 All but the HiMMS policy have rules that include locking down a device by disabling features.
-All but the Edinburgh policy have rules that look at what should happen when a device is lost,
-  and have rules that require employees inform someone when something happens.
+All but the Edinburgh policy have rules that look at what should happen when a device is lost.
+The rest have rules that require employees inform someone when something happens.
 These common concerns suggest where future \ac{MDM} software should focus their efforts.
@@
 The NHS policy, however, is significantly more complicated.
-Apps have to be approved by three different groups (the Integrated Governance Committee, the Employee's manager, and the relevant group within the trust depending on where the employee works) before the trust will say that an app can be installed.
+Apps have to be approved by three different groups\footnote{The Integrated Governance Committee, the Employee's manager, and the relevant group within the trust depending on where the employee works.} before the trust will say that an app can be installed.
 \begin{policyrule}{NHS}
@@
 This does not appear to be the case, however. 
-As part of selecting the apps an IT department or group may chose to use advanced instrumentation and policies~\cite{armando_enabling_2014}, or they may manually chose apps to form a curated app store as some \ac{MDM} vendors allow.
-But from the perspective of the policy itself it is more important \emph{who} makes the decision rather than \emph{what} they chose, if they have an opinion at all.
+As part of selecting the apps an IT department or group may chose to use advanced instrumentation and policies~\cite{armando_enabling_2014}. 
+Alternatively they may manually chose apps to form a curated app store as some \ac{MDM} vendors allow.
+From the perspective of the policy it is more important \emph{who} makes the decision rather than \emph{what} they chose, however.
 
@@
 
-Using the NHS rules for finding approved apps (\autoref{ssec:byod-policies}) as an example, suppose an employee, \emph{Alice}, wished to get an app, \emph{\ttfamily com.microsoft.office}, approved and installed on their device.
+As an example consider the NHS rules for finding approved apps (\autoref{ssec:byod-policies}).
+Suppose an employee, \emph{Alice}, wished to get an app, \emph{\ttfamily com.microsoft.office}, installed on their device.
 To do so Alice would have to convince the device that \lstinline{'nhs-trust' says 'alices-phone' canInstall('com.microsoft.office').}
@@
   \item {\weitemsize \lstinline{'nhs-trust' says 'com.microsoft.office' isInstallable.}\newline}
-    To do this she needs a statement from the Management of Information Group that it has a business use-case, and from the Integrated Governance Comitee that it is finally approved.
+    For this she needs a statement from the Management of Information Group that it has a business use-case.
+    She also needs approval from the Integrated Governance Committee.
     \begin{enumerate}\setcounter{enumi}{0}
@@
 
-Once all the statements have been collected into an \emph{assertion context}.
+When the statements have been collected 
   Alice can use a SecPAL inference tool (such as AppPAL\footnote{https://github.com/apppal/libapppal}) to check the policy has been satisfied.
-As well as making the decision we can output a proof tree (such as \autoref{fig:proof}).
+As well as making the decision we can output a proof tree.% (such as \autoref{fig:proof}).
 
-\begin{figure}\centering
-  \includegraphics[width=\linewidth]{figures/proof.png}
-  \caption{Proof tree for Alice installing her app.}
-  \label{fig:proof}
-\end{figure}
+%% We don't have room for this
+%\begin{figure}\centering
+%  \includegraphics[width=\linewidth]{figures/proof.png}
+%  \caption{Proof tree for Alice installing her app.}
+%  \label{fig:proof}
+%\end{figure}
 
 \section{BYOD Idioms}
+
+When examining the policies we noticed two particular idioms reappearing: acknowledgements and delegation.
+We describe both idioms in greater detail, and show how they can be implemented in \AppPAL{}, below.
+MDM tools and research have focussed so far on implementing restrictions on apps and devices~\cite{_ibm_????,armando_formal_2014,martinelli_enhancing_2016}.
+Implementing these controls is a vital aspect of BYOD policies and all 5 of the policies we looked at had rule that described restrictions~(\autoref{tab:summary}).
+Every policy also contained rules that required employees acknowledgements, however.
+Only the SANS policy (which is configuration focussed) contained more rules that required restrictions than acknowledgements. 
+All the policies contained more rules featuring delegation relationships than functionality restrictions.
+Restricting device functionality is tricky and important, but other aspects of BYOD policies are also interesting to look at.
 
@@
 Each of the policies describe how separate entities may be responsible for making some decisions.
-This can be a delegation to an employee's manager to authorize a decision (as in the NHS policy), or to technical staff to decide what apps are part of a standard install (as in the sirens and SANS policies).
+This can be a delegation to an employee's manager to authorize a decision (as in the NHS policy). 
+It could be to technical staff to decide what apps are part of a standard install (as in the sirens and SANS policies).
 
@@
 Other speakers can be delegated to by making a statement about what they \emph{can-say}.
-When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:summary}),
-  with the exception of the HiMSS policy that is written as a contract the user is agreeing to.
+When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:summary}).
+The HiMSS policy is different: it  is written as a contract the user agrees to.
+In the HiMMS case the user is the primary speaker.
 All the policies describe multiple entities that might make assertions and be delegated to.
@@
 
-In all of the policies we looked at the majority of the decisions are made by three groups of speakers: 
+In all the policies we looked at the majority of the decisions are made by three groups of speakers: 
   the company, the IT-department, and the users or employees.
-All of the policies also delegate to a user (apart from HiMSS where the user is the primary speaker).
+All the policies also delegate to a user (apart from HiMSS where the user is the primary speaker).
 The user is typically responsible for providing information, such as agreements to policies, reporting devices missing, and updating passcodes.
-In the Sirens, SANS and NHS policy each describe an IT-department who are delegated to make some decisions;
-  the HiMSS policy also describes an \emph{xyz-health-system}, who act similarly to an IT-department.
+In the Sirens, SANS and NHS policy each describe an IT-department who are delegated to make some decisions.
+The HiMSS policy describes an \emph{xyz-health-system} who act similarly to an IT-department.
 These decisions are more varied and can overlap with the responsibilities of the company.
@@
 
-In general delegations are used when a policy decision requires input from a third-party.
-An example might be that an employee's manager has to authorise an app install.
+Delegations are used when a policy decision requires an input from a third-party:
+  for example an employee's manager has to authorise an app install.
 The SecPAL \emph{can-say} statement is the basis for a delegation: we can to the HR department to state who is someones manager.
@@
 \end{lstlisting}
-If we wish to delegate to someone where they have a specific relationship to what they're delegating we add a conditional to the can-say statement that enforces the relationship.
+If we wish to delegate to someone we can add conditionals to the can-say statement that enforces any relationship between the delegating and delegated parties.
 \begin{lstlisting}
@@
 
-All of the policies looked at require their subjects to be aware and acknowledge certain rules or policies, 
+All the policies looked at require their subjects to be aware and acknowledge certain rules or policies, 
   or that the company may perform certain actions.
 For example the NHS and HiMSS policies state that the company will wipe devices remotely to protect confidential information if the device is lost.
-Alongside these rules both policies also state that if an employee has personal information on the device then this will be lost also.
+Alongside these rules both policies also state that if an employee has personal information on the device then this will be lost as well.
 The employee is required to be aware of this, and in the case of the HiMSS policy, agree to hold the company harmless for the loss.
@@
 The SANS policy requires that all employees follow an email security, acceptable use, and an ecomerce-security policy.
-The Sirens policy expects an employee to use their devices in an ethical manner and abide by an acceptable use policy.
+The Sirens policy expects an employee to use their devices ethically and abide by an acceptable use policy.
 %\begin{policyrule}{SANS}
@@
 
-Rules of this form are used when there is a (usually separate) document that contains rules employees should be aware of and follow, but which the company may not have a specific set of rules they wish to enforce.
+Acknowledgements are used when there is a (usually separate) document of rules employees should be aware of and follow.
+The company may not have wish to enforce these separate rules automatically, however.
 For instance a company may have an ethical policy that says employees shouldn't use devices for criminal purposes.
-The company isn't interested in, or capable of, defining precisely what is criminal;
-  but they do trust their employees to make the right decision, and they do expect their employees to be aware of the rules.
+The company isn't interested in, or capable of, defining precisely what is criminal.
+They trust their employees to make the right decision and to be aware of the rules.
 
@@
 
-It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and BlackBerry's Enterprise Mobility Management, both leading MDM packages~\cite{rob_smith_magic_2016}.
+It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}; both leading MDM packages~\cite{rob_smith_magic_2016}.
 Each of these tools support enforcing and checking compliance policies. 
@@
 MaaS360 can associate users with devices, and can assign them into groups.
-It can distribute documents (which could include written policies) but there is currently no means to track what policies a user has agreed to or should be agreeing to.
-Administrators could add user's who've acknowledged a policy to a group
+It can distribute electronic-documents but there is no means to track what policies a user should read and follow.
+Administrators could add users who have acknowledged a policy to a group.
 One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a semi-manual process for an administrator to set up and manage.
-In contrast \AppPAL{} is more flexible. 
+Similarly, with \ac{BES} can create users, groups and groups of devices, but there is no facility for managing what users should agree to automatically.
+We can create policies that apply to groups, but not create the groups and policies based on membership of groups.
+
+\AppPAL{} is more flexible. 
 It allows a user to state which policies they have acknowledged, and for policies to delegate the acknowledgement of the policy to the user.
-We can require that a user must acknowledge certain policies if they meet certain criteria;
-and once the assertion has been made it can be imported wherever and used in making other decisions.
+We can require that a user must acknowledge certain policies if they meet certain criteria.
+Once the assertion has been made it can be imported and used in making other decisions as part of the policy.
 
@@
 \end{policyrule}
-Implementing this in an MDM solution such as MaaS360 administrators can mark some devices as being privately owned.
+To implement this using an MDM solution such as MaaS360 or \ac{BES} administrators can mark some devices as being privately owned.
 A group could be set up for approved privately-owned devices.
-MaaS360 allows the creation of groups based on search criteria, (\autoref{fig:maas360search}), though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
+MaaS360 allows the creation of groups based on search criteria, though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
 Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically, and could pass them along for inventorying.
 Again the management of \emph{who} is \emph{technical personnel} is effectively manual, with optional searches to define groups.
-Once we find those devices that we should allow to connect to a network we can add an access control rule to as well as enforcing the general security policy on.
+Once we find those devices that can connect we could add an access control rule to a group containing them to enforce the rule.
+
+In \AppPAL~the policy specifies all roles and relationships.
+We are free to use conformance to one policy as a conditional for applying a second, allowing dynamic groups.
+Roles such as being \emph{technical personnel} can be given to users holding other roles without requiring manual intervention.
+Furthermore, the \AppPAL~rule is presented in much the same form as the rule in the policy document:  device can do action if checks are satisfied.
 
@@
 
-In contrast using \AppPAL~all relationships are specified by the policy.
-We are free to use conformance to one policy as a conditional for applying a second.
-Roles such as being \emph{technical personnel} can be given to users holding other roles without requiring manual intervention.
-Furthermore the \AppPAL~rule is presented in much the same form as the rule in the policy document:  device can do action if checks are satisfied.
+\AppPAL improves upon existing MDM tools by allowing sophisticated delegation relations and by providing a declarative language for expressing policies.
+This gives greater flexibility to policy authors and allows them to write policies that depend on other policies rather than predefined settings and groups.
+It lets us track
 
-\begin{figure}\centering
-  \includegraphics[width=\linewidth]{figures/maas-groupsearch.png}
-  \caption{MaaS360's search interface for devices.}
-  \label{fig:maas360search}
-\end{figure}
+%\begin{figure}\centering
+%  \includegraphics[width=0.5\linewidth]{figures/maas-groupsearch.png}
+%  \caption{MaaS360's search interface for devices.}
+%  \label{fig:maas360search}
+%\end{figure}
 
@@
 Previous work has focussed on the technical staff's decisions and developing new ways to automate their decisions.
-Our work has instead looked at the policies at a higher level, tracking, managing and authorizing policies based on what people have said and what tools were run.
+Our work looks at the policies at a higher level tracking, managing and authorizing policies based on what people have said and what tools were run.
 
@@
 Whilst earlier work on SecPAL based languages~\cite{becker_secpal:_2010,moritz_y_becker_secpal:_2009} has described how assertions can be distributed there hasn't been a well defined protocol for doing so.
-Future work will also work on tying down the distribution mechanisms and defining the distribution protocol and how it should integrate with the public keys signing the assertions.
+Future work will also work on tying down the distribution mechanisms and defining the distribution protocol.
 
-\todo{Nice ending paragraph.}
-\todo{Sort the bibliography.}
-\bibliography{paper}{}
-\bibliographystyle{plain}
+%\bibliographystyle{plain}
+\printbibliography
 

### 2016-12-14 paper.tex
grammarly edits

@@
 \acrodef{BES}{Blackberry Enterprise Services}
+\acrodef{HiMSS}{Healthcare Information Management System Society}
 
@@
   Various tools and \ac{MDM} solutions can implement parts of the policies, but there isn't a means to tie the written policy to its enforcement method.
-  In this paper we show how \AppPAL~can be used to capture policies for BYOD.
+  The \AppPAL~language is shown capturing BYOD policies throughout the paper.
   We identify common BYOD idioms and delegation structures.
@@
   85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}.
-This creates a challenge for IT departments.
-They need to control access to company resources, but have limited control over the devices used to access them.
+Controlling employee's devices is a challenge for IT departments.
+They need to control access to company resources but have limited control over the devices used to access them.
 
-One common solution to control devices is to require users to agree to follow a policy.
-These policies take the form of user agreements, written in natural language, which describe how the devices should be used and configured.
-Various guides are available for companies wishing to implement a policy from governments, standards bodies, and companies seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
-On top of user agreements companies may also use \ac{MDM} software, which can enforce some policies automatically.
-The use of \ac{MDM} software does not guarantee compliance.
+One solution to controlling devices is to require users to agree to follow policies.
+They often take the form of a user agreement, written in natural language, which describes how the devices should be used and configured.
+Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
+On top of user agreements, companies may also use \ac{MDM} software; which can enforce policies.
+But the use of \ac{MDM} software does not guarantee compliance.
 One survey found over 50\% of companies with \ac{MDM} software have devices that do not comply with their policies in their networks~\cite{mobileiron_security_labs_q4_2015}.
 
-The policies are becoming more complex and \ac{MDM} software is becoming more proficient.
+The policies are becoming more intricate.
+\ac{MDM} software is becoming more proficient.
 There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
-Enforcing a policy is only part of the problem, however.
+Implementing a policy is only part of the problem, however.
 BYOD policies are specified informally using natural language, and they contain more than just access control decisions.
@@
 Policies contain rules that require employees to acknowledge risks, and regulations.
-An antivirus or \ac{MDM} program may be used to \emph{implement} part of a policy.
-It is the policy, however, that \emph{specifies} which software to use and when. 
-There is no automatic way to check how a policy has been implemented and by what.
+An antivirus or \ac{MDM} program may be used to \emph{implement} part of the policy.
+But it is the policy, however, that \emph{specifies} which software to use and when. 
+There is no automatic way to check how the policy has been implemented and by what.
 
-Using a formalization of five BYOD policies written in \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
-Our formalizations help identify the common concerns and trust relationships in these policies.
+Using a formalization of five BYOD policies written using \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
+Our formalizations pick out the common concerns and trust relationships in these policies.
 We look at what decisions and trust relationships used in BYOD policies.
-These are BYOD idioms that are capture decisions at common requirements in BYOD policies.
+These are BYOD idioms that are capture decisions and frequently seen requirements in BYOD policies.
 These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.
 
-We show how \AppPAL~can be used to implement these policies, and describe precisely the different trust relationships.
-\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies, and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
-We have found, however, that SecPAL is a good tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
+We show how \AppPAL~can be used to implement these policies and describe precisely the different trust relationships.
+\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
+We have found, however, that SecPAL is a useful tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
 Our AppPAL implementation can also be easily extended to support new types of policies.
@@
 
-Companies have very little visibility into how their policies are implemented.
-When considering what tools a company may use Morrow notes \emph{``particularly with the BYOD trend IT professionals don't know if anti-virus software is installed or if it's current''}~\cite{morrow_byod_2012}.
-Even when devices can implement policies correctly, it is hard to ensure devices are properly configured when they are not corporately owned~\cite{tokuyoshi_security_2013}.
+Companies lack visibility as to how they implement their policies.
+When considering what tools a company may use Morrow notes \emph{``particularly with the BYOD trend IT professionals do not know if anti-virus software is installed or if it's current''}~\cite{morrow_byod_2012}.
+Even when devices can implement policies correctly, it is hard to properly configure devices that are not owned by the company~\cite{tokuyoshi_security_2013}.
 Our work aims to address these problems directly: by using formal languages we can link the policy to the implementation.
 
-Martinelli~\etal{}'s work has looked at creating a dynamic permissions manager, called UC-Droid.
-Their tool can alter what Android permissions an app is given at run time based on policies~\cite{martinelli_enhancing_2016}.
-This allows companies to reconfigure an apps permissions based on whether the employee is at work, in a highly sensitive lab, or working out-of-hours.
-These kinds of policies are far more configurable than the geo-fenced base policies existing MDM packages provide.
+Martinelli~\etal{}'s looks at creating a dynamic permissions manager, called UC-Droid.
+Their tool can alter what an app's Android permissions are at run time based on policies~\cite{martinelli_enhancing_2016}.
+The tool allows companies to reconfigure the apps depending on whether the employee is at work, in a secret lab, or working out-of-hours.
+These kinds of policies are more configurable than the geo-fenced based policies existing MDM packages provide.
 Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
-This allowed a company to verify the devices within their network and what servers and services they had access to.
-It also describe a mechanism for providing different users with different policies.
+The work allowed a company to verify the devices within their network and what servers and services they could access.
+It also describes a mechanism for providing different users with different policies.
 
@@
 Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
-Using their tool they implemented parts of a policy relating to personal networks, and data management for the NATO Communications and Information Agency~\cite{armando_developing_2016}.
-Their work shows how the app-specific parts of a BYOD policy can be check and enforced using tools.
-They do not look at where the checks or polices come from, however.
+Using their tool, they implemented parts of a NATO Communications and Information Agency policy relating to personal networks and data management~\cite{armando_developing_2016}.
+Their work shows how the app-specific sections of a BYOD policy can be check and enforced using tools.
+They did not look at where the checks or policies come from, however.
+
 An \AppPAL~policy might use BYODroid to ensure that parts of a policy are enforced (as well as other tools for other parts).
-One advantage of \AppPAL, however, is that it can be distributed. 
-We can delegate to different marketplaces to decide if an app meets different aspects of a policy.
-We can even create new stores by composing their policies and using the stores own statements about the apps.
-This is very useful when using some static analysis tools which can take a long time to run (e.g.~TaintDroid~\cite{enck_taintdroid:_2014}).
+Using \AppPAL, we can distribute policies by sharing signed statements from different principals.
+We can delegate to other marketplaces to decide if an app meets different parts of policies.
+We can even create new stores by composing their policies and using multiple store's statements about the apps.
+Distributing checks like this is useful when using some static analysis tools which can take a long time to run (e.g.~TaintDroid~\cite{enck_taintdroid:_2014}).
 
@@
 Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
-This has the advantage that the device's underlying OS needn't be modified as the apps can be modified at the source code level.
-Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}, however app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.
+App rewriting has the advantage that the device's underlying OS needn't be modified as the apps are changed at the source code level.
+Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}. 
+However app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.
 
@@
 
-BYOD policies have grown increasingly complex.
-Part of their policies are prescriptive:  if you configure your device in this way you will mitigate that threat.
+As mobile devices have become more common in the workplace, BYOD policies have grown increasingly complex
+Part of their policies are prescriptive:  if you configure your device in this way, you will mitigate that threat.
 The policies contain more than just configuration, however.
@@
 
-To implement this policy an \ac{MDM} program, such as IBM's MaaS360 could be used~\cite{_ibm_????}.
-This software supports app wrapping, and supports security ``checkboxes'' to enable security controls and access control policies.
+A company could make use of an \ac{MDM} program, such as IBM's MaaS360 or \ac{BES}~\cite{_ibm_????,_secure_????} to implement the policy.
+These programs support app wrapping, and security ``checkboxes'' to enable security controls and access control policies.
 Geofencing could also be used to apply additional policies in the area around a lab.
-Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration to implement.
+Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration.
 It talks of \emph{restricted environments} decided by \emph{company risk analysis}; but how is this communicated to the device?
 Does it access the list of restricted environments once from a server, are they fixed or can a device decide them for itself?
-Does it understand how the risk analysis was performed?
-Can it decide on the basis of a policy if a location is restricted?
+Does it understand how the risk was analysed?
+Can it judge using a policy if a location is restricted?
 The rule also gives a security objective: \emph{prevent malicious users from sending proprietary pictures}.
-The guidelines are given, however, for the specific cases of a compliant user using MMS or email.
-This may not be sufficient to stop a sufficiently motivated \emph{malicious} user.
+The guidelines are given, however, for the case of a legitimate user using MMS or email.
+It may not be sufficient to stop a sufficiently motivated \emph{malicious} user.
 
@@
   a tool for linking policies to the tools used to implement them and distributing decisions about apps.
-Our approach does not try to implement the policy by checking the apps code for programming errors:
-  rather we act as a \emph{``glue-layer''} between the high-level policy and the tools and trust relationships used to implement them.
+Our approach does not try to enforce the policy by checking the app's code for programming errors.
+Rather we act as a \emph{``glue-layer''} between the high-level policy and the tools and trust relationships used to implement them.
 We capture the goals of the policy rules so that the delegations of trust, tools implementing the policy and their configuration are made explicit.
@@
 Auditors can see what happened when things go wrong.
-They know who made what decision, and whether they made it on the basis of a policy or as a stated fact.
+They know who made what decision, and whether they made it through following policy rules or as a stated fact.
 
@@
 
-To create \AppPAL~we instantiate SecPAL~with four kinds of facts specific to BYOD: \emph{can, has, is} and \emph{must}.
+To create \AppPAL~we instantiate SecPAL~with four kinds of facts common in BYOD policies: \emph{can, has, is} and \emph{must}.
 Like other SecPAL-based instantiations~\cite{becker_framework_2009,aziz_secpal4dsa:_2011} we extend the syntax of facts to support these constructs.
@@
 
-Facts of the \emph{must}-kind represent obligations, actions that should be completed if a particular scenario presents itself.
-For these facts we add a rule to ensure the obligation is performed.
+Facts of the \emph{must}-kind represent obligations, actions to complete if a particular scenario presents itself.
+For these facts, we add a rule to check the we perform the obligation.
 This rule should be checked periodically to ensure compliance.
@@
 We often need to describe the type of variables.
-This is performed through the \emph{is}-kind facts.
-\AppPAL~inherits from SecPAL's (and Datalog's) safety condition that all variables in the head of an assertion must be referenced in the body.
-This leads to policies which can be noisy to read.
-In one policy we looked at (SANS) there is a rule that states that devices can only connect to wireless access points owned by the company.
+Facts using \emph{is} predicates give types to variables; 
+  however, \AppPAL~inherits from SecPAL's (and Datalog's) safety condition that the body of an assertion must reference all the variables in the head.
+Following the rule leads to policies which can be noisy to read.
+In one policy we looked at (SANS), there is a rule that states that devices can only connect to wireless access points owned by the company.
 \begin{lstlisting}
@@
 \end{lstlisting}
-To simplify the policies we add syntactic sugar for statements giving the type of a variable (\emph{variable \emph{is}Type}).
+To simplify the policies, we add syntactic sugar for statements giving variables their type (\emph{variable \emph{is}Type}).
 Variables in the head of the rule of the form \texttt{\emph{Type}:Variable} are replaced by the variable and a condition \texttt{Variable is\emph{Type}} is added to the body.
-The rule can be re-written as shown below hiding the typing statements, and emphasising the policy rule rather than SecPAL boiler-plate.
+To hide typing statements we rewrite the rule as shown below emphasising the policy rule rather than SecPAL boiler-plate.
 \begin{lstlisting}
@@
 The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
-This is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices.
+This policy is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices.
 Companies are expected to modify the document to suit their needs.
 The policy is general; not specific to any particular industry, device, or country's legislation.
-The second is taken from Healthcare Information Management System Society~\cite{healthcare_information_and_management_systems_society_mobile_2012};
+The second is taken from the \ac{HiMSS}~\cite{healthcare_information_and_management_systems_society_mobile_2012};
   a US non-profit company trying to improve healthcare through IT.
-It is relatively short but interesting because it contains concerns specific to healthcare scenarios, and is written as a contract the users agree to follow.
-In contrast, every other policy we looked at is written as a series of rules users should follow to ensure compliance.
+The \ac{HiMSS} policy is relatively short and contains concerns specific to healthcare scenarios. 
+It is written as a contract the users agree to follow.
+In contrast, every other policy we looked at is written as an organisation imposing rules on users they should follow to ensure compliance.
 The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
-The third is taken from a British hospital trust~\cite{kennington_mobiles_2014}, and describes the BYOD scheme used in practice at the hospital.
-Finally, two simpler policies are taken from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
+The third is taken from a British hospital trust~\cite{kennington_mobiles_2014} and describes the BYOD scheme used in practice at the hospital.
+Finally, we looked at two simpler policies from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
 These policies are simpler than the other policies we looked at comprised of much more general rules and around two pages in length.
@@
   }
+  \caption{Summary of the contents of each of the BYOD policies.}
   \label{tab:summary}
-  \caption{Summary of the contents of each of the BYOD policies.}
 \end{table}
 
-The policies are summarised in \autoref{tab:summary}.
-Each policy contains a series of \emph{rules}, which are implemented by one or more \emph{\AppPAL~assertions}.
-The \emph{policy coverage} represents the number of rules that have a \AppPAL~description attached.
+We summarise the policies in \autoref{tab:summary}.
+Each policy contains a series of \emph{rules}, which we implemented by one or more \emph{\AppPAL~assertions}.
+The \emph{policy coverage} represents the number of rules that have an \AppPAL~description attached.
 
@@
 \end{policyrule}
-The rule does not prohibit any action, and doesn't require the staff to do anything different.
+The rule does not prohibit any action and doesn't require the staff to do anything different.
 It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
@@
 All five of the policies make use of acknowledgements.
-This could be because not all the rules in a policy can be enforced through technical means. 
+The use of acknowledgements could be because enforcing the rules in a policy through technical means is undesirable. 
 It could indicate policy authors care more that the subjects are aware of the rules than they do for rigorous enforcement.
-All but the HiMMS policy have rules that include locking down a device by disabling features.
-All but the Edinburgh policy have rules that look at what should happen when a device is lost.
+All but the \ac{HiMSS} policy have rules that include locking down a device by disabling features.
+All but the Edinburgh policy have rules that look at what should happen if a user loses their device.
 The rest have rules that require employees inform someone when something happens.
-These common concerns suggest where future \ac{MDM} software should focus their efforts.
+Common concerns, such as these, suggest where future \ac{MDM} software should focus their efforts.
 
-Only the NHS and SANS, the two most complex policies, describe when a device can install an app and what kinds of apps are installable.
-In both policies this is limited to a delegation to the appropriate groups to authorize an app.
+Only the NHS and SANS policies, the two most complex policies, describe when a device can install an app and what kinds of apps are installable.
+In both policies, this is limited to a delegation to the appropriate groups to authorize an app.
 For example, in the SANS policy the IT-Department are responsible for deciding what apps can be installed.
@@
 The NHS policy, however, is significantly more complicated.
-Apps have to be approved by three different groups\footnote{The Integrated Governance Committee, the Employee's manager, and the relevant group within the trust depending on where the employee works.} before the trust will say that an app can be installed.
+Apps have to be approved by three different groups\footnote{The Integrated Governance Committee, the Employee's manager, and the relevant group within the trust depending on where the employee works.} before the Trust will say that an employee can install an app.
 \begin{policyrule}{NHS}
@@
 \end{policyrule}
-Given that user's have privacy preferences expressible as permissions~\cite{lin_modeling_2014}, we might expect corporate policies to also express preferences about what permissions an app can have.
+Given that user's have privacy preferences expressible as permissions~\cite{lin_modeling_2014}, we might expect corporate policies to describe what permissions an app can have.
 This does not appear to be the case, however. 
-As part of selecting the apps an IT department or group may chose to use advanced instrumentation and policies~\cite{armando_enabling_2014}. 
-Alternatively they may manually chose apps to form a curated app store as some \ac{MDM} vendors allow.
-From the perspective of the policy it is more important \emph{who} makes the decision rather than \emph{what} they chose, however.
+As part of selecting the apps, an IT department or group may choose to use advanced instrumentation and policies~\cite{armando_enabling_2014}. 
+Alternatively, they may manually chose apps to form a curated app store as some \ac{MDM} vendors allow.
+From the perspective of the policy, it is more important \emph{who} makes the decision rather than \emph{what} they chose, however.
 
@@
 
-As an example consider the NHS rules for finding approved apps (\autoref{ssec:byod-policies}).
+As a worked-example consider the NHS rules for finding approved apps (\autoref{ssec:byod-policies}).
 Suppose an employee, \emph{Alice}, wished to get an app, \emph{\ttfamily com.microsoft.office}, installed on their device.
-To do so Alice would have to convince the device that \lstinline{'nhs-trust' says 'alices-phone' canInstall('com.microsoft.office').}
-Alice wishes to use the app for business so in order to satisfy the policy Alice must collect the following statements:
+To do so, Alice would have to convince the device that \lstinline{'nhs-trust' says 'alices-phone' canInstall('com.microsoft.office').}
+Alice wishes to use the app for business so to satisfy the policy Alice must collect the following statements:
 \begin{itemize}
@@
   \item {\weitemsize \lstinline{'nhs-trust' says 'com.microsoft.office' isInstallable.}\newline}
-    For this she needs a statement from the Management of Information Group that it has a business use-case.
+    For this, she needs a statement from the Management of Information Group that it has a business use-case.
     She also needs approval from the Integrated Governance Committee.
@@
     \end{enumerate}
-  \item Additionally she needs the following typing statements.
+  \item Additionally, she needs the following typing statements.
     \begin{enumerate}\setcounter{enumi}{4}
@@
 Each may either give her the statement she needs or may give her additional rules.
-For example the MIG and IGC may be happy to state their assertions (after a review).
+For example, the MIG and IGC may be happy to state their assertions (after a review).
 When checking if the app is an App in \autoref{item:isapp}, the NHS trust may be instead inclined to delegate further.
-They could reply that if the App is in the Google Play store then they're convinced it is an app.
+They could reply that if the App is in the Google Play store then they are convinced it is an app.
 Alice would then have to obtain additional statements if she wanted to prove this statement.
-As with SecPAL, all assertions should be accompanied by a signature from their speaker proving they said the assertion.
-Alternatively the speaker could refuse to give the statement, either because they don't believe it to be true, or they cannot give an answer.
-In this case Alice would have to look for an alternative means to prove the statement or accept that they cannot install the app.
+As with SecPAL, all assertions should have a signature from their speaker proving they said the assertion.
+Alternatively, the speaker could refuse to give the statement, either because they do not believe it to be true, or they cannot give an answer.
+In this case, Alice would have to look for an alternative means to prove the statement or accept that they cannot install the app.
 
@@
 As well as making the decision we can output a proof tree.% (such as \autoref{fig:proof}).
+The generated proof lets auditors review how any decision was made, and verify the decision-making process.
 
-%% We don't have room for this
+%% We do not have room for this
 %\begin{figure}\centering
@@
 
-When examining the policies we noticed two particular idioms reappearing: acknowledgements and delegation.
+When examining the policies, we noticed two particular idioms reappearing: acknowledgements and delegation.
 We describe both idioms in greater detail, and show how they can be implemented in \AppPAL{}, below.
 MDM tools and research have focussed so far on implementing restrictions on apps and devices~\cite{_ibm_????,armando_formal_2014,martinelli_enhancing_2016}.
-Implementing these controls is a vital aspect of BYOD policies and all 5 of the policies we looked at had rule that described restrictions~(\autoref{tab:summary}).
+Implementing these controls is a vital aspect of BYOD policies and all 5 of the policies we looked at had rules that described restrictions~(\autoref{tab:summary}).
 Every policy also contained rules that required employees acknowledgements, however.
@@
 All the policies contained more rules featuring delegation relationships than functionality restrictions.
-Restricting device functionality is tricky and important, but other aspects of BYOD policies are also interesting to look at.
+Restricting device functionality is tricky and important, but other aspects of BYOD policies are also worth attention.
 
@@
 Delegation is an important part of each of the policies.
-Each of the policies describe how separate entities may be responsible for making some decisions.
-This can be a delegation to an employee's manager to authorize a decision (as in the NHS policy). 
+Each of the policies describes through rules how separate entities may be responsible for making some decisions.
+These rules can be a delegation to an employee's manager to authorize a decision (as in the NHS policy). 
 It could be to technical staff to decide what apps are part of a standard install (as in the sirens and SANS policies).
 
-\AppPAL~requires all assertions be made by an explicit speaker.
-Other speakers can be delegated to by making a statement about what they \emph{can-say}.
-When translating the policies the author of the policy is generally used as the speaker of the rule (\autoref{tab:summary}).
-The HiMSS policy is different: it  is written as a contract the user agrees to.
-In the HiMMS case the user is the primary speaker.
-All the policies describe multiple entities that might make assertions and be delegated to.
+\AppPAL~requires an explicit speaker for each assertion.
+Speakers can delegate to others by making a statement about what they \emph{can-say}.
+When translating the policies, the author of the policy is used as the primary speaker of the policy's rule (\autoref{tab:summary}).
+For the \ac{HiMSS} policy, where the user states what they will do rather than the company stating what they must, the user is the primary speaker.
+All the policies describe multiple entities that might make assertions and delegate.
 With \AppPAL~policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
-This might be to a user to acknowledge a policy, or it might be to other groups in the company who are responsible for certain decisions.
+The delegation might be to a user to acknowledge a policy, or it might be to other groups in the company who are responsible for certain decisions.
 
@@
   the company, the IT-department, and the users or employees.
-All the policies also delegate to a user (apart from HiMSS where the user is the primary speaker).
+All the policies also delegate to a user (apart from \ac{HiMSS} where the user is the primary speaker).
 The user is typically responsible for providing information, such as agreements to policies, reporting devices missing, and updating passcodes.
 In the Sirens, SANS and NHS policy each describe an IT-department who are delegated to make some decisions.
-The HiMSS policy describes an \emph{xyz-health-system} who act similarly to an IT-department.
+The \ac{HiMSS} policy describes an \emph{xyz-health-system} who act similarly to an IT-department.
 These decisions are more varied and can overlap with the responsibilities of the company.
-In the NHS and SANS policies the IT department is responsible for maintaining lists of activated devices.
-In the Sirens and SANS policies the IT department maintains a list of what is installable on a device or not.
+In the NHS and SANS policies, the IT department is responsible for maintaining lists of activated devices.
+In the Sirens and SANS policies, the IT department maintains a list of what is installable on a device or not.
 
@@
 
-Delegations are used when a policy decision requires an input from a third-party:
-  for example an employee's manager has to authorise an app install.
-The SecPAL \emph{can-say} statement is the basis for a delegation: we can to the HR department to state who is someones manager.
+When a policy decision requires input from a third-party delegation is used.
+For example, an employee's manager has to authorise an app install.
+The SecPAL \emph{can-say} statement is the basis for a delegation. 
+We can ask the HR department to state who is someone's manager.
 \begin{lstlisting}
@@
 \end{lstlisting}
-If we wish to delegate to someone we can add conditionals to the can-say statement that enforces any relationship between the delegating and delegated parties.
+If we wish to delegate to someone, we can add conditionals to the can-say statement that enforces any relationship between the delegating and delegated parties.
 \begin{lstlisting}
@@
   or that the company may perform certain actions.
-For example the NHS and HiMSS policies state that the company will wipe devices remotely to protect confidential information if the device is lost.
-Alongside these rules both policies also state that if an employee has personal information on the device then this will be lost as well.
-The employee is required to be aware of this, and in the case of the HiMSS policy, agree to hold the company harmless for the loss.
+For example, the NHS and \ac{HiMSS} policies state that the organisation will wipe devices remotely to protect confidential information a user loses their device.
+Both policies also say that employees would lose personal information if they had it on the device and the company needed to erase it.
+The employee is required to be aware of this, and in the case of the \ac{HiMSS} policy, agree to hold the company harmless for the loss.
 
@@
 
-Both the SANS and the siren-company policies use acknowledgements to link to other sets of rules that employees should be aware of.
-These policies aren't further specified, and in the case of an acceptable use policy may be hard to enforce automatically.
-The SANS policy requires that all employees follow an email security, acceptable use, and an ecomerce-security policy.
+Both the SANS and the siren-company policies use acknowledgements to link to other sets of rules that employees should follow.
+These policies are not further specified, and in the case of an acceptable use policy may be hard to enforce automatically.
+The SANS policy requires that all employees follow an email security, acceptable use, and an eCommerce-security policy.
 The Sirens policy expects an employee to use their devices ethically and abide by an acceptable use policy.
@@
 
-Acknowledgements are used when there is a (usually separate) document of rules employees should be aware of and follow.
+When there is a (usually separate) set of rules and concerns employees should be aware of acknowledgements are used.
 The company may not have wish to enforce these separate rules automatically, however.
-For instance a company may have an ethical policy that says employees shouldn't use devices for criminal purposes.
-The company isn't interested in, or capable of, defining precisely what is criminal.
+For instance, a company may have an ethical policy that says employees should not use devices for criminal purposes.
+The company is not interested in, or capable of, defining what is criminal.
 They trust their employees to make the right decision and to be aware of the rules.
 
-To implement these in \AppPAL~a policy author creates two rules: 
+To implement these in \AppPAL, a policy author creates two rules: 
   the first stating their employees must have acknowledged the policy,
@@
 
-It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}; both leading MDM packages~\cite{rob_smith_magic_2016}.
-Each of these tools support enforcing and checking compliance policies. 
+It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}, both leading MDM packages~\cite{rob_smith_magic_2016}.
+Each of these tools supports enforcing and checking compliance policies. 
 They do not, however, use policy languages to specify policies; rather they supply canned policies and rules that admins can tweak.
 
-Acknowledgements form a large portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
-MaaS360 can associate users with devices, and can assign them into groups.
-It can distribute electronic-documents but there is no means to track what policies a user should read and follow.
+Acknowledgements form a significant portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
+MaaS360 can associate users with devices and can assign them into groups.
+It can distribute electronic documents, but there are no means to track what policies a user should read and follow.
 Administrators could add users who have acknowledged a policy to a group.
 One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a semi-manual process for an administrator to set up and manage.
+One might imagine administrators adding users who have acknowledged some additional rules to new groups
 Similarly, with \ac{BES} can create users, groups and groups of devices, but there is no facility for managing what users should agree to automatically.
@@
 
-\AppPAL{} is more flexible. 
-It allows a user to state which policies they have acknowledged, and for policies to delegate the acknowledgement of the policy to the user.
+\AppPAL{} is more flexible than these MDM suites. 
+It allows a user to state which policies they have acknowledged and for policies to delegate the acknowledgement of the policy to the user.
 We can require that a user must acknowledge certain policies if they meet certain criteria.
-Once the assertion has been made it can be imported and used in making other decisions as part of the policy.
+Once a statement is made, it can be imported into the \AppPAL~knowledge base and used in making other decisions as part of the policy.
 
@@
 MaaS360 allows the creation of groups based on search criteria, though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
-Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically, and could pass them along for inventorying.
+Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically and could pass them along for inventorying.
 Again the management of \emph{who} is \emph{technical personnel} is effectively manual, with optional searches to define groups.
@@
 Roles such as being \emph{technical personnel} can be given to users holding other roles without requiring manual intervention.
-Furthermore, the \AppPAL~rule is presented in much the same form as the rule in the policy document:  device can do action if checks are satisfied.
+Furthermore, the \AppPAL~rule is presented in much the same form as the rule in the policy document: \emph{device} can do \emph{action} if \emph{checks} are satisfied.
 
@@
 
-\AppPAL improves upon existing MDM tools by allowing sophisticated delegation relations and by providing a declarative language for expressing policies.
-This gives greater flexibility to policy authors and allows them to write policies that depend on other policies rather than predefined settings and groups.
-It lets us track
+\AppPAL~improves upon existing MDM tools by allowing sophisticated delegation relations and by providing a declarative language for expressing policies.
+The language gives greater flexibility to policy authors and allows them to write policies that depend on other policies rather than predefined settings and groups.
+It lets us track what users have agreed to, what their policies are, how they are specified, and how they are satisfied.
 
@@
 BYOD policies contain delegation and trust relationships that define who is responsible for making different decisions in a company.
-Sometimes that is administrators and technical staff deciding what is permitted with company property, and sometimes it is the user's themselves agreeing to follow a policy.
+Sometimes that is administrators and technical staff deciding what to permit inside the company, and sometimes it is the user's themselves agreeing to follow a policy.
 Previous work has focussed on the technical staff's decisions and developing new ways to automate their decisions.
@@
 
-Future work will look at integrating BYOD and static analysis tools into \AppPAL~so that they can be incorporated as part of the policy checks.
-Whilst earlier work on SecPAL based languages~\cite{becker_secpal:_2010,moritz_y_becker_secpal:_2009} has described how assertions can be distributed there hasn't been a well defined protocol for doing so.
+Future work will incorporate new BYOD and static analysis tools into \AppPAL. 
+This will allow us to use them as part of policy checks.
+Whilst earlier work on SecPAL based languages~\cite{becker_secpal:_2010,moritz_y_becker_secpal:_2009} has described how to distribute assertions there has not been a well-defined protocol for doing so.
 Future work will also work on tying down the distribution mechanisms and defining the distribution protocol.

### 2016-12-19 paper.tex
updates

@@
   Various tools and \ac{MDM} solutions can implement parts of the policies, but there isn't a means to tie the written policy to its enforcement method.
-  The \AppPAL~language is shown capturing BYOD policies throughout the paper.
-  We identify common BYOD idioms and delegation structures.
+  The \AppPAL~language is introduced as an instantiation of SecPAL for capturing BYOD policies.
+  We identify common BYOD idioms and delegation structures, and compare how \AppPAL~might enforce policies to existing leading MDM packages.
 \end{abstract}

### 2016-12-20 paper.tex
davids feedback

--- 2016-12-14 paper.tex
@@
 \acrodef{HiMSS}{Healthcare Information Management System Society}
+\acrodef{GSN}{Goal Structuring Notation}
 
@@
   stringstyle=\normalfont\sffamily\listingsize,
+  language=AppPAL,
   columns=flexible,
@@
   breaklines=true,
+  xleftmargin=1em,
   numberstyle=\tiny\ttfamily\color{Gray},
@@
 \usepackage{mdframed}
-\newcommand{\todo}[1]{\marginpar{\begin{mdframed}[userdefinedwidth=1cm]\tiny\sffamily\raggedright #1\end{mdframed}}}
+\mdfdefinestyle{supertight}{%
+  leftmargin=0pt,
+  rightmargin=0pt,
+  skipabove=1em,
+  skipbelow=1em,
+  innerleftmargin=2pt,
+  innerrightmargin=2pt,
+  innertopmargin=2pt,
+  innerbottommargin=2pt,
+}
+
+\newcommand{\todo}[1]{\marginpar{\begin{mdframed}[style=supertight]\tiny\sffamily\raggedright #1\end{mdframed}}}
 
 \newenvironment{policyrule}[1]{%
-  \begin{mdframed}\footnotesize
+  \begin{mdframed}[]\footnotesize
       \noindent\textbf{\sffamily #1}:~\itshape%
@@
 \author{Joseph Hallett and David Aspinall}
-\institute{University of Edinburgh}
+\institute{School of Informatics, University of Edinburgh}
 \begin{document}
@@
   BYOD policies are informally specified using natural language.
-  Various tools and \ac{MDM} solutions can implement parts of the policies, but there isn't a means to tie the written policy to its enforcement method.
-  The \AppPAL~language is shown capturing BYOD policies throughout the paper.
-  We identify common BYOD idioms and delegation structures.
+  We show how the \AppPAL~language can help reduce ambiguity in BYOD policies and link the specification of a BYOD policy to its implementation.
+  This paper identifies a BYOD idiom that is ignored by existing MDM tools.
+  We explore the delegation relationships and concerns in 5 BYOD policies, and describe how \AppPAL might be used to implement them using existing tools.
 \end{abstract}
@@
 Controlling employee's devices is a challenge for IT departments.
-They need to control access to company resources but have limited control over the devices used to access them.
+Failure to do so can lead to cause employees to leak confidential information and the waste of company resources.
+Unfortunately the companies have limited control over the devices used to access them.
 
@@
 Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
-On top of user agreements, companies may also use \ac{MDM} software; which can enforce policies.
+On top of user agreements, companies may also use \ac{MDM} software which can help enforce policies.
+\ac{MDM} software can configure a device's security settings, and add 
 But the use of \ac{MDM} software does not guarantee compliance.
-One survey found over 50\% of companies with \ac{MDM} software have devices that do not comply with their policies in their networks~\cite{mobileiron_security_labs_q4_2015}.
+One survey from a leading MDM vendor found over 50\% of companies with their \ac{MDM} software still had devices that do not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.
+Reasons for non-compliance included out-of-date policies and employees tampering with the \ac{MDM} software.
 
-The policies are becoming more intricate.
-\ac{MDM} software is becoming more proficient.
+BYOD policies are becoming more intricate.
 There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
@@
 
+To describe the policies we came up with \AppPAL{}: 
+  a formal language for linking policies to the tools used to implement them and distributing decisions about apps.
 Using a formalization of five BYOD policies written using \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
@@
 
-We show how \AppPAL~can be used to implement these policies and describe precisely the different trust relationships.
-\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
-We have found, however, that SecPAL is a useful tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
-Our AppPAL implementation can also be easily extended to support new types of policies.
-
-\section{Related Work}
-\label{sec:related}
-% 1-2 pages
-
 Companies lack visibility as to how they implement their policies.
@@
 
+\AppPAL~is not designed to replace existing static and dynamic analysis and enforcement tools.
+A company might use multiple different tools, app stores, and even contractual agreements with employees to enforce their policies.
+We aim to help clarify the meaning of ambiguous natural language policy documents, and provide a rigorous means for following them.
+A company can use any \ac{MDM} tool, curated app store or user agreement to enforce their policy.
+\AppPAL~links the specification of the policy to its implementation, giving visibility into exactly how a policy is implemented and satisfied.
+
+We show how \AppPAL~can be used to implement these policies and describe precisely the different trust relationships.
+\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
+We have found, however, that SecPAL is a useful tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
+Our AppPAL implementation can also be easily extended to support new types of policies.
+It also gives us access to tooling we have developed to can check AppPAL policies for completeness and redundant statements.
+For this work additional tooling was developed to help visualise policies and describe policies contents.
+This was helpful for making comparisons between policies and checking our formalisation for mistakes.
+
+%\paragraph{Contributions}
+
+In summary, our work makes the following contributions:
+
+\begin{itemize}
+  \item We present a formalisation of five different BYOD policies in \AppPAL: a new instantiation of SecPAL for describing BYOD policies (\autoref{sec:idea}).
+  \item Using our formalisation of the policies we make comparisons between the different policies. 
+    Unlike previous work which looks at individual policies~\cite{armando_developing_2016}, our work looks at policies across a variety of domains (\autoref{ssec:byod-policies}).
+  \item We identify that delegation and acknowledgements are an important aspect of BYOD policies that current MDM software does not look at (\autoref{sec:idioms}).
+\end{itemize}
+
+\subsection{Related Work}
+\label{sec:related}
+% 1-2 pages
+
 Martinelli~\etal{}'s looks at creating a dynamic permissions manager, called UC-Droid.
@@
 The tool allows companies to reconfigure the apps depending on whether the employee is at work, in a secret lab, or working out-of-hours.
-These kinds of policies are more configurable than the geo-fenced based policies existing MDM packages provide.
+These kinds of policies are more configurable than the geofenced based policies existing MDM packages provide.
 Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
@@
 
-Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_enabling_2014}.
-Their tool allows companies to distribute apps through a secure app~store.
+Armando~\etal~developed BYODroid as a tool for enforcing BYOD policies through a secure marketplace~\cite{armando_bring_2013}.
+Their tool allows companies to distribute apps through a secure app~store~\cite{armando_enabling_2014}.
 The store ensures apps meet policies through a combination of static analysis and app rewriting with dynamic enforcement.
@@
 App rewriting has the advantage that the device's underlying OS needn't be modified as the apps are changed at the source code level.
-Many commercial \ac{MDM} solutions offer app-wrapping as a feature~\cite{_ibm_????,_app_????}. 
 However app wrapping alone without additional analysis is insufficient to enforce policies effectively~\cite{hao_effectiveness_2013}.
+
+Our approach taken with \AppPAL~is similar to work on safety cases.
+A safety case is an argument made to say a system is acceptably safe to be used in a given scenario.
+Industrial safety cases are often described in natural language, which can be ambiguous and unclear.
+\ac{GSN}~\cite{tim_kelly_goal_2004} is one approach to make the safety cases explicit.
+It is a graphical formal notation that lets engineers argue that a system is safe by linking top-level goals, to sub-goals to the arguments made for a system's safety.
+Similarly work developing a formal language for specifying how medical staff should collaborate in a healthcare scenario~\cite{papapanagiotou_formal_2014} again helps clarify how roles are filled in a medical context, on the basis of staff and different healthcare providers.
+
+It is interesting to examine how leading~\cite{rob_smith_magic_2016} MDM~tools such as IBM's MaaS360, MobileIron and \ac{BES}, enforce \emph{BYOD} policies.
+Each of these tools supports enforcing and checking compliance policies. 
+They do not, however, use policy languages to specify policies; rather they provide a limited number of checkboxes that admins can tweak (an excerpt of a policy from MaaS360 is shown in \autoref{fig:maaspolicy}).
+These tools allow administrators to configure a device's settings and provision the devices with company apps.
+Some support app wrapping, which enables them to encrypt app data locally, use a VPN within the app, or prevent apps not being used when the device isn't compliant.
+Because the policies are inflexible and tightly coupled to the device's OS, intervention by an administrator is often required.
+Whilst MDM software is good at configuring devices, selecting which policies to apply is typically a manual process performed by an administrator.
+Removing blacklisted apps is a common feature, but the selection process of which apps to remove is manual.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/maas360-policy.png}
+  \caption{Excerpt of a policy showing network settings from MaaS360.}
+  \label{fig:maaspolicy}
+\end{figure}
 
@@
 
-As mobile devices have become more common in the workplace, BYOD policies have grown increasingly complex
+As mobile devices have become more common in the workplace, BYOD policies have been written to help control the devices.
 Part of their policies are prescriptive:  if you configure your device in this way, you will mitigate that threat.
 The policies contain more than just configuration, however.
-Consider this rule taken from the \emph{Security Policy for the use of handheld devices in corporate environments} by SANS~\cite{nicholas_r._c._guerin_security_2008}:
+Consider this rule taken from the \emph{Security Policy for the use of handheld devices in corporate environments} by SANS~\cite{nicholas_r._c._guerin_security_2008}.
 
@@
 
-A company could make use of an \ac{MDM} program, such as IBM's MaaS360 or \ac{BES}~\cite{_ibm_????,_secure_????} to implement the policy.
-These programs support app wrapping, and security ``checkboxes'' to enable security controls and access control policies.
-Geofencing could also be used to apply additional policies in the area around a lab.
+A company could make use of an \ac{MDM} program.
+Some can use geofencing to apply a policies in the area around a lab.
 Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration.
@@
 Does it access the list of restricted environments once from a server, are they fixed or can a device decide them for itself?
-Does it understand how the risk was analysed?
 Can it judge using a policy if a location is restricted?
@@
 
-To answer these questions we present \AppPAL{}: 
-  a tool for linking policies to the tools used to implement them and distributing decisions about apps.
+% To answer these questions we came up with \AppPAL{}: 
+%   a formal language for linking policies to the tools used to implement them and distributing decisions about apps.
 Our approach does not try to enforce the policy by checking the app's code for programming errors.
@@
 
-\section{Changes from SecPAL}
+\section{Instantiating SecPAL}
 \label{ssec:changes}
+
+SecPAL was developed as a distributed access control language~\cite{becker_secpal:_2010}.
+The SecPAL language is designed to be have a clear readable syntax, and intuitive semantics.
+It is also designed to be extensible, which makes it ideal for extending to create new languages.
+All SecPAL statements are \emph{said} by an explicit authority.
+The authority can say a fact (that something is described by a predicate), a delegation (that someone else \emph{can-say} a fact), or a role assignment (that something \emph{can-act-as} something else).
+This statement optionally contain conditional facts, and constraints that must be satisfied before the authority will say the statement.
 
@@
 Facts of the \emph{must}-kind represent obligations, actions to complete if a particular scenario presents itself.
-For these facts, we add a rule to check the we perform the obligation.
+For these facts, we add a rule to check we perform the obligation.
 This rule should be checked periodically to ensure compliance.
-Our implementation contains tooling to generate these rules automatically.
+Our implementation contains tooling to generate these rules automatically, by parsing the policy.
 \begin{lstlisting}
@@
 
-We often need to describe the type of variables.
-Facts using \emph{is} predicates give types to variables; 
-  however, \AppPAL~inherits from SecPAL's (and Datalog's) safety condition that the body of an assertion must reference all the variables in the head.
-Following the rule leads to policies which can be noisy to read.
-In one policy we looked at (SANS), there is a rule that states that devices can only connect to wireless access points owned by the company.
+Facts using \emph{is} predicates give types to variables. 
+\AppPAL~inherits from SecPAL's (and Datalog's) safety condition that the body of a statements must reference all the variables in the head.
+This can lead to some \emph{boilerplate} code in policies that may obscure their meaning.
+To simplify the policies, we add syntactic sugar for facts giving variables their type (\emph{variable \emph{is}Type}).
+Variables in the head of the statement of the form \texttt{\emph{Type}:Variable} are replaced by the variable and a condition \texttt{Variable is\emph{Type}} is added to the condition.
+The two statements shown below (taken from the SANS policy) are equivalent, however we feel the example on the right is easier to read.
+
+\begin{minipage}{\linewidth}
+  \begin{minipage}{0.5\linewidth}
 \begin{lstlisting}
-'company' says Device canConnectToAP(AP)
-  if AP isOwnedByCompany,
+'company' says Device canConnectToAP(X)
+  if X isOwnedByCompany,
      Device isDevice,
-     AP isAP.
+     X isAP.
 \end{lstlisting}
-To simplify the policies, we add syntactic sugar for statements giving variables their type (\emph{variable \emph{is}Type}).
-Variables in the head of the rule of the form \texttt{\emph{Type}:Variable} are replaced by the variable and a condition \texttt{Variable is\emph{Type}} is added to the body.
-To hide typing statements we rewrite the rule as shown below emphasising the policy rule rather than SecPAL boiler-plate.
+\end{minipage}
+\begin{minipage}{0.5\linewidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
+\end{minipage}
 
@@
 
-We analysed five different policies looking for common idioms.
-The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
+We examined 5 policies and encoded them into \AppPAL{} looking for common idioms.
+We selected these policies as they came from a variety of domains.
+\begin{itemize}
+  \item The first is the \emph{Security Policy Template: Use of Handheld Devices in a Corporate Environment}, published by the SANS Institute~\cite{nicholas_r._c._guerin_security_2008}.
 This policy is a hypothetical policy published to help companies mitigate the threats to corporate assets caused by mobile devices.
@@
 The policy is general; not specific to any particular industry, device, or country's legislation.
-The second is taken from the \ac{HiMSS}~\cite{healthcare_information_and_management_systems_society_mobile_2012};
+\item The second is taken from the \ac{HiMSS}~\cite{healthcare_information_and_management_systems_society_mobile_2012};
   a US non-profit company trying to improve healthcare through IT.
@@
 The policy is designed as a sample agreement for a system trying to manage personal mobile devices in a healthcare environment.
-The third is taken from a British hospital trust~\cite{kennington_mobiles_2014} and describes the BYOD scheme used in practice at the hospital.
-Finally, we looked at two simpler policies from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
-These policies are simpler than the other policies we looked at comprised of much more general rules and around two pages in length.
+\item The third is taken from a British hospital trust~\cite{kennington_mobiles_2014} and describes the BYOD scheme used in practice at the hospital.
+\item Finally, we looked at two simpler policies from The University of Edinburgh~\cite{williamson_bring_2015} and a company specialising in emergency sirens~\cite{code3pse.org_sample_????}.
+  These policies are simpler, and shorter than the other policies we looked at comprised of much more general rules.
+\end{itemize}
 
@@
       Number of rules                 & 33           & 15          & 56          & 20                                & 25          \\
-      \AppPAL~assertions              & 71           & 21          & 58          & 10                                & 39          \\
+      \AppPAL~statements              & 71           & 21          & 58          & 10                                & 39          \\
       Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large nmuber of rules that whilst marked as rules are infact just descriptions of the document.
@@
 We summarise the policies in \autoref{tab:summary}.
-Each policy contains a series of \emph{rules}, which we implemented by one or more \emph{\AppPAL~assertions}.
+Each policy contains a series of \emph{rules}, which we implemented by one or more \emph{\AppPAL~statements}.
 The \emph{policy coverage} represents the number of rules that have an \AppPAL~description attached.
 
-Not all the rules in a policy require implementation.
-For example rule 9.1 in the NHS policy states:
-\begin{policyrule}{NHS}
-  Downloading of personal apps onto a corporately issued mobile device should be avoided where possible.
-  The Trust would not encourage staff members to download apps for personal use onto a corporately issued mobile device.
-  All staff are reminded that they must adhere to the guidance outlined in the Social Media Policy.
-\end{policyrule}
-The rule does not prohibit any action and doesn't require the staff to do anything different.
-It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
-%\begin{lstlisting}
-%'nhs-trust' says Staff:S mustAcknowledge('avoid-installing-personal-apps').
-%'nhs-trust' says Staff:S mustAcknowledge('social-media-policy').
-%\end{lstlisting}
-In this case, however, since nothing is required even an acknowledgement seems excessive.
+% Not all the rules in a policy require implementation.
+% For example rule 9.1 in the NHS policy states:
+% \begin{policyrule}{NHS}
+%   Downloading of personal apps onto a corporately issued mobile device should be avoided where possible.
+%   The Trust would not encourage staff members to download apps for personal use onto a corporately issued mobile device.
+%   All staff are reminded that they must adhere to the guidance outlined in the Social Media Policy.
+% \end{policyrule}
+% The rule does not prohibit any action and doesn't require the staff to do anything different.
+% It does remind staff to be aware of a rule which we could implement with an acknowledgement; a statement that the party has acknowledged a rule.
+% In this case, however, since nothing is required even an acknowledgement seems excessive.
 
@@
 Only the NHS and SANS policies, the two most complex policies, describe when a device can install an app and what kinds of apps are installable.
-In both policies, this is limited to a delegation to the appropriate groups to authorize an app.
+In both policies this expressed as a delegation to the appropriate groups to authorize an app.
 For example, in the SANS policy the IT-Department are responsible for deciding what apps can be installed.
@@
 The NHS policy, however, is significantly more complicated.
-Apps have to be approved by three different groups\footnote{The Integrated Governance Committee, the Employee's manager, and the relevant group within the trust depending on where the employee works.} before the Trust will say that an employee can install an app.
+Apps have to be approved by three different groups (the IGC, the Employee's manager, and the relevant group for either clinical or business cases) before the Trust will say that an employee can install an app.
 \begin{policyrule}{NHS}
@@
   \normalfont
+  \vspace{1em} 
   \begin{lstlisting}
@@
 'nhs-trust' says App isInstallable
-if App hasMet('final-app-approval'),
-   App isUsable.
+  if App hasMet('final-app-approval'), App isUsable.
 'nhs-trust' says 'igc' can-say App hasMet('final-app-approval').
 'nhs-trust' says Device canInstall(App)
-  if App isInstallable,
-     App isApprovedFor(Device).
+  if App isInstallable, App isApprovedFor(Device).
 'nhs-trust' says Employee:Manager can-say
@@
 \end{policyrule}
-Given that user's have privacy preferences expressible as permissions~\cite{lin_modeling_2014}, we might expect corporate policies to describe what permissions an app can have.
+We might expect corporate policies to describe what apps can be installed in terms of the apps functionality.
 This does not appear to be the case, however. 
@@
 
-\section{Worked Example}
+\section{Authorization Example}
 
@@
 Suppose an employee, \emph{Alice}, wished to get an app, \emph{\ttfamily com.microsoft.office}, installed on their device.
-To do so, Alice would have to convince the device that \lstinline{'nhs-trust' says 'alices-phone' canInstall('com.microsoft.office').}
+To do so, Alice would have to convince the device that:
+
+\begin{lstlisting}[frame=none]
+'nhs-trust' says 'alices-phone' canInstall('com.microsoft.office').
+\end{lstlisting}
+
 Alice wishes to use the app for business so to satisfy the policy Alice must collect the following statements:
@@
 Each may either give her the statement she needs or may give her additional rules.
-For example, the MIG and IGC may be happy to state their assertions (after a review).
+For example, the MIG and IGC may be happy to state their statements (after a review).
 When checking if the app is an App in \autoref{item:isapp}, the NHS trust may be instead inclined to delegate further.
@@
 Alice would then have to obtain additional statements if she wanted to prove this statement.
-As with SecPAL, all assertions should have a signature from their speaker proving they said the assertion.
+As with SecPAL, all statements should have a signature from their speaker proving they said the statement.
 Alternatively, the speaker could refuse to give the statement, either because they do not believe it to be true, or they cannot give an answer.
@@
 \section{BYOD Idioms}
+\label{sec:idioms}
 
@@
 All the policies contained more rules featuring delegation relationships than functionality restrictions.
-Restricting device functionality is tricky and important, but other aspects of BYOD policies are also worth attention.
+%Restricting device functionality is tricky and important, but other aspects of BYOD policies are also worth attention.
 
@@
 
-\AppPAL~requires an explicit speaker for each assertion.
+\AppPAL~requires an explicit speaker for each statement.
 Speakers can delegate to others by making a statement about what they \emph{can-say}.
@@
 For the \ac{HiMSS} policy, where the user states what they will do rather than the company stating what they must, the user is the primary speaker.
-All the policies describe multiple entities that might make assertions and delegate.
+All the policies describe multiple entities that might make statements and delegate.
 With \AppPAL~policies any speaker can delegate a decision to another speaker (with restrictions on re-delegation).
@@
 
-\section{Comparison with Existing MDM Software}
+%\section{Comparison with Existing MDM Software}
 
-It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}, both leading MDM packages~\cite{rob_smith_magic_2016}.
-Each of these tools supports enforcing and checking compliance policies. 
-They do not, however, use policy languages to specify policies; rather they supply canned policies and rules that admins can tweak.
+% It is interesting to compare \AppPAL~to existing MDM~tools such as IBM's MaaS360 and \ac{BES}, both leading MDM packages~\cite{rob_smith_magic_2016}.
+% Each of these tools supports enforcing and checking compliance policies. 
+% They do not, however, use policy languages to specify policies; rather they supply canned policies and rules that admins can tweak.
 
-Acknowledgements form a significant portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
-MaaS360 can associate users with devices and can assign them into groups.
-It can distribute electronic documents, but there are no means to track what policies a user should read and follow.
-Administrators could add users who have acknowledged a policy to a group.
-One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a semi-manual process for an administrator to set up and manage.
-One might imagine administrators adding users who have acknowledged some additional rules to new groups
-Similarly, with \ac{BES} can create users, groups and groups of devices, but there is no facility for managing what users should agree to automatically.
-We can create policies that apply to groups, but not create the groups and policies based on membership of groups.
-
-\AppPAL{} is more flexible than these MDM suites. 
-It allows a user to state which policies they have acknowledged and for policies to delegate the acknowledgement of the policy to the user.
-We can require that a user must acknowledge certain policies if they meet certain criteria.
-Once a statement is made, it can be imported into the \AppPAL~knowledge base and used in making other decisions as part of the policy.
-
-This lack of flexibility extends to trying to compose policies. 
-Consider the following rule taken from the SANS policy:
-\begin{policyrule}{SANS}
-Any non business-owned (that, is private) device must be able to connect to Company network MUST first be approved by technical personnel such as those from the Company IT department or desktop support.
-If allowed, privately-owned handheld devices MUST comply with this security policy and MUST be inventoried along with corporate handhelds, but identified as private.
+% Acknowledgements form a significant portion of the BYOD policies we looked at but are not addressed by the MDM software at all.
+% MaaS360 can associate users with devices and can assign them into groups.
+% It can distribute electronic documents, but there are no means to track what policies a user should read and follow.
+% Administrators could add users who have acknowledged a policy to a group.
+% One might imagine that users who have acknowledged some additional rules might be added to additional groups with stricter or more relaxed policies, but this is a semi-manual process for an administrator to set up and manage.
+% One might imagine administrators adding users who have acknowledged some additional rules to new groups
+% Similarly, with \ac{BES} can create users, groups and groups of devices, but there is no facility for managing what users should agree to automatically.
+% We can create policies that apply to groups, but not create the groups and policies based on membership of groups.
+% 
+% \AppPAL{} is more flexible than these MDM suites. 
+% It allows a user to state which policies they have acknowledged and for policies to delegate the acknowledgement of the policy to the user.
+% We can require that a user must acknowledge certain policies if they meet certain criteria.
+% Once a statement is made, it can be imported into the \AppPAL~knowledge base and used in making other decisions as part of the policy.
+% 
+% This lack of flexibility extends to trying to compose policies. 
+% Consider the following rule taken from the SANS policy:
+% \begin{policyrule}{SANS}
+% Any non business-owned (that, is private) device must be able to connect to Company network MUST first be approved by technical personnel such as those from the Company IT department or desktop support.
+% If allowed, privately-owned handheld devices MUST comply with this security policy and MUST be inventoried along with corporate handhelds, but identified as private.
 This is in order to prevent theft of corporate data with unmanaged handhelds (i.e.
 owner of device is not identified).
-\end{policyrule}
-To implement this using an MDM solution such as MaaS360 or \ac{BES} administrators can mark some devices as being privately owned.
-A group could be set up for approved privately-owned devices.
-MaaS360 allows the creation of groups based on search criteria, though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
-Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically and could pass them along for inventorying.
-Again the management of \emph{who} is \emph{technical personnel} is effectively manual, with optional searches to define groups.
-Once we find those devices that can connect we could add an access control rule to a group containing them to enforce the rule.
-
-In \AppPAL~the policy specifies all roles and relationships.
-We are free to use conformance to one policy as a conditional for applying a second, allowing dynamic groups.
-Roles such as being \emph{technical personnel} can be given to users holding other roles without requiring manual intervention.
-Furthermore, the \AppPAL~rule is presented in much the same form as the rule in the policy document: \emph{device} can do \emph{action} if \emph{checks} are satisfied.
-
-\begin{lstlisting}
-'company' says Device:D canConnectTo(Network:N)
-  if N isCompanyOwned,
-     D isPrivatelyOwned,
-     D isApproved,
-     D hasMet('security-policy').
-
-'company' says 'technical-personnel' can-say
-  Device:D isApproved.
-
-'company' says 'it-department' can-act-as 'technical-personnel'.
-'company' says 'desktop-support' can-act-as 'technical-personnel'.
-\end{lstlisting}
-
-\AppPAL~improves upon existing MDM tools by allowing sophisticated delegation relations and by providing a declarative language for expressing policies.
-The language gives greater flexibility to policy authors and allows them to write policies that depend on other policies rather than predefined settings and groups.
-It lets us track what users have agreed to, what their policies are, how they are specified, and how they are satisfied.
+% \end{policyrule}
+% To implement this using an MDM solution such as MaaS360 or \ac{BES} administrators can mark some devices as being privately owned.
+% A group could be set up for approved privately-owned devices.
+% MaaS360 allows the creation of groups based on search criteria, though the groups are not dynamic\footnote{The search would need to be re-run when new devices were added or a device's status changed.}.
+% Technical personnel could be given an administrator login to the MDM software to add devices to this group periodically and could pass them along for inventorying.
+% Again the management of \emph{who} is \emph{technical personnel} is effectively manual, with optional searches to define groups.
+% Once we find those devices that can connect we could add an access control rule to a group containing them to enforce the rule.
+% 
+% In \AppPAL~the policy specifies all roles and relationships.
+% We are free to use conformance to one policy as a conditional for applying a second, allowing dynamic groups.
+% Roles such as being \emph{technical personnel} can be given to users holding other roles without requiring manual intervention.
+% Furthermore, the \AppPAL~rule is presented in much the same form as the rule in the policy document: \emph{device} can do \emph{action} if \emph{checks} are satisfied.
+% 
+% \begin{lstlisting}
+% 'company' says Device:D canConnectTo(Network:N)
+%   if N isCompanyOwned,
+%      D isPrivatelyOwned,
+%      D isApproved,
+%      D hasMet('security-policy').
+% 
+% 'company' says 'technical-personnel' can-say
+%   Device:D isApproved.
+% 
+% 'company' says 'it-department' can-act-as 'technical-personnel'.
+% 'company' says 'desktop-support' can-act-as 'technical-personnel'.
+% \end{lstlisting}
+% 
+% \AppPAL~improves upon existing MDM tools by allowing sophisticated delegation relations and by providing a declarative language for expressing policies.
+% The language gives greater flexibility to policy authors and allows them to write policies that depend on other policies rather than predefined settings and groups.
+% It lets us track what users have agreed to, what their policies are, how they are specified, and how they are satisfied.
 
@@
 We have presented \AppPAL: an instantiation of SecPAL for BYOD policies.
-Using an \AppPAL~formalization of 5 BYOD policies we have identified several concerns that existing BYOD tools ignore.
+Using an \AppPAL~formalization of 5 BYOD policies we have identified that whilst delegation and acknowledgement form a large part of written BYOD policies, existing BYOD tools ignore them.
 BYOD policies contain delegation and trust relationships that define who is responsible for making different decisions in a company.
@@
 
-Future work will incorporate new BYOD and static analysis tools into \AppPAL. 
-This will allow us to use them as part of policy checks.
-Whilst earlier work on SecPAL based languages~\cite{becker_secpal:_2010,moritz_y_becker_secpal:_2009} has described how to distribute assertions there has not been a well-defined protocol for doing so.
-Future work will also work on tying down the distribution mechanisms and defining the distribution protocol.
+\AppPAL~improves upon existing MDM tools by allowing sophisticated delegation relations and by providing a declarative language for expressing policies.
+The language gives greater flexibility to policy authors and allows them to write policies that depend on other policies rather than predefined settings and groups.
+It lets us track what users have agreed to, what their policies are, how they are specified, and how they are satisfied.
+
+Acknowledgements were used in all the policies, but were not a part of \ac{MDM} tools.
+A purely speculative explanation for this might be that the people using the \ac{MDM} software (the IT department) do not care about the acknowledgements, and that another department (HR perhaps) are responsible for tracking what corporate policies employees have agreed to and have their own methods for dealing with that.
+Future work will aim to further explore how these acknowledgements are used within a company and how to manage them in a practical manner.
+
+Related systems, such as \ac{GSN} described in \autoref{sec:related}, use a graphical notation.
+Whilst SecPAL-based languages are designed to be readable, diagrams can help make authors write policies and auditors understand them.
+Future work will look at extending SecPAL's notation to create such diagrams.
 

### 2016-12-20 paper.tex
submitted version

@@
 \acrodef{GSN}{Goal Structuring Notation}
+
+\usepackage{enumitem}
 
@@
   BYOD policies are informally specified using natural language.
-  We show how the \AppPAL~language can help reduce ambiguity in BYOD policies and link the specification of a BYOD policy to its implementation.
-  This paper identifies a BYOD idiom that is ignored by existing MDM tools.
-  We explore the delegation relationships and concerns in 5 BYOD policies, and describe how \AppPAL might be used to implement them using existing tools.
+  We show how the \AppPAL~language can help reduce ambiguity in 5 BYOD policies and link the specification of a BYOD policy to its implementation.
+  Using a formalisation of the 5 policies written in \AppPAL, we make comparisons between them, and explore the delegation relationships within them.
+  We identify that whilst policy acknowledgement is a key part of all 5 policies, they are not managed by existing MDM tools.
 \end{abstract}
@@
 Employees bring their own devices to work.
-In the past employees might have had a dedicated company device.
-Today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016} and, in some fields,
-  85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}.
+In the past employees might have had a dedicated company device; 
+  but today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016}. 
+In some fields, 85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}.
 Controlling employee's devices is a challenge for IT departments.
-Failure to do so can lead to cause employees to leak confidential information and the waste of company resources.
-Unfortunately the companies have limited control over the devices used to access them.
+Failure to manage devices can lead to employees accidentally leaking confidential information and the waste of company resources.
+Unfortunately the companies have limited control over the devices inside their networks if they do not own them.
 
-One solution to controlling devices is to require users to agree to follow policies.
-They often take the form of a user agreement, written in natural language, which describes how the devices should be used and configured.
-Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,hp_byod_????,cesg_byod_2015}.
+One solution to controlling devices is requiring users agree to follow policies.
+They often take the form of a user agreement, written in natural language, which describes how devices should be used and configured.
+Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,cesg_byod_2015}.
 On top of user agreements, companies may also use \ac{MDM} software which can help enforce policies.
-\ac{MDM} software can configure a device's security settings, and add 
+\ac{MDM} software can configure a device's security settings, and add provide security APIs.
 But the use of \ac{MDM} software does not guarantee compliance.
-One survey from a leading MDM vendor found over 50\% of companies with their \ac{MDM} software still had devices that do not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.
+One survey from a leading MDM vendor found over 50\% of companies with their \ac{MDM} software still had devices that did not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.
 Reasons for non-compliance included out-of-date policies and employees tampering with the \ac{MDM} software.
@@
 BYOD policies are becoming more intricate.
-There has been much work looking at developing the \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
+Prior work has looked at developing \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
 Implementing a policy is only part of the problem, however.
@@
 An antivirus or \ac{MDM} program may be used to \emph{implement} part of the policy.
-But it is the policy, however, that \emph{specifies} which software to use and when. 
+But it is the policy that \emph{specifies} which software to use and when. 
 There is no automatic way to check how the policy has been implemented and by what.
 
+Companies lack visibility as to how they implement their policies.
+When considering what tools a company may use Morrow notes \emph{``particularly with the BYOD trend IT professionals do not know if anti-virus software is installed or if it's current''}~\cite{morrow_byod_2012}.
+Even when devices can implement policies correctly, it is hard to configure devices that are not owned by the company~\cite{tokuyoshi_security_2013}.
+Our work aims to address these problems directly: by using formal languages we can link the policy to the implementation.
+
 To describe the policies we came up with \AppPAL{}: 
-  a formal language for linking policies to the tools used to implement them and distributing decisions about apps.
+  a formal language for linking policies to the tools used to implement them and distributing decisions.
 Using a formalization of five BYOD policies written using \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
@@
 We look at what decisions and trust relationships used in BYOD policies.
-These are BYOD idioms that are capture decisions and frequently seen requirements in BYOD policies.
+We identify BYOD idioms that capture frequently seen decisions in BYOD policies.
 These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.
-
-Companies lack visibility as to how they implement their policies.
-When considering what tools a company may use Morrow notes \emph{``particularly with the BYOD trend IT professionals do not know if anti-virus software is installed or if it's current''}~\cite{morrow_byod_2012}.
-Even when devices can implement policies correctly, it is hard to properly configure devices that are not owned by the company~\cite{tokuyoshi_security_2013}.
-Our work aims to address these problems directly: by using formal languages we can link the policy to the implementation.
 
@@
 A company can use any \ac{MDM} tool, curated app store or user agreement to enforce their policy.
-\AppPAL~links the specification of the policy to its implementation, giving visibility into exactly how a policy is implemented and satisfied.
+\AppPAL~links the specification of the policy to its implementation, showing exactly how a policy is implemented and giving a rigorous means to enforce it.
 
-We show how \AppPAL~can be used to implement these policies and describe precisely the different trust relationships.
+We show how \AppPAL~can be used to encode these policies and describe precisely the different trust relationships.
 \AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
-We have found, however, that SecPAL is a useful tool for describing other policies surround the mobile ecosystem as well~\cite{hallett_specifying_2016}.
-Our AppPAL implementation can also be easily extended to support new types of policies.
-It also gives us access to tooling we have developed to can check AppPAL policies for completeness and redundant statements.
-For this work additional tooling was developed to help visualise policies and describe policies contents.
+We have found, however, that SecPAL is a useful tool for describing other policies surrounding mobile ecosystems~\cite{hallett_specifying_2016}.
+Our AppPAL implementation can be easily extended to support new types of policies.
+It also gives us access to tooling we have developed to check AppPAL policies for completeness and redundant statements.
+For this work additional tooling was developed to help visualise policies and describe their contents.
 This was helpful for making comparisons between policies and checking our formalisation for mistakes.
@@
 In summary, our work makes the following contributions:
-
-\begin{itemize}
+\begin{itemize}[topsep=0pt]
   \item We present a formalisation of five different BYOD policies in \AppPAL: a new instantiation of SecPAL for describing BYOD policies (\autoref{sec:idea}).
@@
 
-Martinelli~\etal{}'s looks at creating a dynamic permissions manager, called UC-Droid.
+Martinelli~\etal{}'s work looks at creating a dynamic permissions manager, called UC-Droid.
 Their tool can alter what an app's Android permissions are at run time based on policies~\cite{martinelli_enhancing_2016}.
-The tool allows companies to reconfigure the apps depending on whether the employee is at work, in a secret lab, or working out-of-hours.
-These kinds of policies are more configurable than the geofenced based policies existing MDM packages provide.
+The tool allows companies to reconfigure their apps depending on whether the employee is at work, in a secret lab, or working out-of-hours.
+These kinds of policies are more configurable than the geofenced based policies some \ac{MDM} tools provide.
 Other work has looked at enforcing different policies based on what roles an employee holds~\cite{costantino_towards_2013}.
@@
 The store ensures apps meet policies through a combination of static analysis and app rewriting with dynamic enforcement.
-Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on the JVM state.
+Their policies are low level, based on ConSpec~\cite{aktug_conspec_2008}, allowing checks based on Dalvik VM's state.
 Using their tool, they implemented parts of a NATO Communications and Information Agency policy relating to personal networks and data management~\cite{armando_developing_2016}.
@@
 
-Other work has looked at using app wrapping (where an app is recompiled to use guarded APIs) to enforce more general policies.
-Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping as a possible way to enforce policies.
+Tools, such as Dr.~Android and Mr.~Hyde~\cite{jeon_dr._2012} and Aurasium~\cite{xu_aurasium:_2012}, have suggested app wrapping (where an app is recompiled to use guarded APIs) as a possible way to enforce policies.
 App rewriting has the advantage that the device's underlying OS needn't be modified as the apps are changed at the source code level.
@@
 Industrial safety cases are often described in natural language, which can be ambiguous and unclear.
-\ac{GSN}~\cite{tim_kelly_goal_2004} is one approach to make the safety cases explicit.
-It is a graphical formal notation that lets engineers argue that a system is safe by linking top-level goals, to sub-goals to the arguments made for a system's safety.
-Similarly work developing a formal language for specifying how medical staff should collaborate in a healthcare scenario~\cite{papapanagiotou_formal_2014} again helps clarify how roles are filled in a medical context, on the basis of staff and different healthcare providers.
+\ac{GSN}~\cite{kelly_goal_2004} is one approach to make the safety cases explicit.
+It is a graphical formal notation that lets engineers argue that a system is safe by linking safety goals to the arguments made for a system's safety.
+Similarly, work developing a formal language for specifying how medical staff should collaborate in a healthcare scenario~\cite{papapanagiotou_formal_2014} again helps clarify how roles are filled in a medical context on the basis of staff and different healthcare providers.
 
-It is interesting to examine how leading~\cite{rob_smith_magic_2016} MDM~tools such as IBM's MaaS360, MobileIron and \ac{BES}, enforce \emph{BYOD} policies.
-Each of these tools supports enforcing and checking compliance policies. 
+It is interesting to examine how leading~\cite{rob_smith_magic_2016} MDM~tools such as IBM's MaaS360, or \ac{BES}, enforce \emph{BYOD} policies.
+These tools support enforcing and checking compliance policies. 
 They do not, however, use policy languages to specify policies; rather they provide a limited number of checkboxes that admins can tweak (an excerpt of a policy from MaaS360 is shown in \autoref{fig:maaspolicy}).
@@
 Some support app wrapping, which enables them to encrypt app data locally, use a VPN within the app, or prevent apps not being used when the device isn't compliant.
-Because the policies are inflexible and tightly coupled to the device's OS, intervention by an administrator is often required.
+But because the policies are inflexible and tightly coupled to the device's OS, intervention by an administrator is often required.
 Whilst MDM software is good at configuring devices, selecting which policies to apply is typically a manual process performed by an administrator.
@@
 
-As mobile devices have become more common in the workplace, BYOD policies have been written to help control the devices.
+As mobile devices have become more common in the workplace, BYOD policies have been written to help control them.
 Part of their policies are prescriptive:  if you configure your device in this way, you will mitigate that threat.
@@
 
-A company could make use of an \ac{MDM} program.
-Some can use geofencing to apply a policies in the area around a lab.
+A company could use an \ac{MDM} program to enforce this.
+Some \ac{MDM} tools can use geofencing to apply a policies in the area around a lab.
 Techniques like this would implement the recommendation within the rule, but the rule itself contains more than just configuration.
-It talks of \emph{restricted environments} decided by \emph{company risk analysis}; but how is this communicated to the device?
+It talks of \emph{restricted environments} decided by \emph{company risk analysis}.
+How is this communicated to the device?
 Does it access the list of restricted environments once from a server, are they fixed or can a device decide them for itself?
@@
   and use automatic-tools to uncover problematic aspects of the policy~\cite{hallett_specifying_2016}.
-Continuing with the example above, we can write this in \AppPAL~as:
+Continuing with the example above, we can encode this in \AppPAL~as:
 
@@
 
-After checking a policy written in \AppPAL, we generate a proof tree that shows how the policy was satisfied.
+After checking the policy we generate a proof tree that shows how the policy was satisfied.
 These proof trees not only show how the policy was followed but also provide an audit trail.
@@
 SecPAL was developed as a distributed access control language~\cite{becker_secpal:_2010}.
-The SecPAL language is designed to be have a clear readable syntax, and intuitive semantics.
+It is designed to be have a clear readable syntax, and intuitive semantics.
 It is also designed to be extensible, which makes it ideal for extending to create new languages.
@@
 All five of the policies make use of acknowledgements.
-The use of acknowledgements could be because enforcing the rules in a policy through technical means is undesirable. 
+The use of an acknowledgements could be because enforcing that rule in a policy through technical means is undesirable. 
 It could indicate policy authors care more that the subjects are aware of the rules than they do for rigorous enforcement.
@@
   \normalfont
-  \vspace{1em} 
   \begin{lstlisting}
@@
   Alice can use a SecPAL inference tool (such as AppPAL\footnote{https://github.com/apppal/libapppal}) to check the policy has been satisfied.
-As well as making the decision we can output a proof tree.% (such as \autoref{fig:proof}).
-The generated proof lets auditors review how any decision was made, and verify the decision-making process.
+%As well as making the decision we can output a proof tree.% (such as \autoref{fig:proof}).
+The generated proof from the tool lets auditors review how the decision was made, and verify the decision-making process.
 
@@
 
-When examining the policies, we noticed two particular idioms reappearing: acknowledgements and delegation.
+When examining the policies, we noticed two particular idioms in many policies: acknowledgements and delegation.
 We describe both idioms in greater detail, and show how they can be implemented in \AppPAL{}, below.
@@
 
-All the policies looked at require their subjects to be aware and acknowledge certain rules or policies, 
-  or that the company may perform certain actions.
+All the policies we looked at require their subjects to be aware and acknowledge certain rules or policies, 
+  and that the company may perform certain actions.
 For example, the NHS and \ac{HiMSS} policies state that the organisation will wipe devices remotely to protect confidential information a user loses their device.
@@
 Whilst SecPAL-based languages are designed to be readable, diagrams can help make authors write policies and auditors understand them.
-Future work will look at extending SecPAL's notation to create such diagrams.
+Future work will look at extending SecPAL's notation to create such diagrams and further aid readability.
 
 %\bibliographystyle{plain}
+\newpage
 \printbibliography

### 2017-03-01 paper.tex
Sorted out typos from review.

@@
 
-One solution to controlling devices is requiring users agree to follow policies.
+One solution to controlling devices is requiring users follow policies.
 They often take the form of a user agreement, written in natural language, which describes how devices should be used and configured.
@@
 On top of user agreements, companies may also use \ac{MDM} software which can help enforce policies.
-\ac{MDM} software can configure a device's security settings, and add provide security APIs.
+\ac{MDM} software can configure a device's security settings, and add security APIs.
 But the use of \ac{MDM} software does not guarantee compliance.
@@
 Our formalizations pick out the common concerns and trust relationships in these policies.
-We look at what decisions and trust relationships used in BYOD policies.
+We look at what decisions and trust relationships are used in BYOD policies.
 We identify BYOD idioms that capture frequently seen decisions in BYOD policies.
@@
 \AppPAL~is not designed to replace existing static and dynamic analysis and enforcement tools.
-A company might use multiple different tools, app stores, and even contractual agreements with employees to enforce their policies.
+A company might use multiple tools, app stores, and even contractual agreements with employees to enforce their policies.
 We aim to help clarify the meaning of ambiguous natural language policy documents, and provide a rigorous means for following them.
@@
 \AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
-We have found, however, that SecPAL is a useful tool for describing other policies surrounding mobile ecosystems~\cite{hallett_specifying_2016}.
+SecPAL is also a useful tool for describing other policies surrounding mobile ecosystems~\cite{hallett_specifying_2016}.
 Our AppPAL implementation can be easily extended to support new types of policies.
@@
       \AppPAL~statements              & 71           & 21          & 58          & 10                                & 39          \\
-      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large nmuber of rules that whilst marked as rules are infact just descriptions of the document.
+      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large number of rules that whilst marked as such are infact just descriptions of the document.
 All the policy rules that described restrictions or relationships were implemented in \AppPAL{}.} & 22 (88$\%$) \\
@@
 When there is a (usually separate) set of rules and concerns employees should be aware of acknowledgements are used.
-The company may not have wish to enforce these separate rules automatically, however.
-For instance, a company may have an ethical policy that says employees should not use devices for criminal purposes.
+The company may not wish to enforce these separate rules automatically, however.
+For instance, a company may have an ethics policy that says employees should not use devices for criminal purposes.
 The company is not interested in, or capable of, defining what is criminal.

### 2017-03-01 paper.tex
the rest of the tweaks requested from the reviewers

@@
 \usepackage{tablefootnote}
-\usepackage[backend=bibtex,firstinits=true]{biblatex}
+\usepackage[backend=bibtex,firstinits=true,sorting=none]{biblatex}
 \addbibresource{paper.bib}
@@
 \title{Capturing Policies for BYOD}
-\author{Joseph Hallett and David Aspinall}
+\author{Joseph Hallett\thanks{Work supported by EPSRC App Guarden grant (EP/K032666/1); and the Alan Turing Institute.} and David Aspinall}
 \institute{School of Informatics, University of Edinburgh}
@@
 Industrial safety cases are often described in natural language, which can be ambiguous and unclear.
-\ac{GSN}~\cite{kelly_goal_2004} is one approach to make the safety cases explicit.
+\ac{GSN}~\cite{tim_kelly_goal_2004} is one approach to make the safety cases explicit.
 It is a graphical formal notation that lets engineers argue that a system is safe by linking safety goals to the arguments made for a system's safety.

### 2017-03-06 paper.tex
davids alterations

@@
 \usepackage{amsmath,amssymb,mathtools}
-\usepackage{hyperref}
 \usepackage{booktabs}
@@
 \newcommand{\AppPAL}[0]{SP4BYOD}
+
+\usepackage{hyperref}
+\renewcommand{\sectionautorefname}{Section}
+\renewcommand{\subsectionautorefname}{Subsection}
+\renewcommand{\figureautorefname}{Figure}
+\renewcommand{\tableautorefname}{Table}
+\renewcommand{\itemautorefname}{Item}
 
@@
 \title{Capturing Policies for BYOD}
-\author{Joseph Hallett\thanks{Work supported by EPSRC App Guarden grant (EP/K032666/1); and the Alan Turing Institute.} and David Aspinall}
+\author{Joseph Hallett and David Aspinall\thanks{Work supported by EPSRC App Guarden grant (EP/K032666/1) and the Alan Turing Institute.}}
 \institute{School of Informatics, University of Edinburgh}
@@
   Using a formalisation of the 5 policies written in \AppPAL, we make comparisons between them, and explore the delegation relationships within them.
-  We identify that whilst policy acknowledgement is a key part of all 5 policies, they are not managed by existing MDM tools.
+  We identify that whilst policy acknowledgement is a key part of all 5 policies, this is not managed by existing MDM tools.
 \end{abstract}
@@
 Employees bring their own devices to work.
-In the past employees might have had a dedicated company device; 
-  but today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016}. 
+In the past employees might have had a dedicated company device. 
+Today around 70\% of companies have a BYOD scheme~\cite{schulze_byod_2016}. 
 In some fields, 85\% of staff use their personal devices to look up work-sensitive information~\cite{patel_uk_2015}.
 Controlling employee's devices is a challenge for IT departments.
-Failure to manage devices can lead to employees accidentally leaking confidential information and the waste of company resources.
-Unfortunately the companies have limited control over the devices inside their networks if they do not own them.
+Failure to manage devices can lead to employees accidentally leaking confidential information.
+Unfortunately companies have limited control over the devices inside their networks if they do not own them.
 
-One solution to controlling devices is requiring users follow policies.
-They often take the form of a user agreement, written in natural language, which describes how devices should be used and configured.
+One solution to controlling devices is requiring users follow BYOD policies.
+A BYOD policy takes the form of a user agreement, written in natural language, which describes how devices should be used and configured.
 Various guides are available for companies wishing to implement a policy from governments, standards bodies, and organizations seeking to advise~\cite{nicholas_r._c._guerin_security_2008,souppaya_guidelines_????,cesg_byod_2015}.
 On top of user agreements, companies may also use \ac{MDM} software which can help enforce policies.
-\ac{MDM} software can configure a device's security settings, and add security APIs.
+\ac{MDM} software can configure a device's security settings, and add security APIs, helping enforce some aspects of the policies.
 But the use of \ac{MDM} software does not guarantee compliance.
-One survey from a leading MDM vendor found over 50\% of companies with their \ac{MDM} software still had devices that did not comply with their policies~\cite{mobileiron_security_labs_q4_2015}.
-Reasons for non-compliance included out-of-date policies and employees tampering with the \ac{MDM} software.
+One survey from a leading MDM vendor found over 50\% of companies using their \ac{MDM} software still had devices that did not comply with the policy~\cite{mobileiron_security_labs_q4_2015}.
+Reasons for non-compliance included out-of-date \ac{MDM} configurations that hadn't been updated, and employees tampering with the \ac{MDM} software.
 
 BYOD policies are becoming more intricate.
-Prior work has looked at developing \ac{MDM} software to enforce aspects of BYOD policies~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
-Implementing a policy is only part of the problem, however.
+Prior work has looked at developing \ac{MDM} software to enforce some aspects~\cite{costantino_towards_2013,martinelli_enhancing_2016,armando_enabling_2014}.
+The \ac{MDM} encoding of a policy is only part of the problem, however.
 BYOD policies are specified informally using natural language, and they contain more than just access control decisions.
-These policies describe trust relationships inside the company between the IT departments, users, and HR each who may be delegated to make decisions and provide rules.
+They describe trust relationships inside the company between IT departments, users, and HR, who each may be delegated to provide rules and make decisions.
 Policies contain rules that require employees to acknowledge risks, and regulations.
@@
 
-Companies lack visibility as to how they implement their policies.
+Companies lack visibility as to how well they implement their policies.
 When considering what tools a company may use Morrow notes \emph{``particularly with the BYOD trend IT professionals do not know if anti-virus software is installed or if it's current''}~\cite{morrow_byod_2012}.
@@
 
-To describe the policies we came up with \AppPAL{}: 
-  a formal language for linking policies to the tools used to implement them and distributing decisions.
+To describe the BYOD policies we present \AppPAL{}: 
+  a formal language for linking policies to the tools used to implement them, and distributing decisions to relevant parties.
 Using a formalization of five BYOD policies written using \AppPAL~we identify different idioms and common delegation patterns present in BYOD policies.
@@
 We identify BYOD idioms that capture frequently seen decisions in BYOD policies.
-These give a guide for where future work implementing BYOD tools should focus their efforts to cover more aspects of policies.
+These give a guide for where future work implementing \ac{MDM} tools should focus their efforts to cover more aspects of policies.
 
 \AppPAL~is not designed to replace existing static and dynamic analysis and enforcement tools.
-A company might use multiple tools, app stores, and even contractual agreements with employees to enforce their policies.
+A company might use multiple tools, app stores, or contractual agreements with employees to enforce their policies.
 We aim to help clarify the meaning of ambiguous natural language policy documents, and provide a rigorous means for following them.
@@
 
-We show how \AppPAL~can be used to encode these policies and describe precisely the different trust relationships.
-\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016}.
+We show how \AppPAL~can be used to encode policies and describe precisely the different trust relationships.
+\AppPAL~is an instantiation of the SecPAL authorization language~\cite{becker_secpal:_2010} for mobile device policies and implemented atop of AppPAL~\cite{hallett_apppal_2016},
+  which adds mobile device specific predicates to SecPAL, for example, capturing app permissions.
 SecPAL is also a useful tool for describing other policies surrounding mobile ecosystems~\cite{hallett_specifying_2016}.
-Our AppPAL implementation can be easily extended to support new types of policies.
-It also gives us access to tooling we have developed to check AppPAL policies for completeness and redundant statements.
+Our \AppPAL{}~implementation can be easily extended to support new types of policies.
+It also gives us access to tooling we have developed to check \AppPAL{}~policies for completeness and redundant statements.
 For this work additional tooling was developed to help visualise policies and describe their contents.
@@
       \AppPAL~statements              & 71           & 21          & 58          & 10                                & 39          \\
-      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edindiburgh policy contains a large number of rules that whilst marked as such are infact just descriptions of the document.
+      Policy coverage                 & 33 (100$\%$) & 14 (93$\%$) & 40 (71$\%$) & 10 (100\%)\tablefootnote{The Edinburgh policy contains a large number of rules that whilst marked as such are in fact just descriptions of the document.
 All the policy rules that described restrictions or relationships were implemented in \AppPAL{}.} & 22 (88$\%$) \\
